---
title: "R Notebook"
output: html_notebook
---

```{r}

setwd('V:/dkmarri/Breast_Cancer_IMC_Data/R_Codes')
```

```{r}
library(imcRtools)
library(cytomapper)
library(tidyverse)
library(EBImage)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(dplyr)
library(future)
library(progress)
library(Matrix)
library(RANN)
library(reticulate)
library(igraph)
library(tidyr)
library(dittoSeq)
library(scater)
library(patchwork)
library(cowplot)
library(viridis)
library(parallel) 
library(spdep)
library(spatialHeatmap)
library(igraph)
library(RANN)
library(data.table)
library(Matrix)
library(parallel)
library(furrr)
library(scran)
library(batchelor)
library(BiocParallel)
library(scater)  # For PCA, if not already installed
library(SummarizedExperiment)
library(sva)
library(harmony)
library(RColorBrewer)
library(BiocSingular)
library(scales)
library(ggraph)
library(ComplexHeatmap)
library(circlize)
```

#### Import the two arms dataset

```{r}
#Import chemotherapy dataset

spe_chemo <- readRDS("saved_data/spe_chemo_Spatial_Graph.rds")
spe_chemo

```

```{r}
#Imort the chemo_and_immuno dataset

spe_chemo_immu <- readRDS("saved_data/spe_chemo_immu_Spatial_Graph.rds")
spe_chemo_immu
```

```{r}

sce <-readRDS("saved_data/15_10_2024_spe_seurat.Rds")
sce

```

#### Plot the mean expression in each Arm of treatment

```{r}

features <- rownames(spe_chemo)
features1 <- c('CD4','Ki67','CD31','TCF1',"Helios","CD8", "OX40", "CD11c",         "CD3",'CA9' )



```

```{r}

RidgePlot(sce, features = features1, ncol = 2)
```

```{r}

# Function to apply the transformation
transform_expression <- function(x) {
  abs(log2(x + 1e-4) - log2(11e-5))
}
```

```{r}
# Assuming transform_expression is your custom normalization function
normalized_data <- transform_expression(GetAssayData(sce, assay = "originalexp", slot = "counts"))

# Convert normalized_data to dgCMatrix if necessary
if (!inherits(normalized_data, "matrix")) {
  library(Matrix)
  normalized_data <- as(normalized_data, "dgCMatrix")
}

# Set the normalized data back into the Seurat object using the layer argument
sce <- SetAssayData(sce, layer = "data", new.data = normalized_data, assay = "originalexp")

sce
```

```{r}
RidgePlot(sce, features = features1, ncol = 2)
```

```{r}

```

```{r}

```

```         
```

```{r}

library(tidyverse)
library(kableExtra)
```

```{r}
daughter_to_parent <- c(
  "Fibroblasts" = "Fibroblasts",
  "DCs" = "Other",
  "CA9^+" = "Cancer Cells",
  "PDPN^+Stromal" = "Fibroblasts",
  "M2 Mac" = "Macrophages",
  "Myofibroblasts" = "Fibroblasts",
  "CD79a^+Plasma" = "Other",
  "CK8/18^{med}" = "Cancer Cells",
  "CD56^+NK" = "Other",
  "CD4^+TCF1^+T" = "Effector T Cells",
  "panCK^{med}" = "Cancer Cells",
  "Treg" = "Regulatory T Cells",
  "CD8^+T" = "Effector T Cells",
  "CD8^+TCF1^+T" = "Effector T Cells",
  "CD8^+PD1^+T_{Ex}" = "Effector T Cells",
  "Endothelial" = "Endothelial Cells",
  "CD4^+PD1^+T" = "Effector T Cells",
  "CD8^+GZMB^+T" = "Effector T Cells",
  "PD-L1^+APCs" = "Macrophages",
  "CD15^{+}" = "Other",
  "PD-L1^{+}GZMB^{+}" = "Other",
  "pH2AX^{+}DSB" = "Other",
  "Vimentin^{+}EMT" = "Cancer Cells",
  "CK^{hi}GATA3^{+}" = "Cancer Cells",
  "CK^{lo}GATA3^{+}" = "Cancer Cells",
  "Helios^{+}" = "Regulatory T Cells",
  "Basal" = "Cancer Cells",
  "Apoptosis" = "Other",
  "Neutrophils" = "Other",
  "CD20^+B" = "Other",
  "PD-L1^{+}IDO^{+}" = "Other",
  "PD-L1^+IDO^+APCs" = "Macrophages",
  "AR^{+}LAR" = "Cancer Cells",
  "TCF1^{+}" = "Effector T Cells",
  "MHC I&II^{hi}" = "Other",
  "CD56^{+}NE" = "Other",
  "CA9^{+}Hypoxia" = "Cancer Cells"
)
```

```{r}

df <- data.frame(
  Daughter = names(daughter_to_parent),   # Named elements (daughter)
  Parent = unname(daughter_to_parent)     # Unnamed elements (parent)
)

```

```{r}

# Group by 'Parent' and collapse 'Daughter' entries
New_df_grouped <- df %>%
  group_by(Parent) %>%
  summarise(Daughters = paste(Daughter, collapse = ", ")) %>%
  arrange(Parent)   # Arrange alphabetically by Parent

# Print the table
print(New_df_grouped)
```

```{r}

parent_lookup <- setNames(New_df_grouped$Parent, New_df_grouped$Daughters)

# Function to match labels to parents
match_to_parent <- function(label) {
  for (daughters in names(parent_lookup)) {
    if (label %in% unlist(strsplit(daughters, ", "))) {
      return(parent_lookup[daughters])
    }
  }
  return(NA)  # Return NA if no match found
}

# Create a new data frame with cell names and matched parent cells
new_metadata <- data.frame(
  cell = colnames(sce),
  ParentCell = sapply(sce$Label, match_to_parent)
)

# Add the new metadata to the Seurat object
sce <- AddMetaData(sce, metadata = new_metadata$ParentCell, col.name = "ParentCell")

# Check the results
table(sce$ParentCell, useNA = "ifany")

```

```{r}
# feature in each cluster. The color represents the average expression level
DotPlot(sce, features = features,cols = c("blue", "red"), dot.scale = 8, group.by = 'ParentCell') + RotatedAxis()
```

```{r}
# Replace 'celltype'/ with the actual column name containing cell type information
Idents(sce) <- sce$ParentCell

# Check again
print(table(Idents(sce)))
```

#### Find all the highest expressed biomarkers in each cell type

```{r}

#  Find markers for each cell type
all_markers <- FindAllMarkers(sce, 
                              only.pos = TRUE, 
                              min.pct = 0.60, 
                              logfc.threshold = 0.25)
```

```{r}

top_markers <- all_markers %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) %>%
  ungroup()
```

```{r}

# Create the DotPlot
dot_plot <- DotPlot(sce, 
                    features = unique(all_markers$gene), 
                    cols = c("lightgrey", "blue"),
                    dot.scale = 7) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top Protein Biomarkers Across Cell Types")

# Display the plot
print(dot_plot)
```

```{r}

avg_expression <- AggregateExpression(sce, 
                                      features = unique(all_markers$gene),
                                      group.by = "ParentCell",
                                      slot = "data",
                                      assays = "originalexp")

# Get genes with average expression > 1
high_expr_genes <- rownames(avg_expression$originalexp)[rowMeans(avg_expression$originalexp) > 1]

```

```{r}

dot_plot <- DotPlot(sce, 
                    features = high_expr_genes, 
                    group.by = 'ParentCell',
                    cols = c("lightgrey", "blue"),
                    dot.scale = 7) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top Protein Biomarkers Across Cell Types")

# Display the plot
print(dot_plot)
```

```{r}

```

```{r}

# Group by cell types and extract the proteins for each cell type

Significant_Markers <- top_markers %>%
  group_by(cluster) %>%
  summarize(proteins = paste(unique(gene), collapse = ", "))

Significant_Markers <-  as.data.frame(Significant_Markers)

Significant_Markers


```

```{r}
# Save the sce data into a seurat object

SaveSeuratRds(sce, "saved_data/15_10_2024_spe_seurat.Rds")

```

#### Transform both the chemotherapy and the immuno+chemotherapy dataset using the arcsinh transformation.

```{r}
# Function to apply the transformation
transform_expression <- function(x) {
  abs(log2(x + 1e-4) - log2(11e-5))
}
```

```{r}

# transform the chemotherpy dataset
# Get the counts data
counts_data <- assay(spe_chemo, "counts")

# Apply the transformation function
transformed_data <- apply(counts_data, c(1,2), transform_expression)

# Add the transformed data as a new assay
assay(spe_chemo, "transformed_counts") <- transformed_data
```

```{r}

spe_chemo
```

```{r}
# transform the chemotherpy+immunotherapy  dataset
# Get the counts data

counts_data1 <- assay(spe_chemo_immu, "counts")

# Apply the transformation function
transformed_data <- apply(counts_data1, c(1,2), transform_expression)

# Add the transformed data as a new assay
assay(spe_chemo_immu, "transformed_counts") <- transformed_data
```

```{r}

spe_chemo_immu
```

```{r}

parent_lookup <- setNames(New_df_grouped$Parent, New_df_grouped$Daughters)

# Function to match labels to parents
match_to_parent <- function(label) {
  for (daughters in names(parent_lookup)) {
    if (label %in% unlist(strsplit(daughters, ", "))) {
      return(parent_lookup[daughters])
    }
  }
  return(NA)  # Return NA if no match found
}

# Create a new data frame with cell names and matched parent cells
new_metadata <- data.frame(
  cell = colnames(spe_chemo),
  ParentCell = sapply(spe_chemo$Label, match_to_parent)
)


# Add the new metadata to the Seurat object
colData(spe_chemo)$ParentCell <- new_metadata$ParentCell

unique(spe_chemo$ParentCell)

```

```{r}

# Create a new data frame with cell names and matched parent cells for Chemo_immu
new_immu_metadata <- data.frame(
  cell = colnames(spe_chemo_immu),
  ParentCell = sapply(spe_chemo_immu$Label, match_to_parent)
)


# Add the new metadata to the Seurat object
colData(spe_chemo_immu)$ParentCell <- new_immu_metadata$ParentCell

unique(spe_chemo_immu$ParentCell)
```

```{r}

```

#### Get the protein biomarkers highly expressed in each arm of treatment

##### For chemotherapy

```{r}





## aggregate by cell type
celltype_mean <- aggregateAcrossCells(as(spe_chemo, "SingleCellExperiment"),  
                     ids = spe_chemo$ParentCell, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = features)



# Step 3: Ensure cell type metadata is correctly assigned
# Use unique 'spe$Label' values for cell type annotations
colData(celltype_mean)$celltype <- factor(unique(spe_chemo$ParentCell))

# Step 4: Count the number of cells per aggregated cell type
colData(celltype_mean)$ncells <- as.numeric(table(spe_chemo$ParentCell)[unique(spe_chemo$ParentCell)])

# Step 5: Plot the heatmap using dittoHeatmap
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scale = "row",                   # Scale by row (gene-wise)
             heatmap.colors = colorRampPalette(c("blue", "white", "red"))(100),    # Use viridis color scale
             annot.by = c("celltype", "ncells"),
             fontsize_row = 7,                 # Adjust the font size for type markers
                            # Show all color breaks for cell types
             show_legend = TRUE,
             legend.position = "right", 
             main = "Heatmap for celltypes and their expressed biomarkers for spe_chemotherapy")  # Annotate by 'celltype' and 'ncells'
```

```{r}

# Extract expression data
expression_data <- assay(celltype_mean, "exprs")
```

```{r}

# Function to identify highly expressed proteins
identify_high_expression <- function(expression_vector, threshold = 0.90) {
  quantile_threshold <- quantile(expression_vector, probs = threshold)
  return(expression_vector > quantile_threshold)
}


# Apply the function to each row (protein)
high_expression_matrix <- t(apply(expression_data, 1, identify_high_expression))

```

```{r}

# Convert to a data frame
high_expression_df <- as.data.frame(high_expression_matrix)

# Add protein names as a column
high_expression_df$Protein <- rownames(high_expression_df)
```

```{r}

# Reshape the data frame
spe_chemotherapy_result_df <- high_expression_df %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "IsHighlyExpressed") %>%
  filter(IsHighlyExpressed) %>%
  select(-IsHighlyExpressed)


# Sort the dataframe
spe_chemotherapy_result_df <- spe_chemotherapy_result_df %>%
  arrange(Protein, CellType)

```

```{r}

spe_chemotherapy_celltype_summary <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(NumProteins = n_distinct(Protein)) %>%
  arrange(desc(NumProteins))
spe_chemotherapy_celltype_summary
```

```{r}

# Create a new dataframe with cell types and their highly expressed proteins
spe_chemotherapy_celltype_proteins_df <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

spe_chemotherapy_celltype_proteins_df
```

```{r}
plot_data <- high_expression_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "Protein") %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "Expressed") %>%
  filter(Expressed == TRUE)

# Create the dotplot
dotplot <- ggplot(plot_data, aes(x = CellType, y = Protein)) +
  geom_point(size = 3, color = "blue") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8)
  ) +
  labs(
    title = "Highly Expressed Proteins Across Cell Types",
    x = "Cell Type",
    y = "Protein"
  )

# Display the plot
print(dotplot)
```

```{r}

```

#### For Chemo_Immu therapy

```{r}

## aggregate by cell type
celltype_mean1 <- aggregateAcrossCells(as(spe_chemo_immu, "SingleCellExperiment"),  
                     ids = spe_chemo_immu$ParentCell, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = features)



# Step 3: Ensure cell type metadata is correctly assigned
# Use unique 'spe$Label' values for cell type annotations
colData(celltype_mean1)$celltype <- factor(unique(spe_chemo_immu$ParentCell))

# Step 4: Count the number of cells per aggregated cell type
colData(celltype_mean1)$ncells <- as.numeric(table(spe_chemo_immu$ParentCell)[unique(spe_chemo_immu$ParentCell)])

# Step 5: Plot the heatmap using dittoHeatmap
dittoHeatmap(celltype_mean1,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scale = "row",                   # Scale by row (gene-wise)
             heatmap.colors = colorRampPalette(c("blue", "white", "red"))(100),    # Use viridis color scale
             annot.by = c("celltype", "ncells"),
             fontsize_row = 7,                 # Adjust the font size for type markers
                            # Show all color breaks for cell types
             show_legend = TRUE,
             legend.position = "right", 
             main = "Heatmap for celltypes and their expressed biomarkers for spe_chemo_Immu")  # Annotate by 'celltype' and 'ncells'
```

```{r}

# Extract expression data
expression_data1 <- assay(celltype_mean1, "exprs")
```

```{r}

# Function to identify highly expressed proteins
identify_high_expression1 <- function(expression_vector, threshold = 0.90) {
  quantile_threshold <- quantile(expression_vector, probs = threshold)
  return(expression_vector > quantile_threshold)
}


# Apply the function to each row (protein)
high_expression_matrix1 <- t(apply(expression_data1, 1, identify_high_expression1))

```

```{r}

# Convert to a data frame
high_expression_df1 <- as.data.frame(high_expression_matrix1)

# Add protein names as a column
high_expression_df1$Protein <- rownames(high_expression_df1)
```

```{r}

# Reshape the data frame
spe_chemo_immu_result_df <- high_expression_df1 %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "IsHighlyExpressed") %>%
  filter(IsHighlyExpressed) %>%
  select(-IsHighlyExpressed)


# Sort the dataframe
spe_chemo_immu_result_df <- spe_chemo_immu_result_df %>%
  arrange(Protein, CellType)

```

```{r}

spe_chemo_immu_celltype_summary <- spe_chemo_immu_result_df %>%
  group_by(CellType) %>%
  summarise(NumProteins = n_distinct(Protein)) %>%
  arrange(desc(NumProteins))
spe_chemo_immu_celltype_summary
```

```{r}

# Create a new dataframe with cell types and their highly expressed proteins
spe_chemo_immu_celltype_proteins_df <- spe_chemo_immu_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

spe_chemo_immu_celltype_proteins_df
```

```{r}

plot_data <- high_expression_matrix1 %>%
  as.data.frame() %>%
  rownames_to_column(var = "Protein") %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "Expressed") %>%
  filter(Expressed == TRUE)

# Create the dotplot
dotplot <- ggplot(plot_data, aes(x = CellType, y = Protein)) +
  geom_point(size = 3, color = "blue") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8)
  ) +
  labs(
    title = "Highly Expressed Proteins Across Cell Types",
    x = "Cell Type",
    y = "Protein"
  )

# Display the plot
print(dotplot)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```         
```

##### Save the transformed datasets

```{r}

saveRDS(spe_chemo, "saved_data/transformed_spe_chemo_Spatial_Graph.rds" )
```

```{r}

saveRDS(spe_chemo_immu, "saved_data/transformed_spe_chemo_immu_Spatial_Graph.rds" )
```

#### Extract spe_chemo data with Biopsyphase having 3 columns

```{r}

# Extract the colData from the SpatialExperiment object
col_data <- colData(spe_chemo)

# Check unique values in BiopsyPhase to confirm
unique_biopsy_phases <- unique(col_data$BiopsyPhase)
print(unique_biopsy_phases)

# Convert colData to a data.frame
col_data_df <- as.data.frame(col_data)

# Now you can use dplyr functions on col_data_df
patients_with_all_phases <- col_data_df %>%
  group_by(PatientID) %>%
  filter(all(c("On-treatment", "Baseline", "Post-treatment") %in% BiopsyPhase)) %>%
  summarize() %>%
  pull(PatientID)

# Filter the original data for these PatientIDs
filtered_indices <- which(col_data$PatientID %in% patients_with_all_phases)

# Extract the relevant data
if (length(filtered_indices) > 0) {
    filtered_spe_chemo <- spe_chemo[, filtered_indices]
} else {
    stop("No entries found for patients with all BiopsyPhase values.")
}

# Print out the structure of the filtered SpatialExperiment
print(filtered_spe_chemo)
```

```{r}
```

#### Extract spe_chemo_Immu data with Biopsyphase having 3 columns

```{r}

# Extract the colData from the SpatialExperiment object
col_data1 <- colData(spe_chemo_immu)

# Check unique values in BiopsyPhase to confirm
unique_biopsy_phases1 <- unique(col_data1$BiopsyPhase)
print(unique_biopsy_phases1)

# Convert colData to a data.frame
col_data_df1 <- as.data.frame(col_data1)

# Now you can use dplyr functions on col_data_df
patients_with_all_phases1 <- col_data_df1 %>%
  group_by(PatientID) %>%
  filter(all(c("On-treatment", "Baseline", "Post-treatment") %in% BiopsyPhase)) %>%
  summarize() %>%
  pull(PatientID)

# Filter the original data for these PatientIDs
filtered_indices1 <- which(col_data1$PatientID %in% patients_with_all_phases1)

# Extract the relevant data
if (length(filtered_indices) > 0) {
    filtered_spe_chemo_immu <- spe_chemo_immu[, filtered_indices1]
} else {
    stop("No entries found for patients with all BiopsyPhase values.")
}

# Print out the structure of the filtered SpatialExperiment
print(filtered_spe_chemo_immu)
```

```{r}
sort(unique(filtered_spe_chemo$PatientID))

```

```{r}

```

```{r}
sort(unique(filtered_spe_chemo_immu$PatientID))

```

```{r}

```

#### Find all the highest expressed biomarkers in each cell typeGetting the data for patient with ID NT009 to get parameters for the model using the chemotherapy only dataset.

```{r}
# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy_patient_1 <- spe_chemo[, colData(spe_chemo)$PatientID == "NT009"]

spe_chemotherapy_patient_1
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_1, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT009 Cell types on UMAP for spe_chemotherapy")
```

```{r}

# Function to count cells by type and biopsyphase
count_cells_by_type_and_phase <- function(spe) {
  # Extract cell type and biopsyphase information from colData
  cell_data <- as.data.frame(colData(spe))
  
  # Ensure 'ParentCell' and 'BiopsyPhase' columns exist
  if (!all(c("ParentCell", "BiopsyPhase") %in% colnames(cell_data))) {
    stop("Both 'ParentCell' and 'BiopsyPhase' columns must be present in colData")
  }
  
  # Count cells
  cell_counts <- cell_data %>%
    group_by(ParentCell, BiopsyPhase) %>%
    summarise(count = n(), .groups = "drop") %>%
    pivot_wider(names_from = BiopsyPhase, values_from = count, values_fill = 0)
  
  # Calculate total for each cell type
  cell_counts$Total <- rowSums(cell_counts[, -1])
  
  # Reorder columns to have Total as the second column
  cell_counts <- cell_counts %>%
    select(ParentCell, Total, everything())
  
  return(cell_counts)
}

```

```{r}

# Main analysis script
main_analysis <- function(spe) {
  # Perform cell count analysis
  result <- count_cells_by_type_and_phase(spe)
  
  # Print the result
  print(result)
  
  # Optionally, you can save the result to a CSV file
  # write.csv(result, "cell_count_analysis_result.csv", row.names = FALSE)
  
  return(result)
  
}
```

```{r}

result <- main_analysis(spe_chemotherapy_patient_1)
result <- result %>% filter(ParentCell != "Other")
result
```

```{r}

# Function to calculate column-wise statistics
calculate_column_stats <- function(data, column) {
  total_cells <- sum(data[[column]])
  ratios <- data[[column]] / total_cells
  variance <- var(data[[column]])
  
  tibble(
    ParentCell = data$ParentCell,
    Column = column,
    Count = data[[column]],
    Total_in_Column = total_cells,
    Ratio = round(ratios, 4),
    Proportion_Percent = round(ratios * 100, 2),
    Column_Variance = round(variance, 2)
  )
}

```

```{r}

# Calculate statistics for each column
columns_to_analyze <- c("Total", "Baseline", "On-treatment", "Post-treatment")
all_stats <- map_dfr(columns_to_analyze, ~calculate_column_stats(result, .x))

# Calculate cell type specific statistics
cell_type_stats <- result %>%
  pivot_longer(
    cols = -ParentCell,
    names_to = "Phase",
    values_to = "Count"
  ) %>%
  group_by(ParentCell) %>%
  summarise(
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = sd(Count)/mean(Count),
    Min = min(Count),
    Max = max(Count),
    Range = max(Count) - min(Count),
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Calculate phase-specific statistics
phase_stats <- all_stats %>%
  group_by(Column) %>%
  summarise(
    Total_Cells = sum(Count),
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = SD/Mean_Count,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Create detailed ratio analysis
ratio_analysis <- all_stats %>%
  select(ParentCell, Column, Ratio) %>%
  pivot_wider(
    names_from = Column,
    values_from = Ratio,
    names_prefix = "Ratio_"
  )

# Create ratio variance analysis
ratio_variance <- all_stats %>%
  group_by(ParentCell) %>%
  summarise(
    Ratio_Mean = mean(Ratio),
    Ratio_SD = sd(Ratio),
    Ratio_Variance = var(Ratio),
    Ratio_CV = Ratio_SD/Ratio_Mean,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 4)))

```

```{r}

print(all_stats)

```

```{r}
print(ratio_analysis)

```

```{r}

print(ratio_variance)
```

```{r}

# Function to calculate growth rate between two time points
calculate_growth_rate <- function(N0, N1, time_interval = 1) {
  # Using exponential growth formula: N1 = N0 * e^(r*t)
  # Solving for r: r = ln(N1/N0)/t
  r <- log(N1/N0) / time_interval
  return(r)
}

# Function to estimate carrying capacity using logistic growth model
estimate_carrying_capacity <- function(N0, N1, r, time_interval = 1) {
  # Using logistic growth formula: K = N1*N0*exp(r*t)/(N0*(exp(r*t)-1))
  # Where K is carrying capacity, r is growth rate, t is time interval
  exp_rt <- exp(r * time_interval)
  K <- (N1 * N0 * exp_rt) / (N0 * (exp_rt - 1))
  return(K)
}
```

```{r}

library(minpack.lm) 
```

```{r}

# Calculate carrying capacities and growth parameters
carrying_capacity_analysis <- ratio_analysis %>%
  mutate(
    # Calculate growth rates
    growth_rate = map2_dbl(
      Ratio_Baseline, `Ratio_On-treatment`,
      ~calculate_growth_rate(.x, .y)
    ),
    
    # Calculate basic carrying capacity estimate
    carrying_capacity_basic = map2_dbl(
      Ratio_Baseline, `Ratio_On-treatment`,
      ~max(.x, .y) * 1.5  # Simple initial estimate
    ),
    
    # Calculate refined carrying capacity
    carrying_capacity = pmap_dbl(
      list(Ratio_Baseline, `Ratio_On-treatment`, growth_rate),
      estimate_carrying_capacity
    ),
    
    # Calculate additional metrics
    density_dependent_factor = 1 - (Ratio_Baseline/carrying_capacity),
    population_pressure = Ratio_Baseline/carrying_capacity,
    growth_potential = carrying_capacity - Ratio_Baseline,
    
    # Calculate confidence metrics
    carrying_capacity_lower = carrying_capacity * 0.8,  # 80% CI
    carrying_capacity_upper = carrying_capacity * 1.2   # 80% CI
  ) %>%
  # Clean up extreme values
  mutate(
    carrying_capacity = ifelse(carrying_capacity < 0, NA, carrying_capacity),
    carrying_capacity = ifelse(carrying_capacity > 1e6, NA, carrying_capacity)
  ) %>%
  # Round numeric columns
  mutate(across(where(is.numeric), ~round(., 2)))
```

```{r}
print(carrying_capacity_analysis %>% 
      select(ParentCell, growth_rate, carrying_capacity, 
             carrying_capacity_lower, carrying_capacity_upper,
             population_pressure))

```

```         
```

#### 

```{r}

```

#### Getting the mean expression values of some of the protein biomarkers in the various cell types for spe_chemotherapy Patient 1.

```{r}

extract_mean_expression_tidy <- function(spe, 
                                         cell_type_column = "ParentCell", 
                                         assay_name = "transformed_counts") {
  # Check if the cell type column exists in colData
  if (!cell_type_column %in% colnames(colData(spe))) {
    stop(paste("Column", cell_type_column, "not found in colData"))
  }
  
  # Check if the specified assay exists
  if (!assay_name %in% assayNames(spe)) {
    stop(paste("Assay", assay_name, "not found in the SpatialExperiment object"))
  }
  
  # Extract expression data
  expression_data <- assay(spe, assay_name)
  
  # Extract cell type information
  cell_types <- colData(spe)$ParentCell
  
  # Calculate mean expression for each biomarker and cell type
  mean_expression <- t(apply(expression_data, 1, function(x) tapply(x, cell_types, median, na.rm = TRUE)))
  
  # Convert to a tidy data frame
  mean_expression_df <- as.data.frame(mean_expression) %>%
    rownames_to_column("Biomarker") %>%
    pivot_longer(cols = -Biomarker, names_to = "CellType", values_to = "MeanExpression")
  
  # Sort the data frame by CellType
  mean_expression_df_sorted <- mean_expression_df %>%
    arrange(CellType, Biomarker)
  
  return(mean_expression_df_sorted)
}
```

```{r}

results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_1, cell_type_column = "ParentCell",assay_name = "transformed_counts")

results2 
```

```{r}

biomarkers_of_interest <- c("CD31", "CD8", "CD163", "CD3", "PD-L1 (SP142)", "OX40","CD11c", "TOX", "PD-1", "IDO","ICOS","PD-L1 (73-10)","FOXP3","CD4", "PDGFRB", "PDPN", "Vimentin", "Caveolin-1", "HLA-DR", "SMA","Helios","AR","Ki67","CD45","CD68","Carboplatin","GZMB" )

```

```{r}

# Extract mean expression for selected biomarkers
result3 = results2[results2$Biomarker  %in% biomarkers_of_interest, ]
result3
```

```{r}

```

#### 

#### Find all the highest expressed biomarkers in each cell typeGetting the data for patient with ID NT010 to get parameters for the model using the chemotherapy only dataset.

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy_patient_2 <- spe_chemo[, colData(spe_chemo)$PatientID == "NT010"]

spe_chemotherapy_patient_2
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_2, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT010 Cell types on UMAP for spe_chemotherapy")
```

```{r}

Patient2_result <- main_analysis(spe_chemotherapy_patient_2)
Patient2_result <- Patient2_result %>% filter(ParentCell != "Other")
Patient2_result
```

```{r}

# Calculate statistics for each column
columns_to_analyze <- c("Total", "Baseline", "On-treatment", "Post-treatment")
all_stats <- map_dfr(columns_to_analyze, ~calculate_column_stats(Patient2_result, .x))

# Calculate cell type specific statistics
cell_type_stats <- Patient2_result %>%
  pivot_longer(
    cols = -ParentCell,
    names_to = "Phase",
    values_to = "Count"
  ) %>%
  group_by(ParentCell) %>%
  summarise(
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = sd(Count)/mean(Count),
    Min = min(Count),
    Max = max(Count),
    Range = max(Count) - min(Count),
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Calculate phase-specific statistics
phase_stats <- all_stats %>%
  group_by(Column) %>%
  summarise(
    Total_Cells = sum(Count),
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = SD/Mean_Count,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Create detailed ratio analysis
ratio_analysis <- all_stats %>%
  select(ParentCell, Column, Ratio) %>%
  pivot_wider(
    names_from = Column,
    values_from = Ratio,
    names_prefix = "Ratio_"
  )

# Create ratio variance analysis
ratio_variance <- all_stats %>%
  group_by(ParentCell) %>%
  summarise(
    Ratio_Mean = mean(Ratio),
    Ratio_SD = sd(Ratio),
    Ratio_Variance = var(Ratio),
    Ratio_CV = Ratio_SD/Ratio_Mean,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 4)))

```

```{r}

print(ratio_analysis)
```

```{r}

Patient2_results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_2, cell_type_column = "ParentCell",assay_name = "transformed_counts")

Patient2_results2
```

```{r}

# Extract mean expression for selected biomarkers
Patient2_result3 = Patient2_results2[Patient2_results2$Biomarker  %in% biomarkers_of_interest, ]
Patient2_result3
```

```{r}

```

#### Find all the highest expressed biomarkers in each cell typeGetting the data for patient with ID NT013 to get parameters for the model using the chemotherapy only dataset.

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy_patient_3 <- spe_chemo[, colData(spe_chemo)$PatientID == "NT013"]

spe_chemotherapy_patient_3
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_3, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT013 Cell types on UMAP for spe_chemotherapy")
```

```{r}

Patient3_result <- main_analysis(spe_chemotherapy_patient_3)
Patient3_result <- Patient3_result %>% filter(ParentCell != "Other")
Patient3_result
```

```{r}

# Calculate statistics for each column
columns_to_analyze <- c("Total", "Baseline", "On-treatment", "Post-treatment")
all_stats <- map_dfr(columns_to_analyze, ~calculate_column_stats(Patient3_result, .x))

# Calculate cell type specific statistics
cell_type_stats <- Patient3_result %>%
  pivot_longer(
    cols = -ParentCell,
    names_to = "Phase",
    values_to = "Count"
  ) %>%
  group_by(ParentCell) %>%
  summarise(
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = sd(Count)/mean(Count),
    Min = min(Count),
    Max = max(Count),
    Range = max(Count) - min(Count),
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Calculate phase-specific statistics
phase_stats <- all_stats %>%
  group_by(Column) %>%
  summarise(
    Total_Cells = sum(Count),
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = SD/Mean_Count,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Create detailed ratio analysis
ratio_analysis <- all_stats %>%
  select(ParentCell, Column, Ratio) %>%
  pivot_wider(
    names_from = Column,
    values_from = Ratio,
    names_prefix = "Ratio_"
  )

# Create ratio variance analysis
ratio_variance <- all_stats %>%
  group_by(ParentCell) %>%
  summarise(
    Ratio_Mean = mean(Ratio),
    Ratio_SD = sd(Ratio),
    Ratio_Variance = var(Ratio),
    Ratio_CV = Ratio_SD/Ratio_Mean,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 4)))

```

```{r}

print(ratio_analysis)
```

```{r}

Patient3_results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_3, cell_type_column = "ParentCell",assay_name = "transformed_counts")

Patient3_results2
```

```{r}

# Extract mean expression for selected biomarkers
Patient3_result3 = Patient3_results2[Patient3_results2$Biomarker  %in% biomarkers_of_interest, ]
Patient3_result3
```

```{r}

```

#### Find all the highest expressed biomarkers in each cell type. Getting the data for patient with ID NT022 to get parameters for the model using the chemotherapy only dataset.

```{r}



spe_chemotherapy_patient_4 <- spe_chemo[, colData(spe_chemo)$PatientID == "NT022"]

spe_chemotherapy_patient_4
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_4, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT022 Cell types on UMAP for spe_chemotherapy")
```

```{r}

Patient4_result <- main_analysis(spe_chemotherapy_patient_4)
Patient4_result <- Patient4_result %>% filter(ParentCell != "Other")
Patient4_result
```

```{r}

# Calculate statistics for each column
columns_to_analyze <- c("Total", "Baseline", "On-treatment", "Post-treatment")
all_stats <- map_dfr(columns_to_analyze, ~calculate_column_stats(Patient4_result, .x))

# Calculate cell type specific statistics
cell_type_stats <- Patient4_result %>%
  pivot_longer(
    cols = -ParentCell,
    names_to = "Phase",
    values_to = "Count"
  ) %>%
  group_by(ParentCell) %>%
  summarise(
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = sd(Count)/mean(Count),
    Min = min(Count),
    Max = max(Count),
    Range = max(Count) - min(Count),
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Calculate phase-specific statistics
phase_stats <- all_stats %>%
  group_by(Column) %>%
  summarise(
    Total_Cells = sum(Count),
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = SD/Mean_Count,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Create detailed ratio analysis
ratio_analysis <- all_stats %>%
  select(ParentCell, Column, Ratio) %>%
  pivot_wider(
    names_from = Column,
    values_from = Ratio,
    names_prefix = "Ratio_"
  )

# Create ratio variance analysis
ratio_variance <- all_stats %>%
  group_by(ParentCell) %>%
  summarise(
    Ratio_Mean = mean(Ratio),
    Ratio_SD = sd(Ratio),
    Ratio_Variance = var(Ratio),
    Ratio_CV = Ratio_SD/Ratio_Mean,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 4)))

```

```{r}

print(ratio_analysis)

```

```{r}

Patient4_results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_4, cell_type_column = "ParentCell",assay_name = "transformed_counts")

Patient4_results2
```

```{r}

# Extract mean expression for selected biomarkers
Patient4_result3 = Patient4_results2[Patient4_results2$Biomarker  %in% biomarkers_of_interest, ]
Patient4_result3
```

```{r}

```

#### Find all the highest expressed biomarkers in each cell type. Getting the data for patient with ID NT023 to get parameters for the model using the chemotherapy only dataset.

```{r}

spe_chemotherapy_patient_5 <- spe_chemo[, colData(spe_chemo)$PatientID == "NT023"]

spe_chemotherapy_patient_5
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_5, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT023 Cell types on UMAP for spe_chemotherapy")
```

```{r}

Patient5_result <- main_analysis(spe_chemotherapy_patient_5)
Patient5_result <- Patient5_result %>% filter(ParentCell != "Other")
Patient5_result
```

```{r}

# Calculate statistics for each column
columns_to_analyze <- c("Total", "Baseline", "On-treatment", "Post-treatment")
all_stats <- map_dfr(columns_to_analyze, ~calculate_column_stats(Patient5_result, .x))

# Calculate cell type specific statistics
cell_type_stats <- Patient5_result %>%
  pivot_longer(
    cols = -ParentCell,
    names_to = "Phase",
    values_to = "Count"
  ) %>%
  group_by(ParentCell) %>%
  summarise(
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = sd(Count)/mean(Count),
    Min = min(Count),
    Max = max(Count),
    Range = max(Count) - min(Count),
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Calculate phase-specific statistics
phase_stats <- all_stats %>%
  group_by(Column) %>%
  summarise(
    Total_Cells = sum(Count),
    Mean_Count = mean(Count),
    SD = sd(Count),
    Variance = var(Count),
    CV = SD/Mean_Count,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Create detailed ratio analysis
ratio_analysis <- all_stats %>%
  select(ParentCell, Column, Ratio) %>%
  pivot_wider(
    names_from = Column,
    values_from = Ratio,
    names_prefix = "Ratio_"
  )

# Create ratio variance analysis
ratio_variance <- all_stats %>%
  group_by(ParentCell) %>%
  summarise(
    Ratio_Mean = mean(Ratio),
    Ratio_SD = sd(Ratio),
    Ratio_Variance = var(Ratio),
    Ratio_CV = Ratio_SD/Ratio_Mean,
    .groups = 'drop'
  ) %>%
  mutate(across(where(is.numeric), ~round(., 4)))

```

```{r}

print(ratio_analysis)
```

```{r}

Patient5_results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_5, cell_type_column = "ParentCell",assay_name = "transformed_counts")

Patient5_results2
```

```{r}

# Extract mean expression for selected biomarkers
Patient5_result3 = Patient5_results2[Patient5_results2$Biomarker  %in% biomarkers_of_interest, ]
Patient5_result3
```

```{r}

```

```{r}

```

```{r}

```

```{r}


```

```{r}

```

```{r}

```

#### Chemo & Immuno analysis

#### Find all the highest expressed biomarkers in each cell typeGetting the data for patient with ID NT001 to get parameters for the model using the chemotherapy only dataset.

```         
```

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemo_immu_patient_1 <- spe_chemo_immu[, colData(spe_chemo_immu)$PatientID == "NT001"]

spe_chemo_immu_patient_1
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemo_immu_patient_1, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT001 Cell types on UMAP for spe_chemo_immu_patient_1")
```

```{r}
## Count number of cells in each timepoint

result4 <- main_analysis(spe_chemo_immu_patient_1)
result4
```

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemo_immu_patient_2 <- spe_chemo_immu[, colData(spe_chemo_immu)$PatientID == "NT002"]

spe_chemo_immu_patient_2

result5 <- main_analysis(spe_chemo_immu_patient_2)
result5
```

#### Getting the mean expression values of some of the protein biomarkers in the various cell types for spe_chemotherapy.

```{r}

## aggregate by cell type
celltype_mean <- aggregateAcrossCells(as(spe_chemotherapy, "SingleCellExperiment"),  
                     ids = spe_chemotherapy$ParentCell, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = type_markers)



# Step 3: Ensure cell type metadata is correctly assigned
# Use unique 'spe$Label' values for cell type annotations
colData(celltype_mean)$celltype <- factor(unique(spe_chemotherapy$ParentCell))

# Step 4: Count the number of cells per aggregated cell type
colData(celltype_mean)$ncells <- as.numeric(table(spe_chemotherapy$ParentCell)[unique(spe_chemotherapy$ParentCell)])

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
