---
title: "R Notebook"
output: html_notebook
---

```{r}

setwd('V:/dkmarri/Breast_Cancer_IMC_Data/R_Codes')
```

```{r}

library(readr)
library(dplyr)
library(ggplot2)
library(cluster)
library(factoextra)
library(caret)
library(pROC)
library(glmnet)
library(reshape2)
library(tidyr)
library(readxl)
```

```{r}

#set.seed(23)
```

```{r}

# Load the data (update the file path accordingly)
Chemo_data <- read_excel("saved_data/Chemo_model_for_box_plot.xlsx")

Chemo_data
```

#### Growth Rate

```{r}

# Reshape the data from wide to long format for the selected variables
Chemo_data_growth <- Chemo_data %>%
  select(Patient, Gr_CC, Gr_Treg, Gr_Teff, Gr_M, Gr_EC, Gr_F, Survival)


Chemo_data_growth
```

```{r}

library(tidyverse)
library(ggpubr)
library(rstatix)
library(gridExtra)
library(ggsignif)
```

```{r}

```

```{r}


# Function to create enhanced distribution analysis
analyze_growth_rates <- function(data) {
  # Remove Patient column and gather growth rate variables
  long_data <- data %>%
    select(-Patient) %>%
    pivot_longer(
      cols = starts_with("Gr_"),
      names_to = "Cell_Type",
      values_to = "Growth_Rate"
    ) %>%
    mutate(Cell_Type = str_remove(Cell_Type, "Gr_"))
  
  # Create individual density plots for each cell type and survival status
  density_plots <- list()
  cell_types <- unique(long_data$Cell_Type)
  survival_types <- unique(long_data$Survival)
  
  for(cell in cell_types) {
    for(surv in survival_types) {
      plot_data <- long_data %>% 
        filter(Cell_Type == cell, Survival == surv)
      
      plot_title <- paste("Growth Rate Distribution -", cell, "-", surv)
      plot_color <- if(surv == "pCR") "#2ECC71" else "#E74C3C"
      
      p <- ggplot(plot_data, aes(x = Growth_Rate)) +
        geom_density(fill = plot_color, alpha = 0.7) +
        theme_minimal() +
        labs(
          title = plot_title,
          x = "Growth Rate",
          y = "Density"
        ) +
        theme(
          plot.title = element_text(face = "bold", size = 12),
          axis.title = element_text(size = 10, face = "bold"),
          axis.text = element_text(
            size = 9,
            face = "bold",
            color = "black"
          ),
          axis.text.x = element_text(
            face = "bold",
            color = "black"
          ),
          axis.text.y = element_text(
            face = "bold",
            color = "black"
          ),
          plot.margin = margin(10, 10, 10, 10)
        ) +
        scale_x_continuous(
          limits = c(0, 10),
          breaks = seq(0, 10, 2)
        ) +
        scale_y_continuous(
          expand = expansion(mult = c(0, 0.05)),
          breaks = function(x) pretty(x, n = 5)
        )
      
      density_plots[[paste(cell, surv, sep = "_")]] <- p
    }
  }
  
  # Calculate summary statistics
  summary_stats <- long_data %>%
    group_by(Cell_Type, Survival) %>%
    summarise(
      n = n(),
      mean = mean(Growth_Rate),
      sd = sd(Growth_Rate),
      median = median(Growth_Rate),
      q1 = quantile(Growth_Rate, 0.25),
      q3 = quantile(Growth_Rate, 0.75),
      .groups = "drop"
    )
  
  # Perform statistical tests
  stat_tests <- long_data %>%
    group_by(Cell_Type) %>%
    wilcox_test(Growth_Rate ~ Survival) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()
  
  # Return results as a list
  return(list(
    density_plots = density_plots,
    statistics = stat_tests,
    summary = summary_stats
  ))
}
```

```{r}

results <- analyze_growth_rates(Chemo_data_growth)
```

```{r}
 # Display all plots in a grid (4x3 arrangement for 12 plots)
plot_grid <- do.call(grid.arrange, 
                     c(results$density_plots, 
                       ncol = 3, 
                       nrow = 4))
```

```{r}

```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "CC_pCR_density_plot.png",    # File name
  plot = results$density_plots[["CC_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)
results$density_plots[["CC_pCR"]]
```

```{r}
# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Treg_pCR_density_plot.png",    # File name
  plot = results$density_plots[["Treg_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)
results$density_plots[["Treg_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Teff_pCR_density_plot.png",    # File name
  plot = results$density_plots[["Teff_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


# Or display individual plots
results$density_plots[["Teff_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "M_pCR_density_plot.png",    # File name
  plot = results$density_plots[["M_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["M_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "EC_pCR_density_plot.png",    # File name
  plot = results$density_plots[["EC_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["EC_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "F_pCR_density_plot.png",    # File name
  plot = results$density_plots[["F_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["F_pCR"]]  # View specific plot
```

```{r}

```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "CC_RD_density_plot.png",    # File name
  plot = results$density_plots[["CC_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)
# Or display individual plots
results$density_plots[["CC_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Treg_RD_density_plot.png",    # File name
  plot = results$density_plots[["Treg_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["Treg_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Teff_RD_density_plot.png",    # File name
  plot = results$density_plots[["Teff_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["Teff_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "M_RD_density_plot.png",    # File name
  plot = results$density_plots[["M_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["M_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "EC_RD_density_plot.png",    # File name
  plot = results$density_plots[["EC_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["EC_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "F_RD_density_plot.png",    # File name
  plot = results$density_plots[["F_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results$density_plots[["F_RD"]]  # View specific plot
```

```{r}
#  View summary statistics
 print(results$summary)
```

```{r}

```

```{r}

# Reshape data from wide to long format
long_data <- Chemo_data_growth %>%
  pivot_longer(
    cols = starts_with("Gr_"),
    names_to = "Cell_Type",
    values_to = "Growth_Rate"
  ) %>%
  mutate(Cell_Type = gsub("Gr_", "", Cell_Type))

long_data
```

```{r}


# Basic high-resolution PNG
png(filename = "Growth_rate_box.png",
    width = 3000,
    height = 2000,
    res = 300,
    units = "px")


# Create the box plot
ggplot(long_data, aes(x = Cell_Type, y = Growth_Rate, fill = Survival)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Growth Rate Comparison by Cell Type and Survival Status",
    x = "Cell Type",
    y = "Growth Rate",
    fill = "Survival Status"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    
  ) +
  scale_fill_manual(values = c("pCR" = "#3498db", "RD" = "#e74c3c"))+
  theme(
    text = element_text(size = 16, face = "bold"),        # Set font size and make it bold
    axis.title = element_text(size = 16, face = "bold"), # Bold axis titles
    axis.text = element_text(size = 16),                # Increase font size for axis text
    legend.text = element_text(size = 16),              # Increase font size for legend
    plot.title = element_text(size = 16, face = "bold") # Bold plot title
  )

dev.off()
```

```{r}

```

```{r}

stat.test <- long_data %>%
  group_by(Cell_Type) %>%
  wilcox_test(Growth_Rate ~ Survival) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")

stat.test
```

```{r}
 # Basic high-resolution PNG
 png(filename = "Growth_rate_box_with_stats.png",
     width = 3000,
     height = 2000,
     res = 300,
     units = "px")

# Create a box plot
bxp <- ggboxplot(
  long_data, x = "Cell_Type", y = "Growth_Rate", 
  color = "Survival", palette = c("pCR" = "#3498db", "RD" = "#e74c3c")
  )+
   theme(
    text = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 16, face = "bold")
  )


# Add p-values onto the box plots
stat.test <- stat.test %>%
  add_xy_position(x = "Cell_Type", dodge = 0.8)
bxp + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0, label.size = 12,   
    fontface = "bold" 
  )
# Add 10% spaces between the p-value labels and the plot border
bxp + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0, label.size = 6,   
    fontface = "bold" 
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))


dev.off()
```

```{r}
# Use adjusted p-values as labels
# Remove brackets
bxp + stat_pvalue_manual(
  stat.test,  label = "p.adj", tip.length = 0,
  remove.bracket = TRUE
  )

# Show adjusted p-values and significance levels
# Hide ns (non-significant)
bxp + stat_pvalue_manual(
  stat.test,  label = "{p.adj}{p.adj.signif}", 
  tip.length = 0, hide.ns = TRUE
  )
```

```{r}
# Additional statistical test
stat.test2 <- long_data %>%
  wilcox_test(Growth_Rate ~ Cell_Type, p.adjust.method = "bonferroni")
stat.test2
```

```{r}



stat.test <- stat.test %>%
  add_xy_position(x = "Cell_Type", dodge = 0.8)
bxp.complex <- bxp + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0,label.size = 4,   
    fontface = "bold" 
  )
# 2. Add stat.test2





# Add more space between brackets using `step.increase` 
stat.test2 <- stat.test2 %>% add_xy_position(x = "Cell_Type")
bxp.complex <- bxp.complex + 
  stat_pvalue_manual(
    stat.test2,  label = "p", tip.length = 0.02,
    step.increase = 0.05,label.size = 4,   
    fontface = "bold" 
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))+
  theme(
    text = element_text(size = 16, face = "bold"),        # Set font size and make it bold
    axis.title = element_text(size = 16, face = "bold"), # Bold axis titles
    axis.text = element_text(size = 16),                # Increase font size for axis text
    legend.text = element_text(size = 16),              # Increase font size for legend
    plot.title = element_text(size = 16, face = "bold") # Bold plot title
  )

# 3. Display the plot
bxp.complex 


```

```{r}

# Create the box plot
ggplot(long_data, aes(x = Patient, y = Growth_Rate, fill = Survival)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Growth Rate Comparison by Cell Type and Survival Status",
    x = "Cell Type",
    y = "Growth Rate",
    fill = "Survival Status"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_fill_manual(values = c("pCR" = "#3498db", "RD" = "#e74c3c"))
```

```{r}

```

#### Inhibition Rate

```{r}

# Reshape the data from wide to long format for the selected variables
Chemo_data_inhibit <- Chemo_data %>%
  select(Patient, In_CC, In_Treg, In_Teff, In_M, In_EC, In_F, Survival)


Chemo_data_inhibit
```

```{r}


# Function to create enhanced distribution analysis
analyze_inhibition_rates <- function(data) {
  # Remove Patient column and gather growth rate variables
  long_data <- data %>%
    select(-Patient) %>%
    pivot_longer(
      cols = starts_with("In_"),
      names_to = "Cell_Type",
      values_to = "Inhibition_Rate"
    ) %>%
    mutate(Cell_Type = str_remove(Cell_Type, "In_"))
  
  # Create individual density plots for each cell type and survival status
  density_plots <- list()
  cell_types <- unique(long_data$Cell_Type)
  survival_types <- unique(long_data$Survival)
  
  for(cell in cell_types) {
    for(surv in survival_types) {
      plot_data <- long_data %>% 
        filter(Cell_Type == cell, Survival == surv)
      
      plot_title <- paste("Inhibition Rate Distribution -", cell, "-", surv)
      plot_color <- if(surv == "pCR") "#2ECC71" else "#E74C3C"
      
      p <- ggplot(plot_data, aes(x = Inhibition_Rate)) +
        geom_density(fill = plot_color, alpha = 0.7) +
        theme_minimal() +
        labs(
          title = plot_title,
          x = "Inhibition Rate",
          y = "Density"
        ) +
        theme(
          plot.title = element_text(face = "bold", size = 12),
          axis.title = element_text(size = 10, face = "bold"),
          axis.text = element_text(
            size = 9,
            face = "bold",
            color = "black"
          ),
          axis.text.x = element_text(
            face = "bold",
            color = "black"
          ),
          axis.text.y = element_text(
            face = "bold",
            color = "black"
          ),
          plot.margin = margin(10, 10, 10, 10)
        ) +
        scale_x_continuous(
          limits = c(0, 10),
          breaks = seq(0, 10, 2)
        ) +
        scale_y_continuous(
          expand = expansion(mult = c(0, 0.05)),
          breaks = function(x) pretty(x, n = 5)
        )
      
      density_plots[[paste(cell, surv, sep = "_")]] <- p
    }
  }
  
  # Calculate summary statistics
  summary_stats <- long_data %>%
    group_by(Cell_Type, Survival) %>%
    summarise(
      n = n(),
      mean = mean(Inhibition_Rate),
      sd = sd(Inhibition_Rate),
      median = median(Inhibition_Rate),
      q1 = quantile(Inhibition_Rate, 0.25),
      q3 = quantile(Inhibition_Rate, 0.75),
      .groups = "drop"
    )
  
  # Perform statistical tests
  stat_tests <- long_data %>%
    group_by(Cell_Type) %>%
    wilcox_test(Inhibition_Rate ~ Survival) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()
  
  # Return results as a list
  return(list(
    density_plots = density_plots,
    statistics = stat_tests,
    summary = summary_stats
  ))
}
```

```{r}

results1 <- analyze_inhibition_rates(Chemo_data_inhibit)
```

```{r}
 # Display all plots in a grid (4x3 arrangement for 12 plots)
plot_grid <- do.call(grid.arrange, 
                     c(results1$density_plots, 
                       ncol = 3, 
                       nrow = 4))
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "CC_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["CC_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

# Or display individual plots
results1$density_plots[["CC_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "CC_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["CC_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["CC_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Treg_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["Treg_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["Treg_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Treg_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["Treg_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["Treg_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Teff_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["Teff_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["Teff_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "Teff_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["Teff_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["Teff_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "M_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["M_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["M_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "M_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["M_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["M_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "EC_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["EC_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["EC_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "EC_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["EC_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)


results1$density_plots[["EC_RD"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "F_pCR_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["F_pCR"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)



results1$density_plots[["F_pCR"]]  # View specific plot
```

```{r}

# Save the specific density plot as a high-resolution PNG
ggsave(
  filename = "F_RD_inhibition_density_plot.png",    # File name
  plot = results1$density_plots[["F_RD"]],  # Specify the plot
  width = 4,    # Width in inches (adjustable)
  height = 3,   # Height in inches (adjustable)
  dpi = 1200,    # High resolution
  units = "in"  # Ensure correct scaling
)

results1$density_plots[["F_RD"]]  # View specific plot
```

```{r}

```

```{r}

```

```{r}

# Reshape data from wide to long format
long_data2 <- Chemo_data_inhibit %>%
  pivot_longer(
    cols = starts_with("In_"),
    names_to = "Cell_Type",
    values_to = "Inhibition_Rate"
  ) %>%
  mutate(Cell_Type = gsub("In_", "", Cell_Type))

long_data2
```

```{r}

# Create the box plot

# Basic high-resolution PNG
png(filename = "Inhibition_rate_box.png",
    width = 3000,
    height = 2000,
    res = 300,
    units = "px")




ggplot(long_data2, aes(x = Cell_Type, y = Inhibition_Rate, fill = Survival)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Inhibition Rate Comparison by Cell Type and Survival Status",
    x = "Cell Type",
    y = "Inhibition Rate",
    fill = "Survival Status"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_fill_manual(values = c("pCR" = "#3498db", "RD" = "#e74c3c"))+
  theme(
    text = element_text(size = 16, face = "bold"),        # Set font size and make it bold
    axis.title = element_text(size = 16, face = "bold"), # Bold axis titles
    axis.text = element_text(size = 16),                # Increase font size for axis text
    legend.text = element_text(size = 16),              # Increase font size for legend
    plot.title = element_text(size = 16, face = "bold") # Bold plot title
  )

dev.off()
```

```{r}

stat.test3 <- long_data2 %>%
  group_by(Cell_Type) %>%
  wilcox_test(Inhibition_Rate ~ Survival) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")

stat.test3
```

```{r}
# Basic high-resolution PNG
png(filename = "Inhibition_rate_box_with_stats.png",
    width = 3000,
    height = 2000,
    res = 300,
    units = "px")

# Create a box plot
bxp2 <- ggboxplot(
  long_data2, x = "Cell_Type", y = "Inhibition_Rate", 
  color = "Survival", palette = c("pCR" = "#3498db", "RD" = "#e74c3c")
  )

# Add p-values onto the box plots
stat.test3 <- stat.test3 %>%
  add_xy_position(x = "Cell_Type", dodge = 0.8)
bxp2 + stat_pvalue_manual(
  stat.test3,  label = "p", tip.length = 0
  )

# Add 10% spaces between the p-value labels and the plot border
bxp2 + stat_pvalue_manual(
  stat.test3,  label = "p", tip.length = 0, label.size = 6,   
    fontface = "bold" 
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))+
  theme(
    text = element_text(size = 16, face = "bold"),        # Set font size and make it bold
    axis.title = element_text(size = 16, face = "bold"), # Bold axis titles
    axis.text = element_text(size = 16),                # Increase font size for axis text
    legend.text = element_text(size = 16),              # Increase font size for legend
    plot.title = element_text(size = 16, face = "bold") # Bold plot title
  )

dev.off()
```

```{r}


```

```{r}
# Additional statistical test
stat.test4 <- long_data2 %>%
  wilcox_test(Inhibition_Rate ~ Cell_Type, p.adjust.method = "bonferroni")
stat.test4
```

```{r}
stat.test3 <- stat.test3 %>%
  add_xy_position(x = "Cell_Type", dodge = 0.8)
bxp2.complex <- bxp2 + stat_pvalue_manual(
  stat.test3,  label = "p", tip.length = 0,label.size = 4,   
    fontface = "bold" 
  )
# 2. Add stat.test2





# Add more space between brackets using `step.increase` 
stat.test4 <- stat.test4 %>% add_xy_position(x = "Cell_Type")
bxp2.complex <- bxp2.complex + 
  stat_pvalue_manual(
    stat.test4,  label = "p", tip.length = 0.02,
    step.increase = 0.05,label.size = 4,   
    fontface = "bold" 
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))+
  theme(
    text = element_text(size = 16, face = "bold"),        # Set font size and make it bold
    axis.title = element_text(size = 16, face = "bold"), # Bold axis titles
    axis.text = element_text(size = 16),                # Increase font size for axis text
    legend.text = element_text(size = 16),              # Increase font size for legend
    plot.title = element_text(size = 16, face = "bold") # Bold plot title
  )

# 3. Display the plot
bxp2.complex 


```

```{r}

# Create the box plot
ggplot(long_data2, aes(x = Patient, y = Inhibition_Rate, fill = Survival)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Growth Rate Comparison by Cell Type and Survival Status",
    x = "Patient",
    y = "Inhibition Rate",
    fill = "Survival Status"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_fill_manual(values = c("pCR" = "#3498db", "RD" = "#e74c3c"))
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### Classification

```{r}

# Load the data (update the file path accordingly)
Chemo_data <- read_excel("saved_data/Chemo_model_for_box_plot.xlsx")

Chemo_data
```

```{r}
Chemo_data <-  Chemo_data[, !names(Chemo_data ) %in% c('Gr_F
', 'In_EC', 'CC_Slope', 'Treg_Slope', 'Treg_ACF', 'Teff_Slope', 'Teff_AUC', 'M_Slope', 'EC_AUC', 'F_Slope')]
```

```{r}

# Prepare the feature matrix for clustering
features <- colnames(Chemo_data)[-c(1, ncol(Chemo_data))]
```

```{r}

 # Normalize the features
Chemo_data_fea <- Chemo_data[, features]
Chemo_data_fea

```

```{r}

```

#### Logistic Regression for Classification

```{r}

# Prepare the data for logistic regression
X <- Chemo_data_fea
y <- ifelse(Chemo_data$Survival == "pCR", 1, 0)

```

```{r}

loocv_control <- trainControl(
   method = "LOOCV",
  summaryFunction = twoClassSummary,
  classProbs = TRUE)
```

```{r}

# Prepare data for caret
train_data <- data.frame(X, Survival = factor(Chemo_data$Survival, levels = c("RD", "pCR")))

```

```{r}
# Extract feature names from training data (excluding the target variable)
features <- colnames(train_data)[colnames(train_data) != "Survival"]
```

```{r}

# Compute class weights
total_samples <- nrow(Chemo_data)
class_counts <- table(Chemo_data$Survival)
class_weights <- total_samples / class_counts

# Assign weights based on class
Chemo_data$weights <- sapply(Chemo_data$Survival, function(x) class_weights[x])

```

```{r}


loocv_model <- train(
  Survival ~ ., 
  data = train_data,
  method = "glmnet",
  family = "binomial",
  weights = Chemo_data$weights,
  trControl = loocv_control,
  metric = "ROC"
)
```

```{r}
# Print model summary
print(loocv_model)

```

```{r}

# Observed survival status
observed <- loocv_model$pred$obs

# Predicted probabilities for class "pCR"
probabilities <- loocv_model$pred$pCR
```

```{r}
# Compute ROC curve
roc_curve <- roc(response = observed, predictor = probabilities)
roc_curve
```

```{r}

# Set up a high-resolution PNG device
png(filename = "ROC_Curve_pCR_Prediction.png",
    width = 5500,    # Width in pixels
    height = 4000,   # Height in pixels
    res = 900)  


# Set custom parameters for larger, bold text
par(cex.main = 1,      # Title size
    cex.lab = 1.5,     # Axis labels size
    cex.axis = 1.5,    # Axis numbers size
    font.main = 2,     # Bold title (2 means bold)
    font.lab = 1.5,      # Bold axis labels
    font.axis = 1.5)     # Bold axis numbers

# Plot ROC curve
plot(roc_curve, 
     main = "ROC Curve for pCR Prediction",
     auc.main = TRUE,
     color = TRUE,
     print.auc = TRUE)


dev.off()


# Plot ROC curve
plot(roc_curve, 
     main = "ROC Curve for pCR Prediction",
     auc.main = TRUE,
     color = TRUE,
     print.auc = TRUE)
```

```{r}
# Confusion Matrix
conf_matrix <- confusionMatrix(
  loocv_model$pred$pred, 
  loocv_model$pred$obs
)
print(conf_matrix)
```

```{r}

```

```{r}

# Extract feature names
features <- loocv_model$finalModel$xNames

# Extract coefficients at the optimal lambda
coef_matrix <- coef(loocv_model$finalModel, s = loocv_model$finalModel$lambdaOpt)
coef_values <- as.vector(coef_matrix[-1, 1])  # Exclude intercept

# Create feature importance data frame with normalized importance
feature_importance <- data.frame(
  Feature = features,
  Importance = abs(coef_values)
)

# Normalize importance to sum to 1
feature_importance$Importance <- feature_importance$Importance / sum(feature_importance$Importance)

# Sort by importance in descending order
feature_importance <- feature_importance[order(-feature_importance$Importance), ]

# Verify sum of importance is 1
print(sum(feature_importance$Importance))

# View result
print(feature_importance)
```

```{r}


# Set up a high-resolution PNG device
png(filename = "Logistic_feature_importance.png",
    width = 3000,    # Width in pixels
    height = 5000,   # Height in pixels
    res = 900)

# Optional: Bar plot of feature importance
library(ggplot2)

# Subset to top 15 features based on Importance
top_features <- feature_importance %>%
  arrange(desc(Importance)) %>%
  head(10)

# Create the bar plot for the top 15 features
feature_importance_plot <- ggplot(top_features, 
                                  aes(x = reorder(Feature, Importance), 
                                      y = Importance)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "",
       x = "Features",
       y = "Absolute Coefficient Value") +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold")
  )

# Print the plot
print(feature_importance_plot)

# Close the PNG device
dev.off()

```

```{r}

```

#### Using PRC

```{r}

library(PRROC) # For Precision-Recall Curve AUC calculations

# Custom summary function to compute PR AUC
prcSummary <- function(data, lev = NULL, model = NULL) {
  require(PRROC)
  
  # Check if predictions and probabilities are valid
  if (is.null(data$obs) || is.null(data[, lev[2]])) return(c(PRC = NA))
  
  # Compute precision-recall curve
  precision_recall <- pr.curve(scores.class0 = data[, lev[2]], weights.class0 = as.numeric(data$obs == lev[2]))
  
  # Return PRC (Area under Precision-Recall Curve)
  c(PRC = precision_recall$auc.integral)
}

```

```{r}

# Update trainControl to use the new metric
loocv_control2 <- trainControl(
  method = "LOOCV",
  summaryFunction = prcSummary,
  classProbs = TRUE
)
```

```{r}
set.seed(12)
loocv_model2 <- train(
  Survival ~ ., 
  data = train_data,
  method = "glmnet",
  family = "binomial",
  weights = Chemo_data$weights,
  trControl = loocv_control2,
  metric = "PRC"
)
```

```{r}

# # Train the logistic regression model using PRC metric
# set.seed(22)
# loocv_model2 <- train(
#   Survival ~ ., 
#   data = train_data,
#   method = "glmnet",
#   family = "binomial",
#   trControl = loocv_control2,
#   metric = "PRC" # Use PRC as the performance metric
# )
# 

```

```{r}

# Print the model summary
print(loocv_model2)

```

```{r}

# Observed survival status and predicted probabilities for "pCR"
observed2 <- ifelse(loocv_model2$pred$obs == "pCR", 1, 0)
probabilities2 <- loocv_model2$pred$pCR
```

```{r}

# Compute the Precision-Recall Curve
prc <- pr.curve(scores.class0 = probabilities2, weights.class0 = observed2, curve = TRUE)
prc
```

```{r}

# Set up a high-resolution PNG device
png(filename = "PRC_Curve_pCR_Prediction.png",
    width = 3000,    # Width in pixels
    height = 2000,   # Height in pixels
    res = 900)  


# Set custom parameters for larger, bold text
par(cex.main = 1,      # Title size
    cex.lab = 1.5,     # Axis labels size
    cex.axis = 1.5,    # Axis numbers size
    font.main = 2,     # Bold title (2 means bold)
    font.lab = 2,      # Bold axis labels
    font.axis = 2)     # Bold axis numbers


# Plot the Precision-Recall Curve
plot(prc,
     main = "Precision-Recall Curve for pCR Prediction",
     auc.main = TRUE,
     color = TRUE,
     maxmin = FALSE)

dev.off


```

```{r}

plot(prc,
     main = "Precision-Recall Curve for pCR Prediction",
     auc.main = TRUE,
     color = TRUE,
     maxmin = FALSE)
```

```{r}

```

```{r}

```

```{r}



```

```{r}
# Load the data (update the file path accordingly)
Chemo_data <- read_excel("saved_data/Chemo_model_for_box_plot.xlsx")

Chemo_data
```

```{r}

# Add these diagnostic checks after loading your data
print(table(Chemo_data$Survival))  # Check class distribution
print(sum(is.na(Chemo_data$Survival)))  # Check for NA values
print(dim(Chemo_data))  # Check dimensions
```

#### Using XGBOOST

```{r}

library(xgboost)
library(caret)
library(tidyverse)
library(pROC)
library(parallel)
library(doParallel)
library(foreach)
library(ROSE)
library(dplyr)
```

```{r}
set.seed(5)
```

```{r}


# Data preprocessing function with validation
preprocess_data <- function(data) {
  # Check if data is empty
  if (nrow(data) == 0) {
    stop("Input data is empty")
  }
  
  # Remove Patient ID column
  if ("Patient" %in% names(data)) {
    data$Patient <- NULL
  }
  
  
   # Convert categorical variables to numeric
  categorical_cols <- sapply(data, is.character)
  for (col in names(data)[categorical_cols]) {
    if (col != "Survival") {  # Skip the target variable
      data[[col]] <- as.numeric(as.factor(data[[col]]))
    }
  }
  
  # Convert target to numeric (0/1)
  data$Survival <- as.numeric(data$Survival == "pCR")
  
  # Check class distribution
  class_dist <- table(data$Survival)
  if (length(class_dist) < 2) {
    stop("Dataset contains only one class. Both positive and negative samples are required.")
  }
  
  # Check for NA values
  if (any(is.na(data))) {
    stop("Dataset contains missing values")
  }
  
  return(data)
}
```

```{r}

# Split data into features and target
prepare_matrices <- function(data) {
  # Convert to matrix format required by xgboost
  features <- as.matrix(data[, !names(data) %in% c("Survival")])
  target <- data$Survival
  
  return(list(features = features, target = target))
}

```

```{r}

# Custom evaluation metric for AUC
eval_auc <- function(preds, dtrain) {
  labels <- getinfo(dtrain, "label")
  auc <- roc(labels, preds)$auc
  return(list(metric = "auc", value = auc))
}
```

```{r}
# Parameter grid for tuning
param_grid <- expand.grid(
  max_depth = c(3, 4, 5, 6),
  eta = c(0.01, 0.03, 0.05, 0.1),
  subsample = c(0.7, 0.8, 0.9),
  colsample_bytree = c(0.7, 0.8, 0.9),
  min_child_weight = c(1, 3, 5),
  gamma = c(0, 0.1, 0.3)
)
```

```{r}
# Cross-validation function
cv_xgboost <- function(params, features, target, nfold = 5) {
  dtrain <- xgb.DMatrix(features, label = target)
  
  tryCatch({
    cv_results <- xgb.cv(
      params = params,
      data = dtrain,
      nrounds = 1000,
      nfold = nfold,
      early_stopping_rounds = 50,
      metrics = list("auc"),
      verbose = 0,
      stratified = TRUE,
      prediction = TRUE
    )
    
    # Get the best iteration's AUC
    best_iter <- which.max(cv_results$evaluation_log$test_auc_mean)
    return(list(
      auc = cv_results$evaluation_log$test_auc_mean[best_iter],
      nrounds = best_iter
    ))
    
  }, error = function(e) {
    warning(paste("CV failed with error:", e$message))
    return(list(auc = NA, nrounds = NA))
  })
}

```

```{r}

# Modified main modeling function
train_optimal_model <- function(data) {
  # Preprocess data with validation
  processed_data <- preprocess_data(data)
  prepared_data <- prepare_matrices(processed_data)
  
  # Print diagnostic information
  cat("Data dimensions:", dim(processed_data), "\n")
  cat("Class distribution:", table(processed_data$Survival), "\n")
  
  # Set basic parameters
  base_params <- list(
    objective = "binary:logistic",
    eval_metric = "auc",
    nthread = parallel::detectCores() - 1,
    scale_pos_weight = sum(prepared_data$target == 0) / sum(prepared_data$target == 1)
  )
  
  # Perform grid search with cross-validation
  results <- data.frame()
  
  # Progress counter
  total_combinations <- nrow(param_grid)
  cat("Starting grid search with", total_combinations, "parameter combinations\n")
  
  for(i in 1:nrow(param_grid)) {
    cat(sprintf("\rTesting combination %d/%d", i, total_combinations))
    
    params <- c(
      base_params,
      as.list(param_grid[i,])
    )
    
    cv_result <- cv_xgboost(
      params,
      prepared_data$features,
      prepared_data$target
    )
    
    if (!is.na(cv_result$auc)) {
      results <- rbind(results, 
                      data.frame(param_grid[i,], 
                                auc = cv_result$auc,
                                nrounds = cv_result$nrounds))
    }
  }
  
  cat("\nGrid search completed.\n")
  
  # Check if we have any valid results
  if (nrow(results) == 0) {
    stop("No valid models could be trained with the given parameters")
  }
  
  # Get best parameters
  best_params <- results[which.max(results$auc), ]
  cat("\nBest parameters found:\n")
  print(best_params)
  
  # Train final model with best parameters
  final_params <- c(
    base_params,
    as.list(best_params[1:6])
  )
  
  dtrain <- xgb.DMatrix(prepared_data$features, 
                        label = prepared_data$target)
  
  # Create watchlist for training
  watchlist <- list(train = dtrain)
  
  final_model <- xgb.train(
    params = final_params,
    data = dtrain,
    nrounds = best_params$nrounds,
    early_stopping_rounds = 50,
    watchlist = watchlist,
    verbose = 0
  )
  
  # Get feature importance
  importance <- xgb.importance(
    feature_names = colnames(prepared_data$features),
    model = final_model
  )
  
  return(list(
    model = final_model,
    best_params = best_params,
    feature_importance = importance,
    cv_results = results
  ))
}
```

```{r}
results <- train_optimal_model(Chemo_data)
```

```{r}
#View best parameters
print(results$best_params)
```

```{r}

```

```{r}

# Ensure the data is in the correct format
processed_data <- preprocess_data(Chemo_data)
prepared_data <- prepare_matrices(processed_data)
```

```{r}
# Convert features into an xgb.DMatrix
dtest <- xgb.DMatrix(prepared_data$features)

# Get predictions from the trained model
predictions <- predict(results$model, dtest)

```

```{r}
# Compute ROC curve and AUC
roc_obj <- roc(prepared_data$target, predictions)
auc_value <- auc(roc_obj)
```

```{r}

```

```{r}
# Compute smoothed ROC curve
roc_obj <- roc(prepared_data$target, predictions, smooth = TRUE)

# Extract smoothed curve points
roc_df <- data.frame(
  FPR = 1 - roc_obj$specificities,
  TPR = roc_obj$sensitivities
)

```

```{r}


# Plot smoothed ROC
roc_plot <- ggplot(roc_df, aes(x = FPR, y = TPR)) +
  geom_line(color = "blue", linewidth = 1.5) +
  geom_abline(linetype = "dashed", color = "red") +
  labs(title = paste("ROC Curve (AUC =", round(auc(roc_obj), 3), ")"),
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal()  +
  theme(
    text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  )


# Save the plot as a high-resolution PNG file
ggsave("XGBOOST_ROC_CURVE_NEW.png", 
       plot = roc_plot,
       width = 5,      # Width in inches
       height = 5,     # Height in inches
       dpi = 900)      # Resolution in dots per inch
```

```{r}
roc_plot 
```

#### New with C_I

```{r}

# Revised function to calculate ROC with confidence intervals
calculate_roc_with_ci <- function(labels, predictions, ci_level = 0.95, n_bootstrap = 2000) {
  # Compute smoothed ROC curve
  roc_obj <- roc(labels, predictions, smooth = TRUE)
  
  # Extract smoothed ROC data
  roc_data <- data.frame(FPR = 1 - roc_obj$specificities, 
                         TPR = roc_obj$sensitivities)
  
  # Find optimal cutoff (Youden’s J by default)
  opt_cut <- coords(roc_obj, "best", ret = c("threshold", "sensitivity", "specificity"), best.method = "youden")
  
  # Use bootstrap for CI
  set.seed(42)
  boot_auc <- numeric(n_bootstrap)
  
  indices_class1 <- which(labels == 1)
  indices_class0 <- which(labels == 0)
  
  alpha <- (1 - ci_level) / 2
  fixed_fpr_points <- seq(0, 1, length.out = 100)
  tpr_at_fpr <- matrix(NA, nrow = n_bootstrap, ncol = length(fixed_fpr_points))
  
  for (i in 1:n_bootstrap) {
    bootstrap_class1 <- sample(indices_class1, length(indices_class1), replace = TRUE)
    bootstrap_class0 <- sample(indices_class0, length(indices_class0), replace = TRUE)
    bootstrap_indices <- c(bootstrap_class1, bootstrap_class0)
    
    bootstrap_predictions <- predictions[bootstrap_indices]
    bootstrap_labels <- labels[bootstrap_indices]
    
    boot_roc <- try(roc(bootstrap_labels, bootstrap_predictions), silent = TRUE)
    
    if (!inherits(boot_roc, "try-error")) {
      boot_auc[i] <- auc(boot_roc)
      interp_tpr <- approx(1 - boot_roc$specificities, boot_roc$sensitivities, 
                           fixed_fpr_points, rule = 2)$y
      tpr_at_fpr[i, ] <- interp_tpr
    }
  }
  
  ci_lower_tpr <- apply(tpr_at_fpr, 2, function(x) quantile(x, alpha, na.rm = TRUE))
  ci_upper_tpr <- apply(tpr_at_fpr, 2, function(x) quantile(x, 1 - alpha, na.rm = TRUE))
  
  ci_lower <- data.frame(FPR = fixed_fpr_points, TPR = ci_lower_tpr)
  ci_upper <- data.frame(FPR = fixed_fpr_points, TPR = ci_upper_tpr)
  
  auc_value <- auc(roc_obj)
  auc_ci <- quantile(boot_auc, c(alpha, 1 - alpha), na.rm = TRUE)
  
  return(list(
    roc_data = roc_data,
    ci_lower = ci_lower,
    ci_upper = ci_upper,
    auc = auc_value,
    auc_ci_lower = auc_ci[1],
    auc_ci_upper = auc_ci[2],
    optimal_cut = opt_cut
  ))
}
```

```{r}

# Calculate confidence intervals for PR curve using bootstrap
calculate_pr_with_ci <- function(labels, predictions, n_bootstrap = 2000, ci_level = 0.95) {
  # Original PR curve
  prc_obj <- pr.curve(scores.class0 = predictions, weights.class0 = labels, curve = TRUE)
  
  # Bootstrap for confidence intervals
  precision_matrix <- matrix(nrow = n_bootstrap, ncol = length(prc_obj$curve[, 2]))
  recall_levels <- prc_obj$curve[, 1]
  
  # Get indices for each class
  indices_class1 <- which(labels == 1)
  indices_class0 <- which(labels == 0)
  
  # Calculate quantiles for CI
  alpha <- (1 - ci_level) / 2
  
  # Perform bootstrap
  set.seed(42)  # For reproducibility
  for (i in 1:n_bootstrap) {
    # Resample with replacement from each class
    bootstrap_class1 <- sample(indices_class1, length(indices_class1), replace = TRUE)
    bootstrap_class0 <- sample(indices_class0, length(indices_class0), replace = TRUE)
    
    # Combine the resampled indices
    bootstrap_indices <- c(bootstrap_class1, bootstrap_class0)
    
    # Get bootstrapped predictions and labels
    bootstrap_predictions <- predictions[bootstrap_indices]
    bootstrap_labels <- labels[bootstrap_indices]
    
    # Compute PR curve for this bootstrap sample
    bootstrap_prc <- try(pr.curve(scores.class0 = bootstrap_predictions, 
                                weights.class0 = bootstrap_labels, 
                                curve = TRUE), silent = TRUE)
    
    if (!inherits(bootstrap_prc, "try-error") && length(bootstrap_prc$curve[, 1]) > 0) {
      # Interpolate to get precision values at the original recall levels
      precision_interp <- try(approx(bootstrap_prc$curve[, 1], bootstrap_prc$curve[, 2], 
                                  recall_levels, rule = 2)$y, silent = TRUE)
      
      if (!inherits(precision_interp, "try-error")) {
        precision_matrix[i, ] <- precision_interp
      }
    }
  }
  
  # Calculate confidence intervals for each recall level
  ci_lower <- apply(precision_matrix, 2, function(x) quantile(x, alpha, na.rm = TRUE))
  ci_upper <- apply(precision_matrix, 2, function(x) quantile(x, 1 - alpha, na.rm = TRUE))
  
  # Calculate confidence interval for AUPRC
  auprc_values <- numeric(n_bootstrap)
  for (i in 1:n_bootstrap) {
    bootstrap_class1 <- sample(indices_class1, length(indices_class1), replace = TRUE)
    bootstrap_class0 <- sample(indices_class0, length(indices_class0), replace = TRUE)
    bootstrap_indices <- c(bootstrap_class1, bootstrap_class0)
    
    bootstrap_predictions <- predictions[bootstrap_indices]
    bootstrap_labels <- labels[bootstrap_indices]
    
    pr_result <- try(pr.curve(scores.class0 = bootstrap_predictions, 
                            weights.class0 = bootstrap_labels, 
                            curve = FALSE), silent = TRUE)
    
    if (!inherits(pr_result, "try-error")) {
      auprc_values[i] <- pr_result$auc.integral
    }
  }
  
  auprc_ci <- quantile(auprc_values, c(alpha, 1 - alpha), na.rm = TRUE)
  
  return(list(
    pr_data = data.frame(Recall = recall_levels, Precision = prc_obj$curve[, 2]),
    ci_lower = data.frame(Recall = recall_levels, Precision = ci_lower),
    ci_upper = data.frame(Recall = recall_levels, Precision = ci_upper),
    auprc = prc_obj$auc.integral,
    auprc_ci_lower = auprc_ci[1],
    auprc_ci_upper = auprc_ci[2]
  ))
}
```

```{r}

# Calculate ROC with confidence intervals
roc_with_ci <- suppressMessages(suppressWarnings(calculate_roc_with_ci(prepared_data$target, predictions)))
```

```{r}
# Extract optimal cutoff values safely and coerce to numeric
opt_spec   <- as.numeric(roc_with_ci$optimal_cut[["specificity"]])
opt_sens   <- as.numeric(roc_with_ci$optimal_cut[["sensitivity"]])
opt_thresh <- as.numeric(roc_with_ci$optimal_cut[["threshold"]])

# Enhanced ROC plot with smoothed curve + CI ribbon
roc_plot <- ggplot() +
  geom_ribbon(data = cbind(
    FPR = roc_with_ci$ci_lower$FPR,
    ymin = roc_with_ci$ci_lower$TPR,
    ymax = roc_with_ci$ci_upper$TPR
  ), aes(x = FPR, ymin = ymin, ymax = ymax), fill = "lightblue", alpha = 0.3) +
  geom_line(data = roc_with_ci$roc_data, aes(x = FPR, y = TPR),
            color = "blue", linewidth = 1.5) +
  geom_abline(linetype = "dashed", color = "red") +
  # Add optimal cutoff point
  geom_point(aes(x = 1 - opt_spec, y = opt_sens),
             color = "darkorange", size = 4, shape = 21, fill = "yellow") +
  annotate("text",
           x = 1 - opt_spec, y = opt_sens,
           label = paste0(" ", round(opt_thresh, 3)),
           vjust = -1, hjust = 1.1, size = 5, fontface = "bold", color = "black") +
  labs(title = paste0("(AUC = ", round(roc_with_ci$auc, 3), 
                      ", 95% CI: ", round(roc_with_ci$auc_ci_lower, 3), "-", 
                      round(roc_with_ci$auc_ci_upper, 3), ")"),
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  )


# Save plot
ggsave("Chemo_XGBOOST_ROC_CURVE_WITH_CI_AND_OPTIMAL.png",
       plot = roc_plot, width = 8, height = 6, dpi = 900)

roc_plot
```

```{r}

# Calculate PR curve with confidence intervals
pr_with_ci <- suppressMessages(suppressWarnings(calculate_pr_with_ci(processed_data$Survival, predictions)))

```

```{r}

# Prepare PR curve data properly
pr_data <- pr_with_ci$pr_data
ci_lower <- pr_with_ci$ci_lower
ci_upper <- pr_with_ci$ci_upper

# Create enhanced Precision-Recall plot with confidence interval
prc_plot <- ggplot() +
  # Add confidence interval ribbon
  geom_ribbon(data = data.frame(
    Recall = ci_lower$Recall,
    ymin = ci_lower$Precision,
    ymax = ci_upper$Precision
  ), aes(x = Recall, ymin = ymin, ymax = ymax),
  fill = "lightgreen", alpha = 0.3) +
  
  # Add PR curve
  geom_line(data = pr_data, aes(x = Recall, y = Precision),
            color = "darkgreen", linewidth = 1.5) +
  
  labs(title = paste0("P-R Curve (AUPRC = ", round(pr_with_ci$auprc, 3),
                     ", 95% CI: ", round(pr_with_ci$auprc_ci_lower, 3), "-",
                     round(pr_with_ci$auprc_ci_upper, 3), ")"),
       x = "Recall",
       y = "Precision") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  )

# Save the enhanced PRC plot
ggsave("Chemo_XGBOOST_PRC_CURVE_WITH_CI.png",
       plot = prc_plot,
       width = 8,
       height = 6,
       dpi = 500)



prc_plot
```

```{r}

```

```{r}

library(PRROC)
# Ensure the predictions and actual labels are numeric
predictions <- as.numeric(predictions)
labels <- as.numeric(processed_data$Survival)

# Compute PR curve
prc_obj <- pr.curve(scores.class0 = predictions, weights.class0 = labels, curve = TRUE)

# Print AUPRC score
cat("AUPRC:", round(prc_obj$auc.integral, 3), "\n")
```

```{r}

library(ggplot2)

# Create the Precision-Recall plot with enhanced formatting
prc_plot <- ggplot(data.frame(Recall = prc_obj$curve[, 1], 
                             Precision = prc_obj$curve[, 2]), 
                   aes(x = Recall, y = Precision)) +
  geom_line(color = "darkgreen", linewidth = 1.5) +  
  labs(title = paste("Precision-Recall Curve (AUPRC =", 
                     round(prc_obj$auc.integral, 3), ")"),
       x = "Recall",
       y = "Precision") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  )

# Save the plot as a high-resolution PNG file
ggsave("XGBOOST_PRC_CURVE_NEW.png", 
       plot = prc_plot,
       width = 8,     # Width in inches
       height = 6,     # Height in inches
       dpi = 500)      # Resolution in dots per inch
```

```{r}


```

```{r}

```

```{r}

```

```{r}

```

#### Shapley Values

```{r}

# Required packages
library(tidyverse)
library(caret)
library(xgboost)
library(shapr)
library(Matrix)
library(fastshap)
library(glmnet)
```

```{r}

# Function to prepare data and create models
prepare_and_train_models <- function(data, target_col, features, cor_threshold = 0.8) {
  # Remove Patient ID if present
  if ("Patient" %in% colnames(data)) {
    data <- data[, !colnames(data) %in% "Patient"]
  }
  
  # Convert target to numeric (0/1)
  data[[target_col]] <- as.numeric(factor(data[[target_col]])) - 1
  
  # Validate data
  if (any(is.na(data[[target_col]]))) {
    stop("Target variable contains missing values")
  }
  if (any(is.na(data[features]))) {
    warning("Features contain missing values. They will be handled in preprocessing.")
  }
  
  # Split data into training and test sets
  set.seed(42)
  train_index <- createDataPartition(data[[target_col]], p = 0.75, list = FALSE)
  train_data <- data[train_index, ]
  test_data <- data[-train_index, ]
  
  # Initial scaling of all features
  initial_preprocess <- preProcess(train_data[features], 
                                 method = c("center", "scale", "medianImpute"))
  train_data_scaled <- predict(initial_preprocess, train_data[features])
  test_data_scaled <- predict(initial_preprocess, test_data[features])
  
  # Handle correlations
  cor_matrix <- cor(train_data_scaled)
  high_cor <- findCorrelation(cor_matrix, cutoff = cor_threshold)
  
  if(length(high_cor) > 0) {
    message("Removing highly correlated features: ", 
            paste(features[high_cor], collapse = ", "))
    features_filtered <- features[-high_cor]
  } else {
    features_filtered <- features
  }
  
  # Create new preprocess object with only filtered features
  filtered_preprocess <- preProcess(train_data[features_filtered], 
                                  method = c("center", "scale", "medianImpute"))
  
  # Scale data using filtered features
  train_data_scaled <- predict(filtered_preprocess, train_data[features_filtered])
  test_data_scaled <- predict(filtered_preprocess, test_data[features_filtered])
  
  # Train logistic regression
  x_train <- as.matrix(train_data_scaled)
  y_train <- train_data[[target_col]]
  
  cv_fit <- cv.glmnet(x_train, y_train, family = "binomial", 
                      alpha = 0.5, nfolds = 5)
  
  logistic_model <- glmnet(x_train, y_train, family = "binomial",
                          alpha = 0.5,
                          lambda = cv_fit$lambda.min)
  
  # XGBoost preparation
  dtrain <- xgb.DMatrix(data = x_train, label = y_train)
  dtest <- xgb.DMatrix(data = as.matrix(test_data_scaled), 
                       label = test_data[[target_col]])
  
  xgb_params <- list(
    objective = "binary:logistic",
    eval_metric = "logloss",
    eta = 0.05,
    max_depth = 3,
    min_child_weight = 1,
    subsample = 0.8,
    colsample_bytree = 0.7,
    scale_pos_weight = sum(y_train == 0) / sum(y_train == 1),
    lambda = 2,
    alpha = 1
  )
  
  xgb_model <- xgb.train(
    params = xgb_params,
    data = dtrain,
    nrounds = 150,
    watchlist = list(train = dtrain, test = dtest),
    early_stopping_rounds = 50,
    verbose = 0
  )
  
  return(list(
    train_data = train_data,
    test_data = test_data,
    train_data_scaled = train_data_scaled,
    test_data_scaled = test_data_scaled,
    logistic_model = logistic_model,
    xgb_model = xgb_model,
    features = features_filtered,
    filtered_preprocess = filtered_preprocess,
    diagnostics = list(
      lambda_min = cv_fit$lambda.min,
      xgb_best_iteration = xgb_model$best_iteration
    )
  ))
}
```

```{r}
 


# Function to calculate Shapley values for Logistic


calculate_logistic_shapley <- function(model_output, observation) {
  # Get filtered features
  features_filtered <- model_output$features
  
  # Ensure observation is a data frame and subset with filtered features
  observation <- as.data.frame(observation)  # Convert to data frame
  obs_filtered <- observation[, features_filtered, drop = FALSE]
  
  # Scale the observation using the filtered preprocess object
  obs_scaled <- predict(model_output$filtered_preprocess, obs_filtered)
  
  # Create background data matrix
  background_data <- as.matrix(model_output$train_data_scaled)
  
  # Define the prediction function
  pred_function <- function(object, newdata) {
    predict(
      object, 
      newx = as.matrix(newdata),  # Ensure newdata is a data frame
      type = "response"
    )
  }
  
  # Calculate Shapley values using fastshap
   shapley_values <- fastshap::explain(
    object = model_output$logistic_model,  # Pass the model here
    X = background_data,
    newdata = as.matrix(obs_scaled),
    pred_wrapper = pred_function,
    nsim = 100
  )

  
  # Format results
   shapley_df <- as.data.frame(shapley_values)
  names(shapley_df) <- features_filtered
  
  return(shapley_df)
}
```

```{r}

# Function to calculate Shapley values for XGBoost
calculate_xgb_shapley <- function(model_output, observation) {
  # Get filtered features
  features_filtered <- model_output$features
  
  # Prepare observation
  obs_scaled <- predict(model_output$filtered_preprocess,
                        observation[features_filtered])
  
  # Define prediction wrapper for XGBoost
  pred_wrapper <- function(object, newdata) {
    predict(object, xgb.DMatrix(data = as.matrix(newdata)))
  }
  
  # Calculate Shapley values
  shapley_values <- fastshap::explain(
    object = model_output$xgb_model,  # Pass model here
    X = as.matrix(model_output$train_data_scaled),
    newdata = as.matrix(obs_scaled),
    pred_wrapper = pred_wrapper,
    nsim = 100
  )
  
  return(as.data.frame(shapley_values))
}
```

```{r}

plot_shapley_values <- function(shapley_values, model_type) {

  suppressPackageStartupMessages({
    library(dplyr)
    library(tidyr)
    library(ggplot2)
  })

  # Store original number of observations
  original_rows <- nrow(shapley_values)

  # Convert to long format
  plot_data <- shapley_values %>%
    tidyr::pivot_longer(cols = dplyr::everything(),
                       names_to = "feature",
                       values_to = "shapley_value")

  # Aggregate if needed
  if (original_rows > 1) {
    plot_data <- plot_data %>%
      group_by(feature) %>%
      summarise(shapley_value = sum(shapley_value), .groups = "drop")
    x_label <- "Sum of Shapley Values"
  } else {
    x_label <- "Shapley Value"
  }

  # Remove zero-valued features
  plot_data <- plot_data %>% filter(shapley_value != 0)

  # Round
  plot_data <- plot_data %>% mutate(shapley_value = round(shapley_value, 2))

  # Select top 5 positive & bottom 5 negative
  top5_pos <- plot_data %>% arrange(desc(shapley_value)) %>% head(5)
  bottom5_neg <- plot_data %>% arrange(shapley_value) %>% head(5)

  final_plot_data <- bind_rows(top5_pos, bottom5_neg) %>%
    mutate(
      direction = ifelse(shapley_value >= 0, "Positive", "Negative"),
      feature   = reorder(feature, shapley_value)
    )

  # Create horizontal bar plot (green for +, red for -)
  ggplot(final_plot_data, aes(y = feature, x = shapley_value, fill = direction)) +
    geom_col(width = 0.8) +
    scale_fill_manual(values = c("Positive" = "forestgreen", "Negative" = "red3")) +
    geom_text(
      aes(label = sprintf("%.2f", shapley_value),
          hjust = ifelse(shapley_value >= 0, -0.15, 1.15)),
      size = 3.5,
      color = "black"
    ) +
    scale_x_continuous(expand = expansion(mult = c(0.15, 0.15))) +
    theme_minimal() +
    labs(
      title = paste("Top 5 Positive & Bottom 5 Negative Shapley Features -", model_type),
      x = x_label,
      y = "Features",
      fill = NULL
    ) +
    theme(
      axis.text.y = element_text(size = 10, face = "bold", color = "black",
                                 margin = margin(r = 10)),
      axis.text.x = element_text(size = 12, face = "bold", color = "black"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5,
                                margin = margin(b = 20)),
      axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 15)),
      axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 15)),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      legend.position = "none"  # remove legend to match the example figure vibe
    )
}

```

##### Usage

```{r}

 # Train models
  Shapley_models <- prepare_and_train_models(Chemo_data, "Survival", features)
```

```{r}
# Check diagnostics
print(Shapley_models$diagnostics)
```

```{r}

```

#### Feature Importance

```{r}

```

```{r}
#logistic Regression

set.seed(25)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, Chemo_data)
logistic_shapley
```

```{r}

#XGBoost

set.seed(25)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, Chemo_data)
xgb_shapley

```

```{r}

```

#### Plot

```{r}

png("Shapley_Logistic_All_Patients.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression all patients")
dev.off()

plot_shapley_values(logistic_shapley, "Logistic Regression all patients")
```

```{r}


png("Shapley_XGBOOST_All_Patients.png", width = 4500, height = 3500, res = 900)
par(cex.main = 1, cex.lab = 1, cex.axis = 0.7, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "")
dev.off()
# Plot results
plot_shapley_values(xgb_shapley, "")
```

```{r}

```

#### pCR Patients

```{r}

```

#### Patient N022

```{r}
# Calculate Shapley values for a specific observation
observation <- Chemo_data[4, ]  
observation
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N022.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N022")
dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N022")
```

```{r}


png("Shapley_XGBOOST_N022.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N022")
dev.off()
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N022")
```

```{r}


```

#### Patient N023

```{r}
# Calculate Shapley values for a specific observation
observation2 <- Chemo_data[5, ]  
observation2
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation2)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation2)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N023.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N023")

dev.off()

# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N023")
```

```{r}

png("Shapley_XGBOOST_N023.png",width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N023")

dev.off()
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N023")
```

```{r}

```

#### Patient N113

```{r}
# Calculate Shapley values for a specific observation
observation2 <- Chemo_data[31, ]  
observation2
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation2)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation2)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N113.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N113")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N113")
```

```{r}

png("Shapley_XGBOOST_N113.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N113")

dev.off()

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N113")
```

```{r}

```

#### Patient N152

```{r}
# Calculate Shapley values for a specific observation
observation3 <- Chemo_data[37, ]  
observation3
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation3)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation3)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N152.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N152")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N152")
```

```{r}


png("Shapley_XGBOOST_N152.png", width = 4500, height = 3200, res = 900)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N152")

dev.off()
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N152")
```

```{r}

```

#### Patient N184

```{r}
# Calculate Shapley values for a specific observation
observation4 <- Chemo_data[44, ]  
observation4
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation4)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation4)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N184")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N184")
```

```{r}

```

#### Patient N187

```{r}
# Calculate Shapley values for a specific observation
observation5 <- Chemo_data[45, ]  
observation5
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation5)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation5)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N187")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N187")
```

```{r}

```

#### Patient N197

```{r}
# Calculate Shapley values for a specific observation
observation6 <- Chemo_data[52, ]  
observation6
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation6)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation6)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N197")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N197")
```

```{r}

```

#### Patient N242

```{r}
# Calculate Shapley values for a specific observation
observation7 <- Chemo_data[59, ]  
observation7
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation7)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation7)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N242")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N242")
```

```{r}

```

#### Patient N252

```{r}
# Calculate Shapley values for a specific observation
observation8 <- Chemo_data[66, ]  
observation8
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation8)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation8)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N252")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N252")
```

```{r}


```

#### Patient N260

```{r}
# Calculate Shapley values for a specific observation
observation9 <- Chemo_data[68, ]  
observation9
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation9)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation9)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N260")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N260")
```

```{r}

```

```{r}

```

#### RD Patients

```{r}

```

#### Patient N040

```{r}
# Calculate Shapley values for a specific observation
observation_imm <- Chemo_data[9, ]  
observation_imm
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N040.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N040")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N040")
```

```{r}
png("Shapley_XGBOOST_N040.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N040")

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N040")
```

```{r}

```

```         
```

#### Patient N069

```{r}
# Calculate Shapley values for a specific observation
observation_imm1 <- Chemo_data[16, ]  
observation_imm1
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm1)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm1)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N069.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N069")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N069")
```

```{r}


png("Shapley_XGBOOST_N069.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N069")

dev.off()

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N069")
```

```{r}

```

#### Patient N074

```{r}
# Calculate Shapley values for a specific observation
observation_imm2 <- Chemo_data[19, ]  
observation_imm2
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm2)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm2)
xgb_shapley

```

```{r}

png("Shapley_Logistic_N074.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N074")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N074")
```

```{r}


png("Shapley_XGBOOST_N074.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N074")
dev.off()
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N074")
```

```{r}

```

#### Patient N099

```{r}
# Calculate Shapley values for a specific observation
observation_imm3 <- Chemo_data[27, ]  
observation_imm3
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm3)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm3)
xgb_shapley

```

```{r}


png("Shapley_Logistic_N099.png", width = 4000, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N099")

dev.off()
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N099")
```

```{r}


png("Shapley_XGBOOST_N099.png", width = 4500, height = 3200, res = 500)
par(cex.main = 1, cex.lab = 1, cex.axis = 1, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N099")

dev.off()

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N099")
```

```{r}

```

#### Patient N127

```{r}
# Calculate Shapley values for a specific observation
observation_imm4 <- Chemo_data[34, ]  
observation_imm4
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm4)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm4)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N127")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N127")
```

```{r}

```

#### Patient N172

```{r}
# Calculate Shapley values for a specific observation
observation_imm5 <- Chemo_data[42, ]  
observation_imm5
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm5)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm5)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N172")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N172")
```

```{r}

```

#### Patient N178

```{r}
# Calculate Shapley values for a specific observation
observation_imm6 <- Chemo_data[43, ]  
observation_imm6
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm6)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm6)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N178")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N178")
```

```{r}

```

#### Patient N193

```{r}
# Calculate Shapley values for a specific observation
observation_imm7 <- Chemo_data[50, ]  
observation_imm7
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm7)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm7)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N193")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N193")
```

```{r}

```

#### Patient N248

```{r}
# Calculate Shapley values for a specific observation
observation_imm8 <- Chemo_data[63, ]  
observation_imm8
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm8)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm8)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N248")
```

```{r}

png("Shapley_XGBOOST_N248.png", width = 4500, height = 3200, res = 900)
par(cex.main = 1, cex.lab = 1, cex.axis = 0.7, font.main = 1, font.lab = 1)
# Plot results
plot_shapley_values(xgb_shapley, "")
dev.off()
# Plot results
# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N248")
```

```{r}

```

#### Patient N172

```{r}
# Calculate Shapley values for a specific observation
observation_imm5 <- Chemo_data[42, ]  
observation_imm5
```

```{r}
set.seed(21)
logistic_shapley <- calculate_logistic_shapley(Shapley_models, observation_imm5)
print(logistic_shapley)
```

```{r}

set.seed(21)
xgb_shapley <- calculate_xgb_shapley(Shapley_models, observation_imm5)
xgb_shapley

```

```{r}
# Plot results
plot_shapley_values(logistic_shapley, "Logistic Regression Patient N172")
```

```{r}

# Plot results
plot_shapley_values(xgb_shapley, "XGBoost Patient N172")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
