---
title: "R Notebook"
output: html_notebook
Author: Daniel Marri
Task: New_Analysis
---

```{r}

setwd('V:/dkmarri/Breast_Cancer_IMC_Data/R_Codes')
```

```{r}

library(imcRtools)
library(cytomapper)
library(tidyverse)
library(EBImage)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(dplyr)
library(future)
library(progress)
library(Matrix)
library(RANN)
library(reticulate)
library(igraph)
library(tidyr)
library(dittoSeq)
library(scater)
library(patchwork)
library(cowplot)
library(viridis)
library(parallel) 
library(spdep)
library(spatialHeatmap)
library(igraph)
library(RANN)
library(data.table)
library(Matrix)
library(parallel)
library(furrr)
library(scran)
library(batchelor)
library(BiocParallel)
library(scater)  # For PCA, if not already installed
library(SummarizedExperiment)
library(sva)
library(harmony)
library(RColorBrewer)
library(BiocSingular)
library(scales)
library(ggraph)
library(ComplexHeatmap)
library(circlize)
```

```{r}

spe <- readRDS("saved_data/spatial.rds")
spe
```

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy <- spe[, colData(spe)$Arm == "C"]

```

```{r}
# Subsetting based on the column 'Arm' Chemotherapy

spe_chemo_Immu <- spe[, colData(spe)$Arm == "C&I"]
```

```{r}
## Run PCA on the Chemo data 

# 2. Run PCA with proper parameters
spe_chemotherapy <- runPCA(spe_chemotherapy, 
                           exprs_values = "exprs",   
                           ncomponents = 10, 
                           BSPARAM = ExactParam(), 
                           scale = TRUE,
                           name = "PCA")
```

```{r}

## Run PCA on the Chemo_Immu data 

spe_chemo_Immu <-  runPCA(spe_chemo_Immu, exprs_values = "exprs",   ncomponents = 10, BSPARAM = ExactParam(), scale = TRUE, name = "PCA") 


```

```{r}
##Run Harmony correction on both dataset.
# Start with Chemo data

sce_chemotherapy <- as(spe_chemotherapy, "SingleCellExperiment")

```

```{r}
# Perform batch correction using Harmony
# Define 'batch' as the column name from colData that stores batch information
sce_chemotherapy_harmony <- RunHarmony(
  object = sce_chemotherapy, 
  group.by.vars = "PatientID",   # Replace with your batch variable (e.g., 'batch', 'sample')
  assay.use = "exprs",   # Ensure you're using normalized counts for correction
  reduction = "PCA",         # Use PCA as input to Harmony            
  BSPARAM = ExactParam() # Number of dimensions to use from PCA
)

```

```{r}
# Update the reduced dimensions in the original SpatialExperiment object
reducedDim(spe_chemotherapy, "Harmony") <- reducedDim(sce_chemotherapy_harmony, "PCA")
spe_chemotherapy
```

```{r}

# End with chemo_Immu data

sce_chemo_Immu <- as(spe_chemo_Immu, "SingleCellExperiment")
```

```{r}

# Define 'batch' as the column name from colData that stores batch information
sce_chemo_Immu_harmony <- RunHarmony(
  object = sce_chemo_Immu, 
  group.by.vars = "PatientID",   # Replace with your batch variable (e.g., 'batch', 'sample')
  assay.use = "exprs",   # Ensure you're using normalized counts for correction
  reduction = "PCA",         # Use PCA as input to Harmony            
  BSPARAM = ExactParam() # Number of dimensions to use from PCA
)

```

```{r}

# Update the reduced dimensions in the original SpatialExperiment object
reducedDim(spe_chemo_Immu, "Harmony") <- reducedDim(sce_chemo_Immu_harmony, "PCA")
spe_chemo_Immu
```

```{r}
param <- SnowParam(workers = 4)
#Run UMAP using harmony on Chemo_data

# Modify the runUMAP call with adjusted parameters
spe_chemotherapy <- runUMAP(
  spe_chemotherapy,
  dimred = "Harmony", 
  name = "UMAP_mnnCorrected",
  n_neighbors = 15,        # Reducing the neighbors from the default (15 is a good starting point)
      # Set to 2 components for 2D UMAP
  min_dist = 0.5,          # Increase the min_dist slightly to make computation faster
  BPPARAM = param          # Add parallelization
)

```

```{r}

# Create a random sample of 20% of cells
sample_size <- round(ncol(spe_chemotherapy) * 0.1)
sampled_cells <- sample(colnames(spe_chemotherapy), sample_size)
```

```{r}
# Subset the SPE object
spe_chemotherapy_subset <- spe_chemotherapy[, sampled_cells]
```

```{r}
p1 <- dittoDimPlot(spe_chemotherapy, var = "BiopsyPhase", 
                   reduction.use = "UMAP_mnnCorrected", size = 0.2) + 
    ggtitle("Arm on UMAP before correction")

p1
```

```{r}

```

```{r}

library(scran)
library(uwot)
library(ggplot2)

# Function to generate UMAP with error handling
generate_umap_plot <- function(spe, reduction_method = "Harmony") {
  # Error checks
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform UMAP using uwot package
  set.seed(123)  # For reproducibility
  umap_result <- uwot::umap(rd, n_neighbors = 15, min_dist = 0.5)
  
  # Create a data frame with UMAP coordinates and BiopsyPhase
  umap_df <- data.frame(
    UMAP1 = umap_result[,1],
    UMAP2 = umap_result[,2],
    BiopsyPhase = colData(spe)$BiopsyPhase
  )
  
  # Generate UMAP plot with ggplot2
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = BiopsyPhase)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = paste("Chemotherapy UMAP of BiopsyPhase"),
      #subtitle = "Colored by Biopsy Phase",
      x = "UMAP 1",
      y = "UMAP 2"
    )  +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      # plot.subtitle = element_text(hjust = 0.5, size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_brewer(palette = "Set2", name = "Biopsy Phase")
  
  # Add statistical summary
  phase_counts <- table(umap_df$BiopsyPhase)
  summary_text <- paste(
    "Biopsy Phase Distribution:\n",
    paste(names(phase_counts), ":", phase_counts, collapse = "\n")
  )
  
  # Combine plot with summary
  final_plot <- p + 
    annotate("text", x = Inf, y = -Inf, 
             label = summary_text, 
             hjust = 1.1, vjust = -0.5, 
             size = 3, 
             lineheight = 0.8)
  
  return(final_plot)
}
```

```{r}
# Generate UMAP plot
umap_plot <- generate_umap_plot(spe_chemotherapy_subset)
# Display the plot
print(umap_plot)


# Save the plot
ggsave("Chemotherapy_Umap_by_Biopsyphase.png", 
       umap_plot, 
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )
```

```{r}

# Clustering function
cluster_by_biopsy_phase <- function(spe, reduction_method = "Harmony") {
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each BiopsyPhase
  biopsy_phases <- unique(colData(spe)$BiopsyPhase)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(biopsy_phases)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$BiopsyPhase == biopsy_phases[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(biopsy_phases))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[biopsy_phases[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    BiopsyPhase = rep(names(phase_coordinates), 
                      sapply(phase_coordinates, nrow))
  )
  
  # Generate plot with enhanced styling
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = BiopsyPhase)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy  Cluster by BiopsyPhase",
      # subtitle = "Distinct Spatial Clustering",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      # plot.subtitle = element_text(hjust = 0.5, size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_brewer(palette = "Set2", name = "Biopsy Phase")
  
  # # Add phase distribution information
   phase_counts <- table(umap_df$BiopsyPhase)
   summary_text <- paste(
     "Biopsy Phase Distribution:\n",
     paste(names(phase_counts), ":", phase_counts, collapse = "\n")
  )
  
  # Combine plot with summary
  final_plot <- p + 
    annotate("text", x = Inf, y = -Inf, 
             label = summary_text, 
             hjust = 1.1, vjust = -0.5, 
             size = 3, 
             lineheight = 0.8)
  
  return(final_plot)
}
```

```{r}
# Generate clustered UMAP plot
clustered_umap_plot <- cluster_by_biopsy_phase(spe_chemotherapy_subset)

# Display the plot
print(clustered_umap_plot)



# Save the plot
ggsave("Chemotherapy_Umap_Cluster_by_Biopsyphase.png", 
       clustered_umap_plot, 
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )
```

```{r}

```

```{r}

# Create the plot with the subsetted data
p1b <- dittoDimPlot(spe_chemotherapy_subset, 
                   var = "BiopsyPhase",
                   reduction.use = "UMAP_mnnCorrected", 
                   size = 0.2) +  # Increased point size since we have fewer points
  ggtitle("Arm on UMAP (20% sample)")

p1b 
```

```{r}

```

```{r}
markers1 <- c("GATA3", "CD8", "CD20", "CD3", "FOXP3", "MPO", "SMA", "Ki67")
```

```{r}

plot_list <- multi_dittoDimPlot(spe_chemotherapy_subset, var = markers1, reduction.use = "UMAP_mnnCorrected", assay = "exprs", size = 0.2, list.out = TRUE) 

# Apply a color scale and customize the color bar for each plot
plot_list <- lapply(plot_list, function(x) {
  x + 
    scale_color_viridis_c() +  # Use scale_color_viridis_c() for continuous color scale
    theme(
      legend.title = element_text(size = 12),  # Customize legend title
      legend.text = element_text(size = 10),   # Customize legend text
      legend.position = "right"                 # Position the legend
    ) +
    guides(color = guide_colorbar(title = "Expression", title.position = "top", 
                                  title.hjust = 0.5, barwidth = 1.5, barheight = 10))  # Customize color bar
})

plot_grid(plotlist = plot_list) 

```

```{r}
### Run UMAP FOR CHEM_IMMU


# Modify the runUMAP call with adjusted parameters
spe_chemo_Immu <- runUMAP(
  spe_chemo_Immu,
  dimred = "Harmony", 
  name = "UMAP_mnnCorrected",
  n_neighbors = 15,        # Reducing the neighbors from the default (15 is a good starting point)
      # Set to 2 components for 2D UMAP
  min_dist = 0.5,          # Increase the min_dist slightly to make computation faster
  BPPARAM = param          # Add parallelization
)

spe_chemo_Immu
```

```{r}

p2 <- dittoDimPlot(spe_chemo_Immu, var = "BiopsyPhase", 
                   reduction.use = "UMAP_mnnCorrected", size = 0.2) + 
    ggtitle("Arm on UMAP before correction")

p2
```

```{r}
# Create a random sample of 20% of cells
sample_size1 <- round(ncol(spe_chemo_Immu) * 0.1)
sampled_cells1 <- sample(colnames(spe_chemo_Immu), sample_size1)
```

```{r}
# Subset the SPE object
spe_chemo_Immu_subset <- spe_chemo_Immu[, sampled_cells1]
```

```{r}

```

```{r}
# Function to generate UMAP with error handling
generate_umap_plot1 <- function(spe, reduction_method = "Harmony") {
  # Error checks
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform UMAP using uwot package
  set.seed(123)  # For reproducibility
  umap_result <- uwot::umap(rd, n_neighbors = 15, min_dist = 0.5)
  
  # Create a data frame with UMAP coordinates and BiopsyPhase
  umap_df <- data.frame(
    UMAP1 = umap_result[,1],
    UMAP2 = umap_result[,2],
    BiopsyPhase = colData(spe)$BiopsyPhase
  )
  
  # Generate UMAP plot with ggplot2
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = BiopsyPhase)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = paste("Chemotherapy_&_Immunotherapy UMAP of BiopsyPhase"),
      #subtitle = "Colored by Biopsy Phase",
      x = "UMAP 1",
      y = "UMAP 2"
    )  +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      # plot.subtitle = element_text(hjust = 0.5, size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_brewer(palette = "Set2", name = "Biopsy Phase")
  
  # Add statistical summary
  phase_counts <- table(umap_df$BiopsyPhase)
  summary_text <- paste(
    "Biopsy Phase Distribution:\n",
    paste(names(phase_counts), ":", phase_counts, collapse = "\n")
  )
  
  # Combine plot with summary
  final_plot <- p + 
    annotate("text", x = Inf, y = -Inf, 
             label = summary_text, 
             hjust = 1.1, vjust = -0.5, 
             size = 3, 
             lineheight = 0.8)
  
  return(final_plot)
}
```

```{r}

# Generate UMAP plot
umap_plot1 <- generate_umap_plot1(spe_chemo_Immu_subset)
# Display the plot
print(umap_plot1)


# Save the plot
ggsave("Chemotherapy_&_Immunotherapy_Umap_Cluster_by_Biopsyphase.png", 
       umap_plot1, 
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )
```

```{r}

```

```{r}
# Clustering function
cluster_by_biopsy_phase1 <- function(spe, reduction_method = "Harmony") {
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each BiopsyPhase
  biopsy_phases <- unique(colData(spe)$BiopsyPhase)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(biopsy_phases)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$BiopsyPhase == biopsy_phases[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(biopsy_phases))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[biopsy_phases[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    BiopsyPhase = rep(names(phase_coordinates), 
                      sapply(phase_coordinates, nrow))
  )
  
  # Generate plot with enhanced styling
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = BiopsyPhase)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy_&_Immunotherapy Cluster by BiopsyPhase",
      # subtitle = "Distinct Spatial Clustering",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      # plot.subtitle = element_text(hjust = 0.5, size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_brewer(palette = "Set2", name = "Biopsy Phase")
  
  # # Add phase distribution information
   phase_counts <- table(umap_df$BiopsyPhase)
   summary_text <- paste(
     "Biopsy Phase Distribution:\n",
     paste(names(phase_counts), ":", phase_counts, collapse = "\n")
  )
  
  # Combine plot with summary
  final_plot <- p + 
    annotate("text", x = Inf, y = -Inf, 
             label = summary_text, 
             hjust = 1.1, vjust = -0.5, 
             size = 3, 
             lineheight = 0.8)
  
  return(final_plot)
}
```

```{r}
# Generate clustered UMAP plot
clustered_umap_plot1 <- cluster_by_biopsy_phase1(spe_chemo_Immu_subset)

# Display the plot
print(clustered_umap_plot1)


# Save the plot
ggsave("Chemotherapy_&_Immunotherapy_Umap_Cluster_by_Biopsyphase.png", 
       clustered_umap_plot1, 
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )
```

```{r}

```

```{r}

```

```{r}

```

```{r}
p2a <- dittoDimPlot(spe_chemo_Immu_subset, var = "BiopsyPhase", 
                   reduction.use = "UMAP_mnnCorrected", size = 0.2) + 
    ggtitle("Arm on UMAP before correction")

p2a
```

```{r}

plot_list <- multi_dittoDimPlot(spe_chemo_Immu_subset, var = markers1, reduction.use = "UMAP_mnnCorrected", assay = "exprs", size = 0.2, list.out = TRUE) 

# Apply a color scale and customize the color bar for each plot
plot_list <- lapply(plot_list, function(x) {
  x + 
    scale_color_viridis_c() +  # Use scale_color_viridis_c() for continuous color scale
    theme(
      legend.title = element_text(size = 12),  # Customize legend title
      legend.text = element_text(size = 10),   # Customize legend text
      legend.position = "right"                 # Position the legend
    ) +
    guides(color = guide_colorbar(title = "Expression", title.position = "top", 
                                  title.hjust = 0.5, barwidth = 1.5, barheight = 10))  # Customize color bar
})

plot_grid(plotlist = plot_list) 

```

```{r}

memory.limit(size=10000)
```

```{r}
## Save data

#saveRDS(spe_chemotherapy, "saved_data/spe_chemotherapy_batched_spatial.rds" )
```

```{r}
## Save data

#saveRDS(spe_chemo_Immu, "saved_data/spe_chemo_Immu_batched_spatial.rds" )
```

```{r}
```

### Import the batched dataset

```{r}

spe_chemotherapy <- readRDS("saved_data/transformed_spe_chemo_Spatial_Graph.rds")
spe_chemotherapy
```

```{r}
spe_chemo_Immu <- readRDS("saved_data/transformed_spe_chemo_immu_Spatial_Graph.rds")
spe_chemo_Immu
```

```{r}

```

```         
```

#### Plot the umap to show all the cell types for both treatment arms

```{r}

p3 <- dittoDimPlot(
 spe_chemotherapy, 
  var = "Label", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2,
  do.raster = TRUE,
  legend.show = TRUE) +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cell_types_umap_for_chemo.png", 
  plot = p3,
  width = 3600,    # Width in inches
  height = 2400,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)

# Display the Plot
print(p3)

# Optional: print confirmation message
cat("UMAP plot saved as cell_types_umap.png\n")
```

```{r}
# Create a random sample of 20% of cells
sample_size <- round(ncol(spe_chemotherapy) * 0.1)
sampled_cells <- sample(colnames(spe_chemotherapy), sample_size)
```

```{r}
# Subset the SPE object
spe_chemotherapy_subset1 <- spe_chemotherapy[, sampled_cells]
```

```{r}

p3a <- dittoDimPlot(
spe_chemotherapy_subset1, 
  var = "Label", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2,
  do.raster = TRUE,
  legend.show = TRUE) +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cell_types_umap_for_chemo_a.png", 
  plot = p3a,
  width = 3600,    # Width in inches
  height = 2700,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)

# Display the Plot
print(p3a)

# Optional: print confirmation message
cat("UMAP plot saved as cell_types_umap.png\n")
```

```{r}

```

```{r}

# Clustering function with comprehensive color palette
cluster_by_cell_type <- function(spe, reduction_method = "Harmony") {
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each BiopsyPhase
  cell_type <- unique(colData(spe)$Label)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(cell_type)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$Label == cell_type[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(cell_type))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[cell_type[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    Label = rep(names(phase_coordinates), 
                      sapply(phase_coordinates, nrow))
  )
  
  # Create a comprehensive color palette
  # Use a combination of RColorBrewer and manual colors to ensure enough distinct colors
  library(RColorBrewer)
  library(grDevices)
  
  # Get a base set of colors from multiple palettes
  base_colors <- c(
    brewer.pal(12, "Paired"),
    brewer.pal(9, "Set1"),
    brewer.pal(8, "Set2"),
    brewer.pal(9, "Set3"),
    colorRampPalette(c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", 
                       "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf"))(20)
  )
  
  # Remove duplicates and ensure enough unique colors
  unique_colors <- unique(base_colors)
  
  # Generate plot with comprehensive color palette
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = Label)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy Cluster by Cell Type",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 10),
      legend.position = "right"
    ) +
    scale_color_manual(values = unique_colors[1:length(cell_type)], name = "Cell Type")
  
  return(p)
}
```

```{r}


# Generate clustered UMAP plot
clustered_umap_plot_celltype <- cluster_by_cell_type(spe_chemotherapy_subset1)

# Display the plot
print(clustered_umap_plot_celltype)


# Save the plot
ggsave("Chemotherapy_Umap_Cluster_by_Cell_Type.png",
       clustered_umap_plot_celltype,
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )

```

```{r}

```

```{r}


```

```{r}


p4 <- dittoDimPlot(
 spe_chemo_Immu, 
  var = "Label", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE) +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cell_types_umap_for_chemo_immu.png", 
  plot = p4,
  width = 3600,    # Width in inches
  height = 24700,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)


# Display the Plot
print(p4)
```

```{r}
# Create a random sample of 20% of cells
sample_size1 <- round(ncol(spe_chemo_Immu) * 0.1)
sampled_cells1 <- sample(colnames(spe_chemo_Immu), sample_size1)
```

```{r}
# Subset the SPE object
spe_chemo_Immu_subset1 <- spe_chemo_Immu[, sampled_cells1]
```

```{r}

p4a <- dittoDimPlot(
 spe_chemo_Immu_subset1, 
  var = "Label", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE) +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cell_types_umap_for_chemo_immu_a.png", 
  plot = p4a,
  width = 3600,    # Width in inches
  height = 2700,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)


# Display the Plot
print(p4a)
```

```{r}

```

```{r}
# Clustering function with comprehensive color palette
cluster_by_cell_type1 <- function(spe, reduction_method = "Harmony") {
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each BiopsyPhase
  cell_type <- unique(colData(spe)$Label)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(cell_type)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$Label == cell_type[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(cell_type))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[cell_type[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    Label = rep(names(phase_coordinates), 
                      sapply(phase_coordinates, nrow))
  )
  
  # Create a comprehensive color palette
  # Use a combination of RColorBrewer and manual colors to ensure enough distinct colors
  library(RColorBrewer)
  library(grDevices)
  
  # Get a base set of colors from multiple palettes
  base_colors <- c(
    brewer.pal(12, "Paired"),
    brewer.pal(9, "Set1"),
    brewer.pal(8, "Set2"),
    brewer.pal(9, "Set3"),
    colorRampPalette(c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", 
                       "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf"))(20)
  )
  
  # Remove duplicates and ensure enough unique colors
  unique_colors <- unique(base_colors)
  
  # Generate plot with comprehensive color palette
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = Label)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy_&_Immunotherapy Cluster by Cell Type",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 10),
      legend.position = "right"
    ) +
    scale_color_manual(values = unique_colors[1:length(cell_type)], name = "Cell Type")
  
  return(p)
}
```

```{r}

# Generate clustered UMAP plot
clustered_umap_plot_celltype1 <- cluster_by_cell_type1(spe_chemo_Immu_subset1)

# Display the plot
print(clustered_umap_plot_celltype1)


# Save the plot
ggsave("Chemotherapy_&_Immunotherapy_Umap_Cluster_by_Cell_Type.png",
       clustered_umap_plot_celltype1,
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )

```

```{r}

```

```{r}

```

```{r}

color_vectors <- list()

#For Biopsy phase
BiopsyPhase <- setNames(brewer.pal(length(unique(spe_chemotherapy$BiopsyPhase)), name = "BrBG"), unique(spe_chemotherapy$BiopsyPhase))


#For Arm
Arm <- setNames(c(brewer.pal(6, "YlOrRd")[3:5],
                        brewer.pal(6, "PuBu")[3:6]), unique(spe_chemotherapy$Arm))



# PatientID

num_colors <- length(unique(spe_chemotherapy$PatientID))
custom_palette <- hue_pal()(num_colors)
PatientID <- setNames(custom_palette, unique(spe_chemotherapy$PatientID))


color_vectors$BiopsyPhase <- BiopsyPhase
color_vectors$Arm <- Arm
color_vectors$PatientID <- PatientID

metadata(spe_chemotherapy)$color_vectors <- color_vectors
```

```{r}
color_vectors <- list()

#For Biopsy phase
BiopsyPhase <- setNames(brewer.pal(length(unique(spe_chemo_Immu$BiopsyPhase)), name = "BrBG"), unique(spe_chemo_Immu$BiopsyPhase))


#For Arm
Arm <- setNames(c(brewer.pal(6, "YlOrRd")[3:5],
                        brewer.pal(6, "PuBu")[3:6]), unique(spe_chemo_Immu$Arm))



# PatientID

num_colors <- length(unique(spe_chemo_Immu$PatientID))
custom_palette <- hue_pal()(num_colors)
PatientID <- setNames(custom_palette, unique(spe_chemo_Immu$PatientID))


color_vectors$BiopsyPhase <- BiopsyPhase
color_vectors$Arm <- Arm
color_vectors$PatientID <- PatientID

metadata(spe_chemo_Immu)$color_vectors <- color_vectors
```

```{r}



assay(spe_chemotherapy, "log_exprs") <- log1p(assay(spe_chemotherapy, "exprs"))

# Heatmap visualization - DittoHeatmap
dittoHeatmap(spe_chemotherapy, 
             genes = rownames(spe_chemotherapy),
             assay = "exprs", 
             cluster_cols = FALSE, 
             scale = "row",
             heatmap.colors = viridis(100),
             legend.width = 5,
             annot.by = c("BiopsyPhase", "Label"),
             annotation_colors = list(BiopsyPhase = metadata(spe_chemotherapy)$color_vectors$BiopsyPhase,
                                      Arm = metadata(spe_chemotherapy)$color_vectors$Arm,
                                      PatientID = metadata(spe_chemotherapy)$color_vectors$PatientID))


```

```{r}

assay(spe_chemo_Immu, "log_exprs") <- log1p(assay(spe_chemo_Immu, "exprs"))

# Heatmap visualization - DittoHeatmap
dittoHeatmap(spe_chemo_Immu, 
             genes = rownames(spe_chemo_Immu),
             assay = "log_exprs", 
             cluster_cols = FALSE, 
             scale = "row",
             heatmap.colors = viridis(100), 
             legend.width = 5,
             annot.by = c("BiopsyPhase", "Label"),
             annotation_colors = list(BiopsyPhase = metadata(spe_chemo_Immu)$color_vectors$BiopsyPhase,
                                      Arm = metadata(spe_chemo_Immu)$color_vectors$Arm,
                                      PatientID = metadata(spe_chemo_Immu)$color_vectors$PatientID))


```

```{r}
type_markers <- rownames(spe_chemotherapy)
```

```{r}
library(scuttle)

## aggregate by cell type
celltype_mean <- aggregateAcrossCells(as(spe_chemotherapy, "SingleCellExperiment"),  
                     ids = spe_chemotherapy$Label, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = type_markers)



# Step 3: Ensure cell type metadata is correctly assigned
# Use unique 'spe$Label' values for cell type annotations
colData(celltype_mean)$celltype <- factor(unique(spe_chemotherapy$Label))

# Step 4: Count the number of cells per aggregated cell type
colData(celltype_mean)$ncells <- as.numeric(table(spe_chemotherapy$Label)[unique(spe_chemotherapy$Label)])

# Step 5: Plot the heatmap using dittoHeatmap
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scale = "row",                   # Scale by row (gene-wise)
             heatmap.colors = viridis(100),    # Use viridis color scale
             annot.by = c("celltype"),
             fontsize_row = 6,                 # Adjust the font size for type markers
             legend_breaks = seq(1, 37),       # Show all 37 cell types in the legend
             show_legend = TRUE,               # Ensure the legend is visible
             legend.position = "right",        # Position the legend on the right margin
             legend.width = 5
             
             )  # Annotate by 'celltype' and 'ncells'
```

```{r}
# Aggregate by cell type
celltype_mean <- aggregateAcrossCells(
  as(spe_chemotherapy, "SingleCellExperiment"),
  ids = spe_chemotherapy$Label,
  statistics = "mean",
  use.assay.type = "exprs",
  subset.row = type_markers
)

# Assign cell type metadata
colData(celltype_mean)$celltype <- factor(unique(spe_chemotherapy$Label))

# Count cells per aggregated cell type
colData(celltype_mean)$ncells <- as.numeric(table(spe_chemotherapy$Label)[unique(spe_chemotherapy$Label)])

# Open PNG device with high resolution
png("cell_type_heatmap_for chemo.png", 
    width = 3600,     # Wide width in pixels
    height = 2700,    # Tall height in pixels 
    res = 310,        # High resolution (300 dpi)
    pointsize = 12)   # Slightly larger point size for readability


# Create the heatmap with updated parameters
dittoHeatmap(
  celltype_mean,
  assay = "exprs",
  cluster_cols = TRUE,
  scale = "row",
  heatmap.colors = viridis(100),
  annot.by = c("celltype"),
  fontsize_row = 12,
  fontsize_col = 12,  # Control cell type label size
  show_colnames = TRUE,  # Show all cell type names
  do.raster = TRUE,
  legend.show = TRUE,
  bold_colnames = TRUE,  # Make cell type labels bold
  breaks = seq(0, 1, length.out = 101),  # Explicitly define color breaks
  legend = NULL,  # Remove custom legend settings
  show_legend = TRUE,
  legend.position = "right", 
  row_font_gp = gpar(fontface = "bold"),  # Bold row (gene) labels
  column_font_gp = gpar(fontface = "bold") 
)

# Close the PNG device
dev.off()
```

```{r}

# Scaled to max
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scaled.to.max = TRUE,
             heatmap.colors.max.scaled = inferno(100),
             annot.by = c("celltype"),
             fontsize_row = 6,                 # Adjust the font size for type markers
             legend_breaks = 6,                # Show all color breaks for cell types
             show_legend = TRUE
            )
```

```{r}

# Z score scaled
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             annot.by = c("celltype"),
             )
```

```{r}

```

### For Chemo_Immu

```{r}

## aggregate by cell type
celltype_mean <- aggregateAcrossCells(as(spe_chemo_Immu, "SingleCellExperiment"),  
                     ids = spe_chemo_Immu$Label, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = type_markers)



# Step 3: Ensure cell type metadata is correctly assigned
# Use unique 'spe$Label' values for cell type annotations
colData(celltype_mean)$celltype <- factor(unique(spe_chemo_Immu$Label))

# Step 4: Count the number of cells per aggregated cell type
colData(celltype_mean)$ncells <- as.numeric(table(spe_chemo_Immu$Label)[unique(spe_chemo_Immu$Label)])

# Step 5: Plot the heatmap using dittoHeatmap
# Open PNG device with high resolution
png("cell_type_heatmap_for_chemo_immu.png", 
    width = 3600,     # Wide width in pixels
    height = 2700,    # Tall height in pixels 
    res = 310,        # High resolution (300 dpi)
    pointsize = 12)   # Slightly larger point size for readability


# Create the heatmap with updated parameters
dittoHeatmap(
  celltype_mean,
  assay = "exprs",
  cluster_cols = TRUE,
  scale = "row",
  heatmap.colors = viridis(100),
  annot.by = c("celltype"),
  fontsize_row = 10,
  fontsize_col = 10,  # Control cell type label size
  show_colnames = TRUE,  # Show all cell type names
  do.raster = TRUE,
  legend.show = TRUE,
  bold_colnames = TRUE,  # Make cell type labels bold
  breaks = seq(0, 1, length.out = 101),  # Explicitly define color breaks
  legend = NULL,  # Remove custom legend settings
  show_legend = TRUE,
  legend.position = "right", 
  row_font_gp = gpar(fontface = "bold"),  # Bold row (gene) labels
  column_font_gp = gpar(fontface = "bold") 
)

# Close the PNG device
dev.off()
```

```{r}

# Scaled to max
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scaled.to.max = TRUE,
             heatmap.colors.max.scaled = inferno(100),
             annot.by = c("celltype"),
             fontsize_row = 6,                 # Adjust the font size for type markers
             legend_breaks = 6,                # Show all color breaks for cell types
             show_legend = TRUE
            )
```

```{r}

# Z score scaled
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             annot.by = c("celltype"),
             )
```

```{r}

```

#### Classification of various cell types

```{r}
library(tidyverse)
library(kableExtra)
```

```{r}

# Create a lookup table (named vector) to map daughter cells to parent cells
daughter_to_parent <- c(
  "Fibroblasts" = "Fibroblasts",
  "DCs" = "Other",
  "CA9^+" = "Cancer Cells",
  "PDPN^+Stromal" = "Other",
  "M2 Mac" = "Macrophages",
  "Myofibroblasts" = "Fibroblasts",
  "CD79a^+Plasma" = "Other",
  "CK8/18^{med}" = "Cancer Cells",
  "CD56^+NK" = "Other",
  "CD4^+TCF1^+T" = "Effector T Cells",
  "panCK^{med}" = "Cancer Cells",
  "Treg" = "Regulatory T Cells",
  "CD8^+T" = "Effector T Cells",
  "CD8^+TCF1^+T" = "Effector T Cells",
  "CD8^+PD1^+T_{Ex}" = "Effector T Cells",
  "Endothelial" = "Endothelial Cells",
  "CD4^+PD1^+T" = "Effector T Cells",
  "CD8^+GZMB^+T" = "Effector T Cells",
  "PD-L1^+APCs" = "Other",
  "CD15^{+}" = "Other",
  "PD-L1^{+}GZMB^{+}" = "Other",
  "pH2AX^{+}DSB" = "Other",
  "Vimentin^{+}EMT" = "Cancer Cells",
  "CK^{hi}GATA3^{+}" = "Cancer Cells",
  "CK^{lo}GATA3^{+}" = "Cancer Cells",
  "Helios^{+}" = "Regulatory T Cells",
  "Basal" = "Cancer Cells",
  "Apoptosis" = "Other",
  "Neutrophils" = "Other",
  "CD20^+B" = "Other",
  "PD-L1^{+}IDO^{+}" = "Other",
  "PD-L1^+IDO^+APCs" = "Other",
  "AR^{+}LAR" = "Cancer Cells",
  "TCF1^{+}" = "Effector T Cells",
  "MHC I&II^{hi}" = "Other",
  "CD56^{+}NE" = "Other",
  "CA9^{+}Hypoxia" = "Cancer Cells"
)
```

```{r}
# Step 2: Convert the vector into a data frame
daughter_to_parent_df <- data.frame(
  Daughter = names(daughter_to_parent),
  Parent = unname(daughter_to_parent)
)
```

```{r}
# Group by 'Parent' and collapse 'Daughter' entries
df_grouped <- daughter_to_parent_df %>%
  group_by(Parent) %>%
  summarise(Daughters = paste(Daughter, collapse = ", ")) %>%
  arrange(Parent)   # Arrange alphabetically by Parent

# Print the table
print(df_grouped)
```

```{r}

# Load required libraries
library(dplyr)
library(gridExtra)
library(grid)
library(gtable)

# Group by 'Parent' and collapse 'Daughter' entries
df_grouped <- daughter_to_parent_df %>%
  group_by(Parent) %>%
  summarise(Daughters = paste(Daughter, collapse = ", ")) %>%
  arrange(Parent)

# Open a PNG device with larger dimensions
png("parent_daughter_table.png", 
    width = 12000,      
    height = 5000,     
    res = 500)        

# Create a custom theme with larger, bolder text and text wrapping
custom_theme <- ttheme_default(
  core = list(
    fg_params = list(fontsize = 14, fontface = "bold"),
    bg_params = list(fill = "white"),
    padding = unit(c(4, 6), "mm")      
  ),
  colhead = list(
    fg_params = list(fontsize = 16, fontface = "bold"),
    bg_params = list(fill = "lightgrey"),
    padding = unit(c(4, 6), "mm")      
  )
)

# Create a table grob with the custom theme
table_grob <- tableGrob(df_grouped, 
                       rows = NULL,
                       theme = custom_theme)

# Adjust column widths
table_grob$widths <- unit(c(0.12, 0.88), "npc")

# Set row heights directly
# Make each row taller by multiplying the default height
table_grob$heights <- unit(rep(1.5, length(table_grob$heights)), "cm")

# Add spacing between rows
table_grob <- gtable_add_row_space(table_grob, unit(0.5, "cm"))

# Add a title
title <- textGrob("", 
                 gp = gpar(fontsize = 18, fontface = "bold"))

# Combine title and table
final_grob <- arrangeGrob(title, table_grob, 
                         ncol = 1, 
                         heights = c(0.05, 0.95))

# Plot the table
grid.draw(final_grob)

# Close the PNG device
dev.off()

# Print the table in the console
print(df_grouped)
```

#### Chemotherapy Celltype classification

```{r}
# Create a new column in the colData of the spe_chemotherapy object for parent cells
colData(spe_chemotherapy)$ParentCell <- daughter_to_parent[spe_chemotherapy$Label]

# Verify the addition of the parent cell types
head(colData(spe_chemotherapy)$ParentCell)
```

```{r}
#  Extract expression matrix and metadata
expr_matrix <- assays(spe_chemotherapy)$exprs
markers <- rownames(spe_chemotherapy)
cell_metadata <- as.data.frame(colData(spe_chemotherapy))  # Ensure colData is a dataframe

```

```{r}

# Grouping the expression data by Label (cell type) and ParentCell
grouped_data <- as.data.frame(expr_matrix) %>%
  t() %>%
  as.data.frame() %>%
  mutate(Label = cell_metadata$Label, ParentCell = cell_metadata$ParentCell) %>%
  group_by(Label, ParentCell) %>%
  summarise_all(mean, na.rm = TRUE) %>%
  pivot_longer(-c(Label, ParentCell), names_to = "Marker", values_to = "MeanExpression")

# Check that Label column exists in grouped_data
print(head(grouped_data))
```

```{r}
 
cell_fraction <- data.frame(
  Label = names(table(cell_metadata$Label)),
  Fraction = as.vector(table(cell_metadata$Label))
) %>%
  mutate(FractionPercent = Fraction / sum(Fraction) * 100)


# Check that Label column exists in cell_fraction
print(head(cell_fraction))

```

```{r}

# Ensure that Label column is a character in both dataframes to avoid mismatch during merge
grouped_data$Label <- as.character(grouped_data$Label)
cell_fraction$Label <- as.character(cell_fraction$Label)

# Merge expression data with cell fraction
plot_data <- merge(grouped_data, cell_fraction, by = "Label")

# Check the structure of the merged data
print(head(plot_data))
```

```{r}

# Proceed with plotting
ggplot(plot_data, aes(x = Label, y = ParentCell)) +
  geom_point(aes(size = FractionPercent, color = MeanExpression)) +
  scale_size_continuous(range = c(1, 10)) +  # Scale the size of points (fraction of cells)
  scale_color_gradient(low = "blue", high = "green") +  # Color scale for mean expression
  theme_minimal() +
  labs(
    x = "Parent Cell Group", y = "Cell Types (Label)",
    size = "Fraction of Cells (%)",
    color = "Mean Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid = element_blank()
  )
```

```{r}

# Proceed with plotting
ggplot(plot_data, aes(x = Label, y = ParentCell)) +
  geom_point(aes(size = FractionPercent, color = FractionPercent)) +  # Map color to FractionPercent
  scale_size_continuous(range = c(1, 10)) +  # Scale the size of points (fraction of cells)
  scale_color_viridis(option = "C", direction = 1) +  # Continuous color scale using viridis
  theme_minimal() +
  labs(
    x = "Cell Types (Label)", y = "Parent Cell Group",  # Clarified axis labels
    size = "Fraction of Cells (%)",
    color = "Fraction Percent",
    title = "Labels to Parent Cells"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, face = "bold"),  # Rotate and bold x-axis points
    axis.text.y = element_text(face = "bold"),  # Bold y-axis points
    axis.title.x = element_text(face = "bold"),  # Make x-axis label bold
    axis.title.y = element_text(face = "bold"),  # Make y-axis label bold
    panel.grid = element_blank()  # Remove panel grid for a cleaner look
  )
```

```{r}
# Create a data frame from the ParentCell vector
parent_cells <- data.frame(ParentCell = spe_chemotherapy$ParentCell)

# Calculate the frequency and percentage of each cell type
cell_counts <- parent_cells %>%
  count(ParentCell) %>%
  mutate(percentage = (n / sum(n) * 100)) %>%
  arrange(desc(percentage))

# Display the results
print(cell_counts)
```

```{r}
 #Create a data frame from the ParentCell vector
parent_cells <- data.frame(ParentCell = spe_chemotherapy$ParentCell)

initial_counts <- parent_cells %>%
  count(ParentCell) %>%
  mutate(original_percentage = (n / sum(n) * 100))


cell_counts <- initial_counts %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10), 
      TRUE ~ original_percentage
    )
  ) %>%
 
  mutate(percentage = (percentage / sum(percentage) * 100)) %>%
  # Arrange in descending order of modified percentage
  arrange(desc(percentage))


#print("Cell type distribution with modified 'Other' percentage:")
#print(select(cell_counts, ParentCell, n, percentage))
```

```         
```

```{r}

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 16, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Chemotherapy Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5),
            color = "black", size = 4, fontface = "bold")


# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 10,
  height = 7,
  dpi = 600,
  bg = "white"
)
print(bar_plot)
```

```{r}

```

```{r}
# Define the color palette for cell types
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color 
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#000000", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 16, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Chemotherapy Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5),
            color = "black", size = 4, fontface = "bold")

# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 10,
  height = 7,
  dpi = 600,
  bg = "white"
)

print(bar_plot)
```

```{r}
# Define the color palette for cell types
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#000000", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Chemotherapy Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  # Place text labels ABOVE the bars instead of inside them
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            vjust = -0.5,  # Negative value places text above the bar
            color = "black", size = 4, fontface = "bold")

# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_Parent_Cell_types_Percentages2.png",
  plot = bar_plot,
  width = 10,
  height = 7,
  dpi = 600,
  bg = "white"
)

print(bar_plot)
```

```{r}

```

```{r}

library(scuttle)
library(ComplexHeatmap)

# Step 1: Define a consistent color palette for cell types
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color from the right legend
  "Effector T Cells" = "#00FFFF",  # Cyan color from the right legend
  "Endothelial Cells" = "#000000", # Black color from the right legend
  "Fibroblasts" = "#FF0000",       # Red color from the right legend
  "Macrophages" = "#00FF00",       # Greenblue color from the right legend
  "Other" = "#e06000",             # Dark orange color from the right legend
  "Regulatory T Cells" = "#FD3DB5" # Magenta color from the right legend
)

# Step 2: Aggregate by cell type
celltype_mean <- aggregateAcrossCells(as(spe_chemotherapy, "SingleCellExperiment"),  
                                      ids = spe_chemotherapy$ParentCell, 
                                      statistics = "mean",
                                      use.assay.type = "exprs", 
                                      subset.row = type_markers)

# Step 3: Ensure cell type metadata is correctly assigned and ordered
unique_cell_types <- unique(spe_chemotherapy$ParentCell)

# Ensure the factor levels match the color order
colData(celltype_mean)$celltype <- factor(colData(celltype_mean)$ParentCell, 
                                          levels = names(cell_type_colors))

# Step 4: Open PNG device
png("cluster_cell_type_heatmap_Chemo.png", 
    width = 3800, 
    height = 3000, 
    res = 500)

# Step 5: Generate the heatmap with boldened and rotated x-axis labels
dittoHeatmap(celltype_mean,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scale = "row",
             heatmap.colors = viridis::viridis(100),
             annot.by = "celltype",
             annotation_colors = list(celltype = cell_type_colors), 
             fontsize_row = 8,
             fontsize_col = 10,  # Increased font size for x-axis labels
             column_names_rot = 45,  # Rotates x-axis labels 45 degrees
             show_colnames = TRUE,
             bold_colnames = TRUE,  # This should bold the column names
             column_font_gp = gpar(fontface = "bold"),  # Additional parameter to bold column labels
             bold_rownames = TRUE,
             breaks = seq(0, 1, length.out = 101),
             show_legend = TRUE,
             legend_position = "right")

# Close PNG device
dev.off()
```

```{r}



```

```{r}

# Extract expression data
expression_data <- assay(celltype_mean, "exprs")
```

```{r}

# Function to identify highly expressed proteins
identify_high_expression <- function(expression_vector, threshold = 0.95) {
  quantile_threshold <- quantile(expression_vector, probs = threshold)
  return(expression_vector > quantile_threshold)
}


# Apply the function to each row (protein)
high_expression_matrix <- t(apply(expression_data, 1, identify_high_expression))

```

```{r}

# Convert to a data frame
high_expression_df <- as.data.frame(high_expression_matrix)

# Add protein names as a column
high_expression_df$Protein <- rownames(high_expression_df)

```

```{r}

# Reshape the data frame
spe_chemotherapy_result_df <- high_expression_df %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "IsHighlyExpressed") %>%
  filter(IsHighlyExpressed) %>%
  select(-IsHighlyExpressed)


# Sort the dataframe
spe_chemotherapy_result_df <- spe_chemotherapy_result_df %>%
  arrange(Protein, CellType)

```

```{r}
spe_chemotherapy_celltype_summary <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(NumProteins = n_distinct(Protein)) %>%
  arrange(desc(NumProteins))
spe_chemotherapy_celltype_summary
```

```{r}

# Create a new dataframe with cell types and their highly expressed proteins
spe_chemotherapy_celltype_proteins_df <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

spe_chemotherapy_celltype_proteins_df
```

```{r}
# Create a new dataframe with cell types and their highly expressed proteins
spe_chemotherapy_celltype_proteins_df <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

# Open a PNG device with a specific resolution and size
png("cell_type_proteins_table_for_chemo.png", 
    width = 15, 
    height = 3, 
    units = "in", 
    res = 360)

# Create a table grob (graphical object)
table_grob <- tableGrob(spe_chemotherapy_celltype_proteins_df, 
                        rows = NULL,  # Remove row numbers
                        theme = ttheme_default(
                          core = list(bg_colors = "white"),
                          colhead = list(bg_colors = "lightgrey")
                        ))

# Add a title to the table
title <- textGrob("Cell Type Highly Expressed Proteins", 
                  gp = gpar(fontsize = 14, fontface = "bold"))

# Combine title and table
final_grob <- arrangeGrob(title, table_grob, 
                          ncol = 1, 
                          heights = c(0.1, 0.9))

# Plot the table
grid.draw(final_grob)

# Close the PNG device
dev.off()

# Print the dataframe in the console
print(spe_chemotherapy_celltype_proteins_df)
```

```{r}

library(ggplot2)
library(dplyr)

# Create a new dataframe with cell types and their highly expressed proteins count
spe_chemotherapy_celltype_proteins_df <- spe_chemotherapy_result_df %>%
  group_by(CellType) %>%
  summarise(ProteinCount = n()) %>%
  arrange(desc(ProteinCount))

# Open a PNG device with a specific resolution and size
png("cell_type_proteins_barplot.png", 
    width = 10, 
    height = 6, 
    units = "in", 
    res = 300)

# Create a bar plot
ggplot(spe_chemotherapy_celltype_proteins_df, aes(x = reorder(CellType, -ProteinCount), y = ProteinCount, fill = CellType)) +
  geom_bar(stat = "identity") +
  labs(title = "Highly Expressed Proteins per Cell Type", 
       x = "Cell Type", 
       y = "Number of Highly Expressed Proteins") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set3")

# Close the PNG device
dev.off()

# Print the dataframe in the console
print(spe_chemotherapy_celltype_proteins_df)

```

```         
```

```{r}
# Create a random sample of 20% of cells
sample_size <- round(ncol(spe_chemotherapy) * 0.1)
sampled_cells <- sample(colnames(spe_chemotherapy), sample_size)
```

```{r}

# Subset the SPE object
spe_chemotherapy_subset <- spe_chemotherapy[, sampled_cells]
```

```{r}

# Generate the plot using the color column in colData
p5 <- dittoDimPlot(
 spe_chemotherapy_subset, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = FALSE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cluster_cell_types_umap_for_chemo.png", 
  plot = p5,
  width = 3600,    # Width in inches
  height = 2400,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)


# Display the Plot
print(p5)
```

```{r}

```

```{r}
# Clustering function with specified color palette
cluster_by_parent_cell_type <- function(spe, reduction_method = "Harmony") {
  # Define the color palette for cell types
  cell_type_colors <- c(
    "Cancer Cells" = "#0000FF",      # Blue color 
    "Effector T Cells" = "#00FFFF",  # Cyan color
    "Endothelial Cells" = "#000000", # Black color
    "Fibroblasts" = "#FF0000",       # Red color
    "Macrophages" = "#00FF00",       # Green color
    "Other" = "#e06000",             # Dark orange color
    "Regulatory T Cells" = "#FD3DB5" # Magenta color
  )
  
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each ParentCell type
  cell_type <- unique(colData(spe)$ParentCell)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(cell_type)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$ParentCell == cell_type[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(cell_type))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[cell_type[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    ParentCell = rep(names(phase_coordinates), 
                     sapply(phase_coordinates, nrow))
  )
  
  # Generate plot with specified color palette
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = ParentCell)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy Cluster by Parent Cell Type",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_manual(
      values = cell_type_colors[names(cell_type_colors) %in% unique(umap_df$ParentCell)], 
      name = "Parent Cell Type"
    )
  
  return(p)
}
```

```{r}
clustered_umap_plot_parent_celltype <- cluster_by_parent_cell_type(spe_chemotherapy_subset)
print(clustered_umap_plot_parent_celltype)


# Save the plot
ggsave("Chemotherapy_Umap_Cluster_by_Parent_Cell_Type.png",
       clustered_umap_plot_parent_celltype,
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )

```

```{r}


```

```{r}

```

```{r}

```

```{r}

```

```         
```

```{r}

```

#### Chemo_Immu Celltype classification

```{r}

# Create a new column in the colData of the spe_chemo_immu object for parent cells
colData(spe_chemo_Immu)$ParentCell <- daughter_to_parent[spe_chemo_Immu$Label]

# Verify the addition of the parent cell types
head(colData(spe_chemo_Immu)$ParentCell)
```

```         
```

```{r}
# Step 1: Define a consistent color palette for cell types
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color from the right legend
  "Effector T Cells" = "#00FFFF",  # Cyan color from the right legend
  "Endothelial Cells" = "#000000", # Black color from the right legend
  "Fibroblasts" = "#FF0000",       # Red color from the right legend
  "Macrophages" = "#00FF00",       # Greenblue color from the right legend
  "Other" = "#e06000",             # Dark orange color from the right legend
  "Regulatory T Cells" = "#FD3DB5" # Magenta color from the right legend
)

## aggregate by cell type
celltype_mean1 <- aggregateAcrossCells(as(spe_chemo_Immu, "SingleCellExperiment"), ids = spe_chemo_Immu$ParentCell, 
                     statistics = "mean",
                     use.assay.type = "exprs", 
                     subset.row = type_markers)



# Step 3: Ensure cell type metadata is correctly assigned and ordered
unique_cell_types <- unique( spe_chemo_Immu$ParentCell)


# Step 4: Count the number of cells per aggregated cell type
# Ensure the factor levels match the color order
colData(celltype_mean1)$celltype <- factor(colData(celltype_mean1)$ParentCell, 
                                          levels = names(cell_type_colors))


# Open PNG device with high resolution
png("cluster_cell_type_heatmap_Chemo_immu.png", 
    width = 3800,     # Wide width in pixels
    height = 3000,    # Tall height in pixels 
    res = 500,        # High resolution (300 dpi)
    pointsize = 12)   # Slightly larger point size for readability


# Step 5: Plot the heatmap using dittoHeatmap
dittoHeatmap(celltype_mean1,
             assay = "exprs", 
             cluster_cols = TRUE, 
             scale = "row",
             heatmap.colors = viridis::viridis(100),
             annot.by = "celltype",
             annotation_colors = list(celltype = cell_type_colors), 
             fontsize_row = 8,
             fontsize_col = 10,  # Increased font size for x-axis labels
             column_names_rot = 45,  # Rotates x-axis labels 45 degrees
             show_colnames = TRUE,
             bold_colnames = TRUE,  # This should bold the column names
             column_font_gp = gpar(fontface = "bold"),  # Additional parameter to bold column labels
             bold_rownames = TRUE,
             breaks = seq(0, 1, length.out = 101),
             show_legend = TRUE,
             legend_position = "right")

# Close PNG device
dev.off()
```

```{r}



```

```{r}
#Create a data frame from the ParentCell vector
parent_cells <- data.frame(ParentCell = spe_chemo_Immu$ParentCell)

initial_counts <- parent_cells %>%
  count(ParentCell) %>%
  mutate(original_percentage = (n / sum(n) * 100))


cell_counts <- initial_counts %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10), 
      TRUE ~ original_percentage
    )
  ) %>%
 
  mutate(percentage = (percentage / sum(percentage) * 100)) %>%
  # Arrange in descending order of modified percentage
  arrange(desc(percentage))
```

```{r}

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Chemotherapy_&_Immunotherapy Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5),
            color = "black", size = 4, fontface = "bold")


# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_&_Immunotherapy_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 10,
  height = 7,
  dpi = 600,
  bg = "white"
)
print(bar_plot)
```

```{r}
# Define the color palette for cell types
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color 
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#000000", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 16, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Chemotherapy Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
 # Place text labels ABOVE the bars instead of inside them
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            vjust = -0.5,  # Negative value places text above the bar
            color = "black", size = 4, fontface = "bold")


# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_Immunotherapy_Parent_Cell_types_Percentages2.png",
  plot = bar_plot,
  width = 10,
  height = 7,
  dpi = 600,
  bg = "white"
)

print(bar_plot)
```

```{r}

```

```{r}

```

```{r}

# Extract expression data
expression_data1 <- assay(celltype_mean1, "exprs")
```

```{r}

# Function to identify highly expressed proteins
identify_high_expression1 <- function(expression_vector, threshold = 0.95) {
  quantile_threshold <- quantile(expression_vector, probs = threshold)
  return(expression_vector > quantile_threshold)
}


# Apply the function to each row (protein)
high_expression_matrix1 <- t(apply(expression_data1, 1, identify_high_expression1))

```

```{r}

# Convert to a data frame
high_expression_df1 <- as.data.frame(high_expression_matrix1)

# Add protein names as a column
high_expression_df1$Protein <- rownames(high_expression_df1)

```

```{r}

# Reshape the data frame
spe_chemo_immu_result_df <- high_expression_df1 %>%
  pivot_longer(cols = -Protein, names_to = "CellType", values_to = "IsHighlyExpressed") %>%
  filter(IsHighlyExpressed) %>%
  select(-IsHighlyExpressed)


# Sort the dataframe
spe_chemo_immu_result_df <- spe_chemo_immu_result_df %>%
  arrange(Protein, CellType)

```

```{r}

spe_chemo_immu_celltype_summary <- spe_chemo_immu_result_df %>%
  group_by(CellType) %>%
  summarise(NumProteins = n_distinct(Protein)) %>%
  arrange(desc(NumProteins))
spe_chemo_immu_celltype_summary 
```

```{r}

# Create a new dataframe with cell types and their highly expressed proteins
spe_chemo_immu_celltype_proteins_df <- spe_chemo_immu_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

spe_chemo_immu_celltype_proteins_df
```

```{r}
# Create a new dataframe with cell types and their highly expressed proteins
spe_chemo_immu_celltype_proteins_df <- spe_chemo_immu_result_df %>%
  group_by(CellType) %>%
  summarise(HighlyExpressedProteins = paste(Protein, collapse = ", ")) %>%
  arrange(CellType)

# Open a PNG device with a specific resolution and size
png("cell_type_proteins_table_for_chemo_immu.png", 
    width = 15, 
    height = 3, 
    units = "in", 
    res = 360)

# Create a table grob (graphical object)
table_grob <- tableGrob(spe_chemo_immu_celltype_proteins_df, 
                        rows = NULL,  # Remove row numbers
                        theme = ttheme_default(
                          core = list(bg_colors = "white"),
                          colhead = list(bg_colors = "lightgrey")
                        ))

# Add a title to the table
title <- textGrob("Cell Type Highly Expressed Proteins", 
                  gp = gpar(fontsize = 14, fontface = "bold"))

# Combine title and table
final_grob <- arrangeGrob(title, table_grob, 
                          ncol = 1, 
                          heights = c(0.1, 0.9))

# Plot the table
grid.draw(final_grob)

# Close the PNG device
dev.off()

# Print the dataframe in the console
print(spe_chemo_immu_celltype_proteins_df)

```

```{r}

```

```{r}
# Generate the plot using the color column in colData
p6 <- dittoDimPlot(
  spe_chemo_Immu, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
)  +
  theme(
    legend.title = element_blank(),
    # Bold and larger text for axes and title
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 10),
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5)
  ) +
  ggtitle("")

# Save the plot as a high-resolution PNG
ggsave(
  "cluster_cell_types_umap_for_chemo_Immu.png", 
  plot = p6,
  width = 3600,    # Width in inches
  height = 2400,    # Height in inches
  dpi = 500,     # High resolution
  units = "px"   # Specify units as inches
)


# Display the Plot
print(p6)

```

```{r}

```

```{r}

# Create a random sample of 20% of cells
sample_size4 <- round(ncol(spe_chemo_Immu) * 0.1)
sampled_cells4 <- sample(colnames(spe_chemo_Immu), sample_size4)
```

```{r}


# Subset the SPE object
spe_chemo_immune_subset2 <- spe_chemo_Immu[, sampled_cells4]
```

```{r}


# Clustering function with specified color palette
cluster_by_parent_cell_type1 <- function(spe, reduction_method = "Harmony") {
  # Define the color palette for cell types
  cell_type_colors <- c(
    "Cancer Cells" = "#0000FF",      # Blue color 
    "Effector T Cells" = "#00FFFF",  # Cyan color
    "Endothelial Cells" = "#000000", # Black color
    "Fibroblasts" = "#FF0000",       # Red color
    "Macrophages" = "#00FF00",       # Green color
    "Other" = "#e06000",             # Dark orange color
    "Regulatory T Cells" = "#FD3DB5" # Magenta color
  )
  
  # Error checking
  if(!reduction_method %in% reducedDimNames(spe)) {
    stop(paste("Reduction method", reduction_method, "not found in the dataset"))
  }
  
  # Extract reduced dimensions
  rd <- reducedDim(spe, reduction_method)
  
  # Perform clustering for each ParentCell type
  cell_type <- unique(colData(spe)$ParentCell)
  
  # Initialize empty list to store phase-specific coordinates
  phase_coordinates <- list()
  
  # Set seed for reproducibility
  set.seed(42)
  
  # Generate distinct clusters for each phase
  for (i in seq_along(cell_type)) {
    # Select subset for current phase
    phase_subset <- rd[colData(spe)$ParentCell == cell_type[i], ]
    
    # Perform UMAP on phase subset
    phase_umap <- uwot::umap(phase_subset, 
                              n_neighbors = 15, 
                              min_dist = 0.5,
                              spread = 1)
    
    # Adjust coordinates to spread phases
    angle <- (i-1) * (2*pi / length(cell_type))
    radius <- 5  # Separation distance
    
    phase_umap[,1] <- phase_umap[,1] + radius * cos(angle)
    phase_umap[,2] <- phase_umap[,2] + radius * sin(angle)
    
    phase_coordinates[[cell_type[i]]] <- phase_umap
  }
  
  # Combine all coordinates
  all_umap_coords <- do.call(rbind, phase_coordinates)
  
  # Create data frame for plotting
  umap_df <- data.frame(
    UMAP1 = all_umap_coords[,1],
    UMAP2 = all_umap_coords[,2],
    ParentCell = rep(names(phase_coordinates), 
                     sapply(phase_coordinates, nrow))
  )
  
  # Generate plot with specified color palette
  p <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = ParentCell)) +
    geom_point(alpha = 0.7, size = 2) +
    theme_minimal() +
    labs(
      title = "Chemotherapy_&_Immunotherapy Cluster by Parent Cell Type",
      x = "UMAP 1",
      y = "UMAP 2"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14),
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 12),
      legend.position = "right"
    ) +
    scale_color_manual(
      values = cell_type_colors[names(cell_type_colors) %in% unique(umap_df$ParentCell)], 
      name = "Parent Cell Type"
    )
  
  return(p)
}
```

```{r}

clustered_umap_plot_parent_celltype1 <- cluster_by_parent_cell_type1(spe_chemo_immune_subset2)
print(clustered_umap_plot_parent_celltype1)

# Save the plot
ggsave("Chemotherapy_&_Immunotherapy_Umap_Cluster_by_Parent_Cell_Type.png",
       clustered_umap_plot_parent_celltype1,
       width = 3600,    # Width in inches
      height = 2400,    # Height in inches
      dpi = 500,     # High resolution
      units = "px" # Specify units as inches
     )


```

```{r}

```

```{r}

```

#### Getting the data for patient with ID NT004 to get parameters for the model using the chemotherapy only dataset.

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy_patient_1 <- spe_chemotherapy[, colData(spe_chemotherapy)$PatientID == "NT004"]

spe_chemotherapy_patient_1
```

```{r}

# Generate the plot using the color column in colData
dittoDimPlot(
  spe_chemotherapy_patient_1, 
  var = "ParentCell", 
  reduction.use = "UMAP_mnnCorrected", 
  size = 0.2,
  do.label = TRUE,
  labels.size = 2.5,
  do.raster = TRUE,
  legend.show = TRUE
) +
  theme(legend.title = element_blank()) +
  ggtitle("Patient NT004 Cell types on UMAP for spe_chemotherapy")
```

```{r}

# Function to count cells by type and biopsyphase
count_cells_by_type_and_phase <- function(spe) {
  # Extract cell type and biopsyphase information from colData
  cell_data <- as.data.frame(colData(spe))
  
  # Ensure 'ParentCell' and 'BiopsyPhase' columns exist
  if (!all(c("ParentCell", "BiopsyPhase") %in% colnames(cell_data))) {
    stop("Both 'ParentCell' and 'BiopsyPhase' columns must be present in colData")
  }
  
  # Count cells
  cell_counts <- cell_data %>%
    group_by(ParentCell, BiopsyPhase) %>%
    summarise(count = n(), .groups = "drop") %>%
    pivot_wider(names_from = BiopsyPhase, values_from = count, values_fill = 0)
  
  # Calculate total for each cell type
  cell_counts$Total <- rowSums(cell_counts[, -1])
  
  # Reorder columns to have Total as the second column
  cell_counts <- cell_counts %>%
    select(ParentCell, Total, everything())
  
  return(cell_counts)
}

```

```{r}

# Main analysis script
main_analysis <- function(spe) {
  # Perform cell count analysis
  result <- count_cells_by_type_and_phase(spe)
  
  # Print the result
  print(result)
  
  # Optionally, you can save the result to a CSV file
  # write.csv(result, "cell_count_analysis_result.csv", row.names = FALSE)
  
  return(result)
}
```

```{r}

result4 <- main_analysis(spe_chemotherapy_patient_1)
result4
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### Getting the mean expression values of some of the protein biomarkers in the various cell types for spe_chemotherapy.

```{r}

extract_mean_expression_tidy <- function(spe, 
                                         cell_type_column = "ParentCell", 
                                         assay_name = "exprs") {
  # Check if the cell type column exists in colData
  if (!cell_type_column %in% colnames(colData(spe))) {
    stop(paste("Column", cell_type_column, "not found in colData"))
  }
  
  # Check if the specified assay exists
  if (!assay_name %in% assayNames(spe)) {
    stop(paste("Assay", assay_name, "not found in the SpatialExperiment object"))
  }
  
  # Extract expression data
  expression_data <- assay(spe, assay_name)
  
  # Extract cell type information
  cell_types <- colData(spe)$ParentCell
  
  # Calculate mean expression for each biomarker and cell type
  mean_expression <- t(apply(expression_data, 1, function(x) tapply(x, cell_types, mean, na.rm = TRUE)))
  
  # Convert to a tidy data frame
  mean_expression_df <- as.data.frame(mean_expression) %>%
    rownames_to_column("Biomarker") %>%
    pivot_longer(cols = -Biomarker, names_to = "CellType", values_to = "MeanExpression")
  
  # Sort the data frame by CellType
  mean_expression_df_sorted <- mean_expression_df %>%
    arrange(CellType, Biomarker)
  
  return(mean_expression_df_sorted)
}
```

```{r}


results2 <- extract_mean_expression_tidy(spe_chemotherapy_patient_1, cell_type_column = "ParentCell",assay_name = "exprs")

results2 
```

```{r}
biomarkers_of_interest <- c("CD31", "CD8", "CD163", "CD3", "PD-L1 (SP142)", "OX40","CD11c", "TOX", "PD-1", "IDO","ICOS","PD-L1 (73-10)","FOXP3","CD4", "PDGFRB", "PDPN", "Vimentin", "Caveolin-1", "HLA-DR"  )

```

```{r}
# Extract mean expression for selected biomarkers
result3 = results2[results2$Biomarker  %in% biomarkers_of_interest, ]
result3
```

```{r}


```

```{r}
sort(results2$mean_expression)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### Save datasets

```{r}

#saveRDS(spe_chemotherapy, "saved_data/spe_chemotherapy_phenotyping_spatial.rds" )

```

```{r}
#saveRDS(spe_chemo_Immu, "saved_data/spe_chemo_Immu_phenotyping_spatial.rds" )
```

```{r}

```

```{r}

```

#### Spatial Analysis for both chemo and chemo_immu

```{r}
## Importing the datasets: Chemotherapy

spe_chemo <- readRDS("saved_data/spe_chemotherapy_Celltype_phenotyping_spatial.rds")
spe_chemo
```

```{r}

## Importing the datasets: Chemotherapy

spe_chemo_immu <- readRDS("saved_data/spe_chemo_Immu_Celltype_phenotyping_spatial.rds")
spe_chemo_immu
```

#### Measuring biomarkers expression similariteis for spe_chemo

```{r}

# Get the expression matrix (cells x proteins)
expr_matrix_spe_chemo <- assay(spe_chemo, 'exprs')

```

```{r}
# Compute similarity matrix (correlation between cells based on expression)
similarity_matrix <- cor(t(expr_matrix_spe_chemo), method = "pearson")  # Can also use "spearman"
```

```{r}
similarity_matrix2 <- cor(t(expr_matrix_spe_chemo), method = "spearman")  #  using "spearman"
```

```{r}

# Create a heatmap of the similarity matrix
heatmap(similarity_matrix, main = "Protein Expression Similarity Between Cells in spe_chemo")
```

```{r}

 #Create a heatmap of the similarity matrix
heatmap(similarity_matrix2, main = "Protein Expression Similarity Between Cells in spe_chemo")
```

```{r}

```

#### Measuring biomarkers expression similariteis for spe_chemo_immu

```{r}

# Get the expression matrix (cells x proteins)
expr_matrix_spe_chemo_immu <- assay(spe_chemo_immu, 'exprs')

```

```{r}

# Compute similarity matrix (correlation between cells based on expression)
similarity_matrix3 <- cor(t(expr_matrix_spe_chemo_immu), method = "pearson")  # Can also use "spearman"
```

```{r}

similarity_matrix4 <- cor(t(expr_matrix_spe_chemo_immu), method = "spearman")  #  using "spearman"
```

```{r}

# Create a heatmap of the similarity matrix
heatmap(similarity_matrix3, main = "Protein Expression Similarity Between Cells in spe_chemo_immu")
```

```{r}

#Create a heatmap of the similarity matrix
heatmap(similarity_matrix4, main = "Protein Expression Similarity Between Cells in spe_chemo_immu")
```

```{r}

```

#### Measuring spatial proximity interactions with spe_chemo_data

```{r}

# Updated function to build spatial graph with error handling
build_spatial_graph <- function(spe_chemo, k = 5) {
  # Extract spatial coordinates and ImageID
  coords <- spatialCoords(spe_chemo)
  image_ids <- spe_chemo$ImageID
  
  # Initialize progress bar
  pb <- progress_bar$new(
    format = "[:bar] :percent ETA: :eta",
    total = length(unique(image_ids)),
    clear = FALSE,
    width = 60
  )
  
  # Function to compute k-NN for a single image
  compute_knn_for_image <- function(img_id) {
    img_indices <- which(image_ids == img_id)
    img_coords <- coords[img_indices, ]
    
    if (nrow(img_coords) <= k) {
      # If there are fewer points than k, connect all points to each other
      knn_result <- list(
        nn.idx = matrix(rep(1:nrow(img_coords), each = nrow(img_coords)), ncol = nrow(img_coords)),
        nn.dists = as.matrix(dist(img_coords))
      )
    } else {
      knn_result <- RANN::nn2(img_coords, k = k + 1)
    }
    
    pb$tick()
    return(list(indices = img_indices, knn = knn_result))
  }
  
  # Compute k-NN for each image
  knn_results <- lapply(unique(image_ids), compute_knn_for_image)
  
  # Create edge list
  edges <- do.call(rbind, lapply(knn_results, function(res) {
    idx <- res$indices[as.vector(res$knn$nn.idx[, -1])]
    cbind(rep(res$indices, each = k), idx)
  }))
  
  # Create igraph object
  g <- graph_from_edgelist(edges, directed = FALSE)
  
  # Set edge weights with error handling
  tryCatch({
    weights <- unlist(lapply(knn_results, function(res) as.vector(res$knn$nn.dists[, -1])))
    if (any(!is.numeric(weights))) {
      stop("Non-numeric values found in distance calculations")
    }
    if (any(weights == 0)) {
      warning("Zero distances found. Setting weights to a small non-zero value.")
      weights[weights == 0] <- 1e-10
    }
    E(g)$weight <- 1 / weights
  }, error = function(e) {
    cat("Error in weight calculation:", conditionMessage(e), "\n")
    cat("Debugging information:\n")
    print(summary(unlist(lapply(knn_results, function(res) as.vector(res$knn$nn.dists[, -1])))))
    stop("Weight calculation failed")
  })
  
  # Set node attributes
  V(g)$celltype <- spe_chemo$ParentCell
  V(g)$x <- coords[, 1]
  V(g)$y <- coords[, 2]
  V(g)$ImageID <- image_ids
  
  return(g)
}
```

```{r}

visualize_spatial_graph <- function(graph, spe_chemo, image_id, node_color_by = "celltype") {
  # Get the row indices for the specified ImageID
  nodes_to_keep <- which(spe_chemo$ImageID == image_id)
  
  # Since vertices don't have names, we'll use numeric indices directly
  subgraph <- induced_subgraph(graph, as.numeric(nodes_to_keep))
  
  # Print graph information for debugging
  cat("Full graph - Nodes:", vcount(graph), "Edges:", ecount(graph), "\n")
  cat("Subgraph - Nodes:", vcount(subgraph), "Edges:", ecount(subgraph), "\n")
  
  if (vcount(subgraph) == 0) {
    warning("No nodes found for the specified ImageID.")
    return(NULL)
  }
  
  node_attrs <- data.frame(
    celltype = V(subgraph)$celltype,
    x = V(subgraph)$x,
    y = V(subgraph)$y
  )
  
  layout <- as.matrix(node_attrs[, c("x", "y")])
  
  p <- ggraph(subgraph, layout = layout) +
    geom_edge_link(edge_colour = "gray", edge_alpha = 0.2, show.legend = FALSE) +
    geom_node_point(aes(color = .data[[node_color_by]]), size = 1, alpha = 0.7) +
    scale_color_discrete() +
    theme_void() +
    labs(title = paste("ImageID:", image_id),
         color = node_color_by) +
    coord_fixed()
  
  return(p)
}
```

```{r}

# Usage example
spatial_graph <- build_spatial_graph(spe_chemo, k = 5)

```

```{r}

# Save the graph as metadata in the SPE object
metadata(spe_chemo)$spatial_graph <- spatial_graph

```

#### Single ImageID visualization

```{r}

# Choose a specific ImageID to visualize
image_id_to_visualize <- "NTImg0010"  # Replace with your desired ImageID

# Visualize the graph for the specific ImageID (using celltype for node color)
plot <- visualize_spatial_graph(spatial_graph, spe_chemo, image_id = image_id_to_visualize, node_color_by = "celltype")


# Display the plot
print(plot)
```

```{r}

```

### Multi-plot for ImageID's

```{r}

visualize_multiple_spatial_graphs <- function(graph, spe_chemo, image_ids, node_color_by = "celltype", ncol = 2) {
  # Check if image_ids is provided
  if (missing(image_ids) || length(image_ids) == 0) {
    stop("Please provide at least one ImageID to visualize.")
  }

  # Create a list to store individual plots
  plot_list <- list()

  # Generate plots for each ImageID
  for (img_id in image_ids) {
    plot <- visualize_spatial_graph(graph, spe_chemo, image_id = img_id, node_color_by = node_color_by)
    if (!is.null(plot)) {
      plot_list[[img_id]] <- plot
    }
  }

  # Check if any plots were generated
  if (length(plot_list) == 0) {
    stop("No valid plots were generated. Check your ImageIDs.")
  }

  # Combine plots using patchwork
  combined_plot <- wrap_plots(plot_list, ncol = ncol)

  # Add a title to the combined plot
  final_plot <- combined_plot +
    plot_annotation(
      theme = theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
                    plot.subtitle = element_text(hjust = 0.5, size = 12))
    )

  return(final_plot)
}
```

```{r}

# Usage example:
image_ids_to_visualize <- c("NTImg0001", "NTImg0002", "NTImg0003", "NTImg0004")
multiplot <- visualize_multiple_spatial_graphs(spatial_graph, spe_chemo, image_ids = image_ids_to_visualize, node_color_by = "celltype", ncol = 2)
print(multiplot)
```

```{r}

```

### Single PatientID visualization

```{r}

visualize_patient_spatial_graph <- function(graph, spe_chemo, patient_id, node_color_by = "celltype", edge_alpha = 0.1, node_size = 1, node_alpha = 0.7, max_edges = 10000) {
  # Check if PatientID column exists in spe
  if (!"PatientID" %in% colnames(colData(spe_chemo))) {
    stop("PatientID column not found in the SpatialExperiment object.")
  }
  
  # Filter the graph for the specified PatientID
  nodes_to_keep <- which(spe_chemo$PatientID == patient_id)
  subgraph <- induced_subgraph(graph, nodes_to_keep)
  
  # Remove isolated nodes
  subgraph <- delete_vertices(subgraph, degree(subgraph) == 0)
  
  # Print graph information
  cat("Full graph - Nodes:", vcount(graph), "Edges:", ecount(graph), "\n")
  cat("Patient subgraph - Nodes:", vcount(subgraph), "Edges:", ecount(subgraph), "\n")
  
  # Check if the subgraph is empty
  if (vcount(subgraph) == 0) {
    warning("No nodes found for the specified PatientID.")
    return(NULL)
  }
  
  # Subsample edges if there are too many
  if (ecount(subgraph) > max_edges) {
    set.seed(42)  # for reproducibility
    edges_to_keep <- sample(1:ecount(subgraph), max_edges)
    edge_list <- as_edgelist(subgraph)[edges_to_keep, ]
    subgraph <- graph_from_edgelist(edge_list, directed = FALSE)
    
    # Reapply vertex attributes
    vertex_attr_names <- list.vertex.attributes(graph)
    for (attr in vertex_attr_names) {
      vertex_attr(subgraph, attr) <- vertex_attr(graph, attr)[V(subgraph)]
    }
    
    cat("Subsampled edges - New edge count:", ecount(subgraph), "\n")
  }
  
  # Get node attributes
  node_attrs <- data.frame(
    celltype = V(subgraph)$celltype,
    x = V(subgraph)$x,
    y = V(subgraph)$y,
    ImageID = V(subgraph)$ImageID
  )
  
  # Create a layout based on spatial coordinates
  layout <- as.matrix(node_attrs[, c("x", "y")])
  
  # Create ggraph plot with nodes colored by the specified attribute
  p <- ggraph(subgraph, layout = layout) +
    geom_edge_link(edge_colour = "gray", edge_alpha = edge_alpha, show.legend = FALSE) +
    geom_node_point(aes(color = .data[[node_color_by]]), size = node_size, alpha = node_alpha) +
    scale_color_discrete() +
    theme_void() +
    labs(title = paste("PatientID:", patient_id),
         subtitle = paste("Number of Images:", length(unique(node_attrs$ImageID))),
         color = node_color_by) +
    coord_fixed()  # Ensure the aspect ratio is 1:1
  
  # Add faceting by ImageID if there are multiple images
  if (length(unique(node_attrs$ImageID)) > 1) {
    p <- p + facet_wrap(~ImageID)
  }
  
  return(p)
}
```

```{r}

# Usage example:
patient_id_to_visualize <- "NT008"  # Replace with your desired PatientID
plot <- visualize_patient_spatial_graph(spatial_graph, spe_chemo, patient_id = patient_id_to_visualize, node_color_by = "celltype")
print(plot)
```

```{r}

```

```{r}

```

#### Measuring spatial proximity interactions with spe_chemo_immu_data

```{r}

# Updated function to build spatial graph with error handling
build_spatial_graph1 <- function(spe_chemo_immu, k = 5) {
  # Extract spatial coordinates and ImageID
  coords <- spatialCoords(spe_chemo_immu)
  image_ids <- spe_chemo_immu$ImageID
  
  # Initialize progress bar
  pb <- progress_bar$new(
    format = "[:bar] :percent ETA: :eta",
    total = length(unique(image_ids)),
    clear = FALSE,
    width = 60
  )
  
  # Function to compute k-NN for a single image
  compute_knn_for_image <- function(img_id) {
    img_indices <- which(image_ids == img_id)
    img_coords <- coords[img_indices, ]
    
    if (nrow(img_coords) <= k) {
      # If there are fewer points than k, connect all points to each other
      knn_result <- list(
        nn.idx = matrix(rep(1:nrow(img_coords), each = nrow(img_coords)), ncol = nrow(img_coords)),
        nn.dists = as.matrix(dist(img_coords))
      )
    } else {
      knn_result <- RANN::nn2(img_coords, k = k + 1)
    }
    
    pb$tick()
    return(list(indices = img_indices, knn = knn_result))
  }
  
  # Compute k-NN for each image
  knn_results <- lapply(unique(image_ids), compute_knn_for_image)
  
  # Create edge list
  edges <- do.call(rbind, lapply(knn_results, function(res) {
    idx <- res$indices[as.vector(res$knn$nn.idx[, -1])]
    cbind(rep(res$indices, each = k), idx)
  }))
  
  # Create igraph object
  g <- graph_from_edgelist(edges, directed = FALSE)
  
  # Set edge weights with error handling
  tryCatch({
    weights <- unlist(lapply(knn_results, function(res) as.vector(res$knn$nn.dists[, -1])))
    if (any(!is.numeric(weights))) {
      stop("Non-numeric values found in distance calculations")
    }
    if (any(weights == 0)) {
      warning("Zero distances found. Setting weights to a small non-zero value.")
      weights[weights == 0] <- 1e-10
    }
    E(g)$weight <- 1 / weights
  }, error = function(e) {
    cat("Error in weight calculation:", conditionMessage(e), "\n")
    cat("Debugging information:\n")
    print(summary(unlist(lapply(knn_results, function(res) as.vector(res$knn$nn.dists[, -1])))))
    stop("Weight calculation failed")
  })
  
  # Set node attributes
  V(g)$celltype <- spe_chemo_immu$ParentCell
  V(g)$x <- coords[, 1]
  V(g)$y <- coords[, 2]
  V(g)$ImageID <- image_ids
  
  return(g)
}

```

```{r}

# Visualization function using ggraph
visualize_spatial_graph1 <- function(graph, spe_chemo_immu, image_id, node_color_by = "celltype") {
  # Filter the graph for the specified ImageID
  nodes_to_keep <- which(spe_chemo_immu$ImageID == image_id)
  subgraph <- induced_subgraph(graph, nodes_to_keep)
  
  # Remove isolated nodes
  subgraph <- delete_vertices(subgraph, degree(subgraph) == 0)
  
  # Print graph information
  cat("Full graph - Nodes:", vcount(graph), "Edges:", ecount(graph), "\n")
  cat("Subgraph - Nodes:", vcount(subgraph), "Edges:", ecount(subgraph), "\n")
  
  # Check if the subgraph is empty
  if (vcount(subgraph) == 0) {
    warning("No nodes found for the specified ImageID.")
    return(NULL)
  }
  
  # Get node attributes
  node_attrs <- data.frame(
    celltype = V(subgraph)$celltype,
    x = V(subgraph)$x,
    y = V(subgraph)$y
  )
  
  # Create a layout based on spatial coordinates
  layout <- as.matrix(node_attrs[, c("x", "y")])
  
  # Create ggraph plot with nodes colored by the specified attribute
  p <- ggraph(subgraph, layout = layout) +
    geom_edge_link(edge_colour = "gray", edge_alpha = 0.2, show.legend = FALSE) +
    geom_node_point(aes(color = .data[[node_color_by]]), size = 1, alpha = 0.7) +
    scale_color_discrete() +
    theme_void() +
    labs(title = paste("ImageID:", image_id),
         color = node_color_by) +
    coord_fixed()  # Ensure the aspect ratio is 1:1
  
  return(p)
}

```

```{r}

# Usage example
spatial_graph1 <- build_spatial_graph1(spe_chemo_immu, k = 5)

```

```{r}
# Save the graph as metadata in the SPE object
metadata(spe_chemo_immu)$spatial_graph <- spatial_graph1

```

```{r}

# Choose a specific ImageID to visualize
image_id_to_visualize <- "NTImg0010"  # Replace with your desired ImageID

# Visualize the graph for the specific ImageID (using celltype for node color)
plot <- visualize_spatial_graph1(spatial_graph1, spe_chemo_immu, image_id = image_id_to_visualize, node_color_by = "celltype")


# Display the plot
print(plot)
```

```{r}

```

### Multi-plot for ImageID's

```{r}

visualize_multiple_spatial_graphs1 <- function(graph, spe_chemo_immu, image_ids, node_color_by = "celltype", ncol = 2) {
  # Check if image_ids is provided
  if (missing(image_ids) || length(image_ids) == 0) {
    stop("Please provide at least one ImageID to visualize.")
  }

  # Create a list to store individual plots
  plot_list <- list()

  # Generate plots for each ImageID
  for (img_id in image_ids) {
    plot <- visualize_spatial_graph(graph, spe_chemo_immu, image_id = img_id, node_color_by = node_color_by)
    if (!is.null(plot)) {
      plot_list[[img_id]] <- plot
    }
  }

  # Check if any plots were generated
  if (length(plot_list) == 0) {
    stop("No valid plots were generated. Check your ImageIDs.")
  }

  # Combine plots using patchwork
  combined_plot <- wrap_plots(plot_list, ncol = ncol)

  # Add a title to the combined plot
  final_plot <- combined_plot +
    plot_annotation(
      theme = theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
                    plot.subtitle = element_text(hjust = 0.5, size = 12))
    )

  return(final_plot)
}
```

```{r}

# Usage example:
image_ids_to_visualize <- c("NTImg0001", "NTImg0002", "NTImg0003", "NTImg0004")
multiplot <- visualize_multiple_spatial_graphs1(spatial_graph1, spe_chemo_immu, image_ids = image_ids_to_visualize, node_color_by = "celltype", ncol = 2)
print(multiplot)
```

```{r}

```

### Single PatientID visualization

```{r}

visualize_patient_spatial_graph1 <- function(graph, spe_chemo_immu, patient_id, node_color_by = "celltype", edge_alpha = 0.1, node_size = 1, node_alpha = 0.7, max_edges = 10000) {
  # Check if PatientID column exists in spe
  if (!"PatientID" %in% colnames(colData(spe_chemo_immu))) {
    stop("PatientID column not found in the SpatialExperiment object.")
  }
  
  # Filter the graph for the specified PatientID
  nodes_to_keep <- which(spe_chemo_immu$PatientID == patient_id)
  subgraph <- induced_subgraph(graph, nodes_to_keep)
  
  # Remove isolated nodes
  subgraph <- delete_vertices(subgraph, degree(subgraph) == 0)
  
  # Print graph information
  cat("Full graph - Nodes:", vcount(graph), "Edges:", ecount(graph), "\n")
  cat("Patient subgraph - Nodes:", vcount(subgraph), "Edges:", ecount(subgraph), "\n")
  
  # Check if the subgraph is empty
  if (vcount(subgraph) == 0) {
    warning("No nodes found for the specified PatientID.")
    return(NULL)
  }
  
  # Subsample edges if there are too many
  if (ecount(subgraph) > max_edges) {
    set.seed(42)  # for reproducibility
    edges_to_keep <- sample(1:ecount(subgraph), max_edges)
    edge_list <- as_edgelist(subgraph)[edges_to_keep, ]
    subgraph <- graph_from_edgelist(edge_list, directed = FALSE)
    
    # Reapply vertex attributes
    vertex_attr_names <- list.vertex.attributes(graph)
    for (attr in vertex_attr_names) {
      vertex_attr(subgraph, attr) <- vertex_attr(graph, attr)[V(subgraph)]
    }
    
    cat("Subsampled edges - New edge count:", ecount(subgraph), "\n")
  }
  
  # Get node attributes
  node_attrs <- data.frame(
    celltype = V(subgraph)$celltype,
    x = V(subgraph)$x,
    y = V(subgraph)$y,
    ImageID = V(subgraph)$ImageID
  )
  
  # Create a layout based on spatial coordinates
  layout <- as.matrix(node_attrs[, c("x", "y")])
  
  # Create ggraph plot with nodes colored by the specified attribute
  p <- ggraph(subgraph, layout = layout) +
    geom_edge_link(edge_colour = "gray", edge_alpha = edge_alpha, show.legend = FALSE) +
    geom_node_point(aes(color = .data[[node_color_by]]), size = node_size, alpha = node_alpha) +
    scale_color_discrete() +
    theme_void() +
    labs(title = paste("PatientID:", patient_id),
         subtitle = paste("Number of Images:", length(unique(node_attrs$ImageID))),
         color = node_color_by) +
    coord_fixed()  # Ensure the aspect ratio is 1:1
  
  # Add faceting by ImageID if there are multiple images
  if (length(unique(node_attrs$ImageID)) > 1) {
    p <- p + facet_wrap(~ImageID)
  }
  
  return(p)
}
```

```{r}

# Usage example:
patient_id_to_visualize <- "NT008"  # Replace with your desired PatientID
plot <- visualize_patient_spatial_graph1(spatial_graph1, spe_chemo_immu, patient_id = patient_id_to_visualize, node_color_by = "celltype")
print(plot)
```

```{r}

```

#### Save Datasets

```{r}
### Chemotherapy

## Save data

#saveRDS(spe_chemo, "saved_data/spe_chemo_Spatial_Graph.rds" )
```

```{r}

### Chemotherapy

## Save data

#saveRDS(spe_chemo_immu, "saved_data/spe_chemo_immu_Spatial_Graph.rds" )
```

```{r}
  
```

```{r}

spe <- readRDS("saved_data/Spatial_Graph.rds")
spe
```

```{r}

# Subsetting based on the column 'Arm' Chemotherapy

spe_chemotherapy <- spe[, colData(spe)$Arm == "C"]
spe_chemotherapy
```

```{r}

spe_chemotherapy_Immu <- spe[, colData(spe)$Arm == "C&I"]
spe_chemotherapy_Immu
```

```{r}

#saveRDS(spe_chemotherapy, "saved_data/spe_chemo_Spatial_Graph.rds" )
```

```{r}
#saveRDS(spe_chemotherapy_Immu, "saved_data/spe_chemo_immu_Spatial_Graph.rds" )
```

```{r}

```

```         
```
