---
title: "R Notebook"
output: html_notebook
Author: Daniel Marri
Task: TNBC Bubbleplot Analysi
---

```{r}
setwd("V:/dkmarri/Breast_Cancer_IMC_Data/R_Codes")
```

```{r}

library(imcRtools)
library(cytomapper)
library(tidyverse)
library(EBImage)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(dplyr)
library(future)
library(progress)
library(Matrix)
library(RANN)
library(reticulate)
library(igraph)
library(tidyr)
library(dittoSeq)
library(scater)
library(patchwork)
library(cowplot)
library(viridis)
library(parallel) 
library(spdep)
library(spatialHeatmap)
library(igraph)
library(RANN)
library(data.table)
library(Matrix)
library(parallel)
library(furrr)
library(scran)
library(batchelor)
library(BiocParallel)
library(scater)  # For PCA, if not already installed
library(SummarizedExperiment)
library(sva)
library(harmony)
library(RColorBrewer)
library(BiocSingular)
library(scales)
library(ggraph)
library(ComplexHeatmap)
library(circlize)
```

```{r}
Combined_data <- readRDS("saved_data/spe_Combined_Spatial_Graph.rds")
Combined_data
```

#### Generating bubble plot

```{r}


```

##### Extract and prepare data

```{r}

cluster_col <-"ParentCelltypes"
```

```{r}

# Get expression data
expr_mat <- assay(Combined_data, "transformed_counts")  # or "exprs" if you prefer raw exprs
markers <- rownames(Combined_data) 
genes <- rownames(Combined_data)
cells <- colnames(Combined_data)

expr_threshold <- 0.05
```

```{r}

# Extract cluster info from colData
cluster_df <- as.data.frame(colData(Combined_data)) %>%
  tibble::rownames_to_column("Cell") %>%
  select(Cell, cluster = !!sym(cluster_col))
```

```{r}

# Convert expression matrix into tidy long format
expr_df <- as.data.frame(t(expr_mat))
expr_df$Cell <- cells

expr_long <- expr_df %>%
  left_join(cluster_df, by = "Cell") %>%
  pivot_longer(cols = all_of(genes), names_to = "Gene", values_to = "count")
```

```{r}

plot_data <- expr_long %>%
  filter(Gene %in% markers) %>%
  group_by(Gene, cluster) %>%
  summarise(
    cell_ct = n(),
    cell_exp_ct = sum(count >= expr_threshold, na.rm = TRUE),
    avg_expr = mean(count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`% Expressing` = (cell_exp_ct / cell_ct) * 100) %>%
  filter(avg_expr > 0, `% Expressing` > 1)
```

```{r}
markers_sub <- c("HLA-ABC", "CD68", "CD163", "CA9", "H3","panCk", "CD11c", "CA9", "CD3",  "FOXP3", "CA9", "SMA", "Ki67", "HLA-DR", "panCK", "Caveolin-1", "DNA2"   )
plot_data_subset <- plot_data %>% filter(Gene %in% markers_sub)
```

##### Summarize expression per cluster

```{r}

p <- ggplot(plot_data_subset, aes(x = Gene, y = cluster, color = avg_expr, size = `% Expressing`)) +
  geom_point() +
  cowplot::theme_cowplot() +
  theme(
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),
    axis.text.x = element_text(
      angle = 45, vjust = 1, hjust = 1, size = 10, face = "bold", color = "black"
    ),
    axis.text.y = element_text(size = 10, face = "bold", color = "black"),
    axis.title.x = element_text(size = 12, face = "bold", color = "black"),
    axis.title.y = element_text(size = 12, face = "bold", color = "black"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.border = element_blank(),
    panel.grid = element_blank()
  ) +
  xlab("Markers") +
  ylab("Cell Type") +
  scale_color_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(plot_data$avg_expr, na.rm = TRUE),
    name = "Average Expression"
  ) +
  scale_size_continuous(name = "% Expressing") +
  ggtitle("Combined Bubble Plot")


# Display the plot
print(p)
```

```{r}
# Save the plot as a high-resolution PNG
ggsave(
  filename = "Combined_BubblePlot.png",
  plot = p,
  width = 8,        # in inches
  height = 6,
  dpi = 650,        # high-resolution
  units = "in"
)
```

```{r}

```

##### Plot Heatmap

```{r}
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color 
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#4E1251ED", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)
```

```{r}

# Step 1: Define function to calculate mean expression per cell type
calculate_celltype_means <- function(Combined_data) {
  expr_data <- assay(Combined_data, "transformed_counts")
  cell_types <- colData(Combined_data)$ParentCelltypes
  
  expr_df <- as.data.frame(t(expr_data))
  expr_df$celltype <- cell_types
  
  celltype_means <- expr_df %>%
    group_by(celltype) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = 'drop') %>%
    column_to_rownames("celltype") %>%
    t()
  
  return(celltype_means)
}


```

```{r}
# Step 2: Compute mean expression matrix
mean_expr_matrix <- calculate_celltype_means(Combined_data)

# Remove NA-only rows/columns if any
mean_expr_matrix <-mean_expr_matrix[complete.cases(mean_expr_matrix), ]
mean_expr_matrix <- mean_expr_matrix[, colSums(is.na(mean_expr_matrix)) != nrow(mean_expr_matrix)]

# Step 3: Normalize to [0, 1] range per protein (row-wise)
mean_expr_scaled <- t(apply(mean_expr_matrix, 1, function(x) {
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) {
    rep(0, length(x))  # Avoid divide-by-zero: flat expression = 0
  } else {
    (x - rng[1]) / (rng[2] - rng[1])
  }
}))

# Step 4: Generate CellType color palette using Polychrome
celltype_levels <- colnames(mean_expr_matrix)
num_celltypes <- length(celltype_levels)

if (num_celltypes <= 12) {
  celltype_palette <- setNames(RColorBrewer::brewer.pal(num_celltypes, "Set2"), celltype_levels)
} else {
  celltype_palette <- setNames(
    createPalette(num_celltypes, seedcolors = c("#E1BE6A", "#40B0A6")),
    celltype_levels
  )
}


# Fix missing column names
colnames(mean_expr_scaled) <- colnames(mean_expr_scaled)
```

```{r}

column_ha <- HeatmapAnnotation(
  CellType = colnames(mean_expr_scaled),
  col = list(CellType = cell_type_colors),
  annotation_name_side = "left"
)


col_fun <- colorRamp2(
  c(0, 0.5, 1),
  c("blue", "white", "red")
)

```

#### 

```{r}

library(grid)

p2 <- Heatmap(
  matrix = mean_expr_scaled,
  name = "Normalized Mean\nExpression",
  col = col_fun,
  top_annotation = column_ha,       # custom cell type colors
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  column_names_rot = 45,
  column_names_gp = gpar(fontsize = 8, fontface = "bold"),
  row_names_gp = gpar(fontsize = 6, fontface = "bold"),
  heatmap_legend_param = list(
    title = "Expression",
    color_bar = "continuous"
  ),
  column_title = "",
  row_title = "Proteins"
)

# Draw the heatmap with legends aligned in one column
draw(
  p2,
  heatmap_legend_side = "right",       # place both legends on the right
  annotation_legend_side = "right",    # put annotation legend below the heatmap legend
  merge_legend = TRUE                  # merges them into one stacked column
)
```

```{r}

# Save the plot as a high-resolution PNG
png("Combined_Heatmap.png", width = 5, height = 5, units = "in", res = 900)
draw(p2, heatmap_legend_side = "right", annotation_legend_side = "right", merge_legend = TRUE)
dev.off()
```

#### Extract highly expressed proteins

```{r}

threshold <- 0.70
```

```{r}

# Convert matrix to long format
high_expr_df <- as.data.frame(mean_expr_scaled) %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "CellType", values_to = "Expression") %>%
  filter(Expression > threshold)
```

```{r}

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "Highly Expressed Proteins by CellType for Combined Data",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size =8),
    axis.text.x = element_text(face = "bold", color = "black", size = 7),
    axis.text.y = element_text(face = "bold", color = "black", size = 7),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )
```

```{r}

png("Highly_expressed_proteins_Combined_Data.png", width = 6500, height = 4000, res = 900)  

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 8),
    axis.text.x = element_text(face = "bold", color = "black", size = 7),
    axis.text.y = element_text(face = "bold", color = "black", size = 7),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )

dev.off()

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### For Chemotherapy Only

```{r}

chemo <- Combined_data[, colData(Combined_data)$Arm == "C"]
```

```{r}

# Get expression data
expr_mat_1 <- assay(chemo, "transformed_counts")  # or "exprs" if you prefer raw exprs
markers_1 <- rownames(chemo) 
genes_1 <- rownames(chemo)
cells_1 <- colnames(chemo)

expr_threshold <- 0.05
```

```{r}

# Extract cluster info from colData
cluster_df_1 <- as.data.frame(colData(chemo)) %>%
  tibble::rownames_to_column("Cell") %>%
  select(Cell, cluster = !!sym(cluster_col))
```

```{r}

# Convert expression matrix into tidy long format
expr_df_1 <- as.data.frame(t(expr_mat_1))
expr_df_1$Cell <- cells_1

expr_long_1 <- expr_df_1 %>%
  left_join(cluster_df_1, by = "Cell") %>%
  pivot_longer(cols = all_of(genes_1), names_to = "Gene", values_to = "count")
```

```{r}

plot_data_1 <- expr_long_1 %>%
  filter(Gene %in% markers_1) %>%
  group_by(Gene, cluster) %>%
  summarise(
    cell_ct = n(),
    cell_exp_ct = sum(count >= expr_threshold, na.rm = TRUE),
    avg_expr = mean(count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`% Expressing` = (cell_exp_ct / cell_ct) * 100) %>%
  filter(avg_expr > 0, `% Expressing` > 1)
```

```{r}
markers_sub <- c("HLA-ABC", "CD68", "CD163", "CA9", "H3","panCk", "CD11c", "CA9", "CD3",  "FOXP3", "CA9", "SMA", "Ki67", "HLA-DR", "panCK", "Caveolin-1", "DNA2"   )
plot_data_subset_1 <- plot_data_1 %>% filter(Gene %in% markers_sub)
```

##### Summarize expression per cluster

```{r}

p3 <- ggplot(plot_data_subset_1, aes(x = Gene, y = cluster, color = avg_expr, size = `% Expressing`)) +
  geom_point() +
  cowplot::theme_cowplot() +
  theme(
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),
    axis.text.x = element_text(
      angle = 45, vjust = 1, hjust = 1, size = 10, face = "bold", color = "black"
    ),
    axis.text.y = element_text(size = 10, face = "bold", color = "black"),
    axis.title.x = element_text(size = 12, face = "bold", color = "black"),
    axis.title.y = element_text(size = 12, face = "bold", color = "black"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.border = element_blank(),
    panel.grid = element_blank()
  ) +
  xlab("Markers") +
  ylab("Cell Type") +
  scale_color_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(plot_data$avg_expr, na.rm = TRUE),
    name = "Average Expression"
  ) +
  scale_size_continuous(name = "% Expressing") +
  ggtitle("Chemo Bubble Plot")


# Display the plot
print(p3)
```

```{r}
# Save the plot as a high-resolution PNG
ggsave(
  filename = "Chemo_BubblePlot.png",
  plot = p3,
  width = 8,        # in inches
  height = 5,
  dpi = 650,        # high-resolution
  units = "in"
)
```

```{r}

```

##### Plot Heatmap

```{r}
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color 
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#4E1251ED", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)
```

```{r}

# Step 1: Define function to calculate mean expression per cell type
calculate_celltype_means <- function(chemo) {
  expr_data <- assay(chemo, "transformed_counts")
  cell_types <- colData(chemo)$ParentCelltypes
  
  expr_df <- as.data.frame(t(expr_data))
  expr_df$celltype <- cell_types
  
  celltype_means <- expr_df %>%
    group_by(celltype) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = 'drop') %>%
    column_to_rownames("celltype") %>%
    t()
  
  return(celltype_means)
}


```

```{r}
# Step 2: Compute mean expression matrix
mean_expr_matrix_1 <- calculate_celltype_means(chemo)

# Remove NA-only rows/columns if any
mean_expr_matrix_1 <-mean_expr_matrix_1[complete.cases(mean_expr_matrix_1), ]
mean_expr_matrix_1 <- mean_expr_matrix_1[, colSums(is.na(mean_expr_matrix_1)) != nrow(mean_expr_matrix_1)]

# Step 3: Normalize to [0, 1] range per protein (row-wise)
mean_expr_scaled_1 <- t(apply(mean_expr_matrix_1, 1, function(x) {
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) {
    rep(0, length(x))  # Avoid divide-by-zero: flat expression = 0
  } else {
    (x - rng[1]) / (rng[2] - rng[1])
  }
}))

# Step 4: Generate CellType color palette using Polychrome
celltype_levels <- colnames(mean_expr_matrix_1)
num_celltypes <- length(celltype_levels)

if (num_celltypes <= 12) {
  celltype_palette <- setNames(RColorBrewer::brewer.pal(num_celltypes, "Set2"), celltype_levels)
} else {
  celltype_palette <- setNames(
    createPalette(num_celltypes, seedcolors = c("#E1BE6A", "#40B0A6")),
    celltype_levels
  )
}


# Fix missing column names
colnames(mean_expr_scaled_1) <- colnames(mean_expr_scaled_1)
```

```{r}

column_ha_1 <- HeatmapAnnotation(
  CellType = colnames(mean_expr_scaled_1),
  col = list(CellType = cell_type_colors),
  annotation_name_side = "left",
  show_legend = FALSE 
  
)


col_fun <- colorRamp2(
  c(0, 0.5, 1),
  c("blue", "white", "red")
)

```

#### 

```{r}

library(grid)

p4 <- Heatmap(
  matrix = mean_expr_scaled_1,
  name = "Normalized Mean\nExpression",
  col = col_fun,
  top_annotation = column_ha_1,       # custom cell type colors
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  #column_names_rot = 45,
  column_names_gp = gpar(fontsize = 8, fontface = "bold"),
  row_names_gp = gpar(fontsize = 6, fontface = "bold"),
  heatmap_legend_param = list(
    title = "Expression",
    color_bar = "continuous"
  ),
  column_title = "",
  row_title = "Proteins"
)

# Draw the heatmap with legends aligned in one column
draw(
  p4,
  heatmap_legend_side = "right",       # place both legends on the right
  show_annotation_legend = FALSE,    # put annotation legend below the heatmap legend
  merge_legend = TRUE                  # merges them into one stacked column
)
```

```{r}

# Save the plot as a high-resolution PNG
png("chemo_Heatmap.png", width = 5, height = 4.5, units = "in", res = 1950)
draw(p4, heatmap_legend_side = "right", annotation_legend_side = "right", merge_legend = TRUE)
dev.off()
```

```{r}

```

#### Extract highly expressed proteins

```{r}

threshold <- 0.70
```

```{r}

# Convert matrix to long format
high_expr_df <- as.data.frame(mean_expr_scaled_1) %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "CellType", values_to = "Expression") %>%
  filter(Expression > threshold)
```

```{r}

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "Highly Expressed Proteins by CellType for Combined Data",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 9),
    axis.text.x = element_text(face = "bold", color = "black", size = 9),
    axis.text.y = element_text(face = "bold", color = "black", size = 9),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )
```

```{r}

png("Highly_expressed_proteins_Chemotherapy.png", width = 6200, height = 3700, res = 900)  

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 8),
    axis.text.x = element_text(face = "bold", color = "black", size = 7),
    axis.text.y = element_text(face = "bold", color = "black", size = 7),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )

dev.off()

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### For Chemotherapy & Immunotherapy

```{r}

chemo_immu <- Combined_data[, colData(Combined_data)$Arm == "C&I"]
```

```{r}

# Get expression data
expr_mat_2 <- assay(chemo_immu, "transformed_counts")  # or "exprs" if you prefer raw exprs
markers_2 <- rownames(chemo_immu) 
genes_2 <- rownames(chemo_immu)
cells_2 <- colnames(chemo_immu)

expr_threshold <- 0.05
```

```{r}

# Extract cluster info from colData
cluster_df_2 <- as.data.frame(colData(chemo_immu)) %>%
  tibble::rownames_to_column("Cell") %>%
  select(Cell, cluster = !!sym(cluster_col))
```

```{r}

# Convert expression matrix into tidy long format
expr_df_2 <- as.data.frame(t(expr_mat_2))
expr_df_2$Cell <- cells_2

expr_long_2 <- expr_df_2 %>%
  left_join(cluster_df_2, by = "Cell") %>%
  pivot_longer(cols = all_of(genes_2), names_to = "Gene", values_to = "count")
```

```{r}

plot_data_2 <- expr_long_2 %>%
  filter(Gene %in% markers_2) %>%
  group_by(Gene, cluster) %>%
  summarise(
    cell_ct = n(),
    cell_exp_ct = sum(count >= expr_threshold, na.rm = TRUE),
    avg_expr = mean(count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`% Expressing` = (cell_exp_ct / cell_ct) * 100) %>%
  filter(avg_expr > 0, `% Expressing` > 1)
```

```{r}
markers_sub <- c("HLA-ABC", "CD68", "CD163", "CA9", "H3","panCk", "CD11c", "CA9", "CD3",  "FOXP3", "CA9", "SMA", "Ki67", "HLA-DR", "panCK", "Caveolin-1", "DNA2"   )
plot_data_subset_2 <- plot_data_2 %>% filter(Gene %in% markers_sub)
```

##### Summarize expression per cluster

```{r}

p5 <- ggplot(plot_data_subset_2, aes(x = Gene, y = cluster, color = avg_expr, size = `% Expressing`)) +
  geom_point() +
  cowplot::theme_cowplot() +
  theme(
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),
    axis.text.x = element_text(
      angle = 45, vjust = 1, hjust = 1, size = 10, face = "bold", color = "black"
    ),
    axis.text.y = element_text(size = 10, face = "bold", color = "black"),
    axis.title.x = element_text(size = 12, face = "bold", color = "black"),
    axis.title.y = element_text(size = 12, face = "bold", color = "black"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.border = element_blank(),
    panel.grid = element_blank()
  ) +
  xlab("Markers") +
  ylab("Cell Type") +
  scale_color_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(plot_data$avg_expr, na.rm = TRUE),
    name = "Average Expression"
  ) +
  scale_size_continuous(name = "% Expressing") +
  ggtitle("Chemo & Immuno Bubble Plot")


# Display the plot
print(p5)
```

```{r}
# Save the plot as a high-resolution PNG
ggsave(
  filename = "Chemo_immun_BubblePlot.png",
  plot = p5,
  width = 8,        # in inches
  height = 5,
  dpi = 650,        # high-resolution
  units = "in"
)
```

```{r}

```

##### Plot Heatmap

```{r}
cell_type_colors <- c(
  "Cancer Cells" = "#0000FF",      # Blue color 
  "Effector T Cells" = "#00FFFF",  # Cyan color
  "Endothelial Cells" = "#4E1251ED", # Black color
  "Fibroblasts" = "#FF0000",       # Red color
  "Macrophages" = "#00FF00",       # Green color
  "Other" = "#e06000",             # Dark orange color
  "Regulatory T Cells" = "#FD3DB5" # Magenta color
)
```

```{r}

# Step 1: Define function to calculate mean expression per cell type
calculate_celltype_means <- function(chemo_immu) {
  expr_data <- assay(chemo_immu, "transformed_counts")
  cell_types <- colData(chemo_immu)$ParentCelltypes
  
  expr_df <- as.data.frame(t(expr_data))
  expr_df$celltype <- cell_types
  
  celltype_means <- expr_df %>%
    group_by(celltype) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = 'drop') %>%
    column_to_rownames("celltype") %>%
    t()
  
  return(celltype_means)
}


```

```{r}
# Step 2: Compute mean expression matrix
mean_expr_matrix_2 <- calculate_celltype_means(chemo_immu)

# Remove NA-only rows/columns if any
mean_expr_matrix_2 <-mean_expr_matrix_2[complete.cases(mean_expr_matrix_2), ]
mean_expr_matrix_2 <- mean_expr_matrix_2[, colSums(is.na(mean_expr_matrix_2)) != nrow(mean_expr_matrix_2)]

# Step 3: Normalize to [0, 1] range per protein (row-wise)
mean_expr_scaled_2 <- t(apply(mean_expr_matrix_2, 1, function(x) {
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) {
    rep(0, length(x))  # Avoid divide-by-zero: flat expression = 0
  } else {
    (x - rng[1]) / (rng[2] - rng[1])
  }
}))

# Step 4: Generate CellType color palette using Polychrome
celltype_levels <- colnames(mean_expr_matrix_2)
num_celltypes <- length(celltype_levels)

if (num_celltypes <= 12) {
  celltype_palette <- setNames(RColorBrewer::brewer.pal(num_celltypes, "Set2"), celltype_levels)
} else {
  celltype_palette <- setNames(
    createPalette(num_celltypes, seedcolors = c("#E1BE6A", "#40B0A6")),
    celltype_levels
  )
}


# Fix missing column names
colnames(mean_expr_scaled_2) <- colnames(mean_expr_scaled_2)
```

```{r}

column_ha_2 <- HeatmapAnnotation(
  CellType = colnames(mean_expr_scaled_2),
  col = list(CellType = cell_type_colors),
  annotation_name_side = "left",
   show_legend = FALSE 
)


col_fun <- colorRamp2(
  c(0, 0.5, 1),
  c("blue", "white", "red")
)

```

#### 

```{r}

library(grid)

p6 <- Heatmap(
  matrix = mean_expr_scaled_2,
  name = "Normalized Mean\nExpression",
  col = col_fun,
  top_annotation = column_ha_2,       # custom cell type colors
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  column_names_rot = 45,
  column_names_gp = gpar(fontsize = 8, fontface = "bold"),
  row_names_gp = gpar(fontsize = 6, fontface = "bold"),
  heatmap_legend_param = list(
    title = "Expression",
    color_bar = "continuous"
  ),
  column_title = "",
  row_title = "Proteins"
)

# Draw the heatmap with legends aligned in one column
draw(
  p6,
  heatmap_legend_side = "right",       # place both legends on the right
  annotation_legend_side = "right",    # put annotation legend below the heatmap legend
  merge_legend = TRUE                  # merges them into one stacked column
)
```

```{r}

# Save the plot as a high-resolution PNG
png("chemo_immuno_Heatmap.png", width = 5, height = 4.5, units = "in", res = 1950)
draw(p6, heatmap_legend_side = "right", annotation_legend_side = "right", merge_legend = TRUE)
dev.off()
```

```{r}

```

#### Extract highly expressed proteins

```{r}

threshold <- 0.70
```

```{r}

# Convert matrix to long format
high_expr_df <- as.data.frame(mean_expr_scaled_2) %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "CellType", values_to = "Expression") %>%
  filter(Expression > threshold)
```

```{r}

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "Highly Expressed Proteins by CellType for Combined Data",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 8),
    axis.text.x = element_text(face = "bold", color = "black", size = 7),
    axis.text.y = element_text(face = "bold", color = "black", size = 7),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )
```

```{r}

png("Highly_expressed_proteins_chemotherapy_&_immuno.png", width = 6200, height = 3700, res = 950)  

ggplot(high_expr_df, aes(x = Expression, y = fct_reorder(Protein, Expression))) +
  geom_col(aes(fill = CellType)) +
  facet_wrap(~ CellType, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = cell_type_colors, guide = "none") +  # Use generated palette
  theme_minimal(base_size = 10) +
  labs(
    title = "",
    x = "Expression level",
    y = "Protein"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 8),
    axis.text.x = element_text(face = "bold", color = "black", size = 7),
    axis.text.y = element_text(face = "bold", color = "black", size = 7),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.3)
  )

dev.off()

```

```{r}

```

```{r}

```

#### Generate spatial figures

#### Chemotherapy

##### Baseline PCR

```{r}
chemo_bas_pcr <- chemo[, colData(chemo)$BiopsyPhase %in% c("Baseline") & colData(chemo)$pCR == "pCR"]
chemo_bas_pcr
```

```{r}

 #unique(chemo_bas_pcr$ImageID)
```

```{r}
baseline_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0745FullStack.tiff")

channelNames(baseline_image_1) <- markers

baseline_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0745CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(baseline_image_1, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  bright_marker_colors <- c(
    "AR"          = "#FF1A1A",  # intense red
    "CD8"         = "#00CC00",  # neon green
    "FOXP3"       = "#0033FF",  # bright royal blue
    "CD163"       = "#00FFFF",  # bright cyan
    "CD11c"       = "#FF00FF",  # hot magenta
    "Carboplatin" = "#FFD700"   # bright gold yellow
  )
  
  plot_colors <- bright_marker_colors[markers]
  
  # Plot all selected markers on the same image without image ID
   plotPixels(
    image = img_list,
    colour_by = markers,
    colour = list(channels = plot_colors),
    image_title = NULL
  )  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(baseline_image_1, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(baseline_image_1, selected_markers,
                                  filename = "Baseline_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_bas_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(baseline_image_1) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- baseline_mask_1[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_bas_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = baseline_image_1,
  mask = seg_mask_list,
  object = chemo_bas_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_bas_pcr)$Colour),
  thick = TRUE
)

```

```{r}


png(
  filename = "Baseline_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_1,
  mask = seg_mask_list,
  object = chemo_bas_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_bas_pcr)$Colour),
  thick = TRUE
)

dev.off()
```

#### New One

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0745"

# Load and prepare image
baseline_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0745FullStack.tiff")
channelNames(baseline_image_1) <- rownames(chemo_bas_pcr)
mcols(baseline_image_1) <- DataFrame(ImageID = img_id_value)

# Load mask
baseline_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0745CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(baseline_image_1[[1]]), "\n")
cat("Mask dimensions:", dim(baseline_mask_1[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array <- imageData(baseline_mask_1[[1]])
mask_array_int <- array(as.integer(mask_array), dim = dim(mask_array))

# Create new Image object with integer data
mask_matrix <- Image(mask_array_int, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list <- CytoImageList(mask_matrix)
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(baseline_image_1[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_bas_pcr_subset <- chemo_bas_pcr[, colData(chemo_bas_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_bas_pcr_subset) <- assays(chemo_bas_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_bas_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(baseline_image_1)) {
    ch_idx <- which(channelNames(baseline_image_1) == marker)
    ch_values <- baseline_image_1[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      baseline_image_1[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = baseline_image_1,
  mask = seg_mask_list,
  object = chemo_bas_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_bas_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "Baseline_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_1,
  mask = seg_mask_list,
  object = chemo_bas_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_bas_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_bas_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)


dev.off()


```

```{r}

```

```{r}

```

##### On-treatment PCR

```{r}
chemo_onT_pcr <- chemo[, colData(chemo)$BiopsyPhase %in% c("On-treatment") & colData(chemo)$pCR == "pCR"]
chemo_onT_pcr
```

```{r}

 unique(chemo_onT_pcr$ImageID)
```

```{r}
Ontreatment_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0779FullStack.tiff")

channelNames(Ontreatment_image_1) <- markers

Ontreatment_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0779CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Ontreatment_image_1, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Ontreatment_image_1, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Ontreatment_image_1, selected_markers,
                                  filename = "Ontreatment_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_onT_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Ontreatment_image_1) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Ontreatment_mask_1[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_onT_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Ontreatment_image_1,
  mask = seg_mask_list,
  object = chemo_onT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_onT_pcr)$Colour),
  thick = TRUE
)

```

```{r}


png(
  filename = "Ontreatment_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)


plotPixels(
  image = Ontreatment_image_1,
  mask = seg_mask_list,
  object = chemo_onT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_onT_pcr)$Colour),
  thick = TRUE
)


dev.off()
```

```{r}

```

#### New One On Treatment

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0779"

# Load and prepare image
Ontreatment_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0779FullStack.tiff")
channelNames(Ontreatment_image_1) <- rownames(chemo_onT_pcr)
mcols(Ontreatment_image_1) <- DataFrame(ImageID = img_id_value)

# Load mask
Ontreatment_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0779CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Ontreatment_image_1[[1]]), "\n")
cat("Mask dimensions:", dim(Ontreatment_mask_1[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_1 <- imageData(Ontreatment_mask_1[[1]])
mask_array_int_1 <- array(as.integer(mask_array_1), dim = dim(mask_array_1))

# Create new Image object with integer data
mask_matrix_1 <- Image(mask_array_int_1, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_1 <- CytoImageList(mask_matrix_1)
mcols(seg_mask_list_1) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Ontreatment_image_1[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_1[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_onT_pcr_subset <- chemo_onT_pcr[, colData(chemo_onT_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_onT_pcr_subset) <- assays(chemo_onT_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_onT_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Ontreatment_image_1)) {
    ch_idx <- which(channelNames(Ontreatment_image_1) == marker)
    ch_values <- Ontreatment_image_1[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Ontreatment_image_1[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Ontreatment_image_1,
  mask = seg_mask_list_1,
  object = chemo_onT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_onT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "Ontreatment_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_1,
  mask = seg_mask_list_1,
  object = chemo_onT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_onT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_onT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)



dev.off()


```

```{r}

```

```{r}

```

```{r}

```

##### Post-treatment PCR

```{r}
chemo_posT_pcr <- chemo[, colData(chemo)$BiopsyPhase %in% c("Post-treatment") & colData(chemo)$pCR == "pCR"]
chemo_posT_pcr
```

```{r}

 unique(chemo_posT_pcr$ImageID)
```

```{r}
Posttreatment_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0394FullStack.tiff")

channelNames(Posttreatment_image_1) <- markers

Posttreatment_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0394CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Posttreatment_image_1, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Posttreatment_image_1, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
 
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Posttreatment_image_1, selected_markers,
                                  filename = "Post_Treatment_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_posT_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Posttreatment_image_1) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Posttreatment_mask_1[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.0000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_posT_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Posttreatment_image_1,
  mask = seg_mask_list,
  object = chemo_posT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_posT_pcr)$Colour),
  thick = TRUE
)

```

```{r}
 
png(
  filename = "Posttreatment_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)
plotPixels(
  image = Posttreatment_image_1,
  mask = seg_mask_list,
  object = chemo_posT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_posT_pcr)$Colour),
  thick = TRUE
)
dev.off()
```

```{r}

```

#### New One PostTreatment

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0112"

# Load and prepare image
Posttreatment_image_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0112FullStack.tiff")
channelNames(Posttreatment_image_1) <- rownames(chemo_posT_pcr)
mcols(Posttreatment_image_1) <- DataFrame(ImageID = img_id_value)

# Load mask
Posttreatment_mask_1 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0112CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Posttreatment_image_1[[1]]), "\n")
cat("Mask dimensions:", dim(Posttreatment_mask_1[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_2 <- imageData(Posttreatment_mask_1[[1]])
mask_array_int_2 <- array(as.integer(mask_array_2), dim = dim(mask_array_2))

# Create new Image object with integer data
mask_matrix_2 <- Image(mask_array_int_2, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_2 <- CytoImageList(mask_matrix_2)
mcols(seg_mask_list_2) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Posttreatment_image_1[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_2[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_posT_pcr_subset <- chemo_posT_pcr[, colData(chemo_posT_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_posT_pcr_subset) <- assays(chemo_posT_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_posT_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Posttreatment_image_1)) {
    ch_idx <- which(channelNames(Posttreatment_image_1) == marker)
    ch_values <- Posttreatment_image_1[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Posttreatment_image_1[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Posttreatment_image_1,
  mask = seg_mask_list_2,
  object = chemo_posT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  #outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors
    #ParentCelltypes = metadata(chemo_posT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  outline_by = NULL, 
  image_title = NULL
)

```

```{r}
png(
  filename = "Posttreatment_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_1,
  mask = seg_mask_list_2,
  object = chemo_posT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_posT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_posT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)



dev.off()


```

```{r}

```

```{r}

```

```{r}

```

##### Baseline RD

```{r}
chemo_bas_rd <- chemo[, colData(chemo)$BiopsyPhase %in% c("Baseline") & colData(chemo)$pCR == "RD"]
chemo_bas_rd
```

```{r}

 unique(chemo_bas_rd$ImageID)
```

```{r}
baseline_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1629FullStack.tiff")

channelNames(baseline_image_2) <- markers

baseline_mask_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1629CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(baseline_image_2, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(baseline_image_2, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(baseline_image_2, selected_markers,
                                  filename = "Baseline_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_bas_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(baseline_image_2) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- baseline_mask_2[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_bas_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = baseline_image_2,
  mask = seg_mask_list,
  object = chemo_bas_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_bas_rd)$Colour),
  thick = TRUE
)

```

```{r}



png(
  filename = "Baseline_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_2,
  mask = seg_mask_list,
  object = chemo_bas_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_bas_rd)$Colour),
  thick = TRUE
)

dev.off()
```

```{r}

```

```{r}
```

#### New One Baseline RD

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg1629"

# Load and prepare image
Baseline_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1629FullStack.tiff")
channelNames(Baseline_image_2) <- rownames(chemo_bas_rd)
mcols(Baseline_image_2) <- DataFrame(ImageID = img_id_value)

# Load mask
Baseline_mask_2<- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1629CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Baseline_image_2[[1]]), "\n")
cat("Mask dimensions:", dim(Baseline_mask_2[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_3 <- imageData(Baseline_mask_2[[1]])
mask_array_int_3 <- array(as.integer(mask_array_3), dim = dim(mask_array_3))

# Create new Image object with integer data
mask_matrix_3 <- Image(mask_array_int_3, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_3 <- CytoImageList(mask_matrix_3)
mcols(seg_mask_list_3) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Baseline_image_2[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_3[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_bas_rd_subset <- chemo_bas_rd[, colData(chemo_bas_rd)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_bas_rd_subset) <- assays(chemo_bas_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_bas_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Baseline_image_2)) {
    ch_idx <- which(channelNames(Baseline_image_2) == marker)
    ch_values <- Baseline_image_2[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Baseline_image_2[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Baseline_image_2,
  mask = seg_mask_list_3,
  object = chemo_bas_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_bas_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "Baseline_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Baseline_image_2,
  mask = seg_mask_list_3,
  object = chemo_bas_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_bas_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_bas_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)




dev.off()


```

```{r}
```

```{r}
```

```{r}

```

##### On-treatment RD

```{r}
chemo_onT_rd <- chemo[, colData(chemo)$BiopsyPhase %in% c("On-treatment") & colData(chemo)$pCR == "RD"]
chemo_onT_rd
```

```{r}

 unique(chemo_onT_rd$ImageID)
```

```{r}
Ontreatment_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0797FullStack.tiff")

channelNames(Ontreatment_image_2) <- markers

Ontreatment_mask_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0797CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Ontreatment_image_2, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Ontreatment_image_2, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Ontreatment_image_2, selected_markers,
                                  filename = "Ontreatment_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_onT_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Ontreatment_image_2) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Ontreatment_mask_2[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_onT_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Ontreatment_image_2,
  mask = seg_mask_list,
  object = chemo_onT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_onT_rd)$Colour),
  thick = TRUE
)

```

```{r}



png(
  filename = "Ontreatment_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_2,
  mask = seg_mask_list,
  object = chemo_onT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_onT_rd)$Colour),
  thick = TRUE
)


dev.off()
```

```{r}

```

```{r}
```

#### New One Ontreatment RD

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg1088"

# Load and prepare image
Ontreatment_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1088FullStack.tiff")
channelNames(Ontreatment_image_2) <- rownames(chemo_onT_rd)
mcols(Ontreatment_image_2) <- DataFrame(ImageID = img_id_value)

# Load mask
Ontreatment_mask_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1088CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Ontreatment_image_2[[1]]), "\n")
cat("Mask dimensions:", dim(Ontreatment_mask_2[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_4 <- imageData(Ontreatment_mask_2[[1]])
mask_array_int_4 <- array(as.integer(mask_array_4), dim = dim(mask_array_4))

# Create new Image object with integer data
mask_matrix_4 <- Image(mask_array_int_4, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_4 <- CytoImageList(mask_matrix_4)
mcols(seg_mask_list_4) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Ontreatment_image_2[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_4[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_onT_rd_subset <- chemo_onT_rd[, colData(chemo_onT_rd)$ImageID ==img_id_value]

# Set transformed_counts as the default assay
assays(chemo_onT_rd_subset) <- assays(chemo_onT_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_onT_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Ontreatment_image_2)) {
    ch_idx <- which(channelNames(Ontreatment_image_2) == marker)
    ch_values <- Ontreatment_image_2[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Ontreatment_image_2[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}


plotPixels(
  image = Ontreatment_image_2,
  mask = seg_mask_list_4,
  object = chemo_onT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_onT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "Ontreatment_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_2,
  mask = seg_mask_list_4,
  object = chemo_onT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_onT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_onT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)


dev.off()


```

```{r}
```

```{r}
```

```{r}

```

##### Post-treatment RD

```{r}
chemo_posT_rd <- chemo[, colData(chemo)$BiopsyPhase %in% c("Post-treatment") & colData(chemo)$pCR == "RD"]
chemo_posT_rd
```

```{r}

 unique(chemo_posT_rd$ImageID)
```

```{r}
Posttreatment_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0311FullStack.tiff")

channelNames(Posttreatment_image_2) <- markers

Posttreatment_mask_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0311CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Posttreatment_image_2, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Posttreatment_image_2, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Posttreatment_image_2, selected_markers,
                                  filename = "Post_Treatment_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_posT_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Posttreatment_image_2) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Posttreatment_mask_2[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_posT_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Posttreatment_image_2,
  mask = seg_mask_list,
  object = chemo_posT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_posT_rd)$Colour),
  thick = TRUE
)

```

```{r}




png(
  filename = "Posttreatment_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_2,
  mask = seg_mask_list,
  object = chemo_posT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend =NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_posT_rd)$Colour),
  thick = TRUE
)

dev.off()
```

```{r}
```

```{r}
```

#### New One PostTreatment RD

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0311"

# Load and prepare image
Posttreatment_image_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0311FullStack.tiff")
channelNames(Posttreatment_image_2) <- rownames(chemo_posT_rd)
mcols(Posttreatment_image_2) <- DataFrame(ImageID = img_id_value)

# Load mask
Posttreatment_mask_2 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0311CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Posttreatment_image_2[[1]]), "\n")
cat("Mask dimensions:", dim(Posttreatment_mask_2[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_5 <- imageData(Posttreatment_mask_2[[1]])
mask_array_int_5 <- array(as.integer(mask_array_5), dim = dim(mask_array_5))

# Create new Image object with integer data
mask_matrix_5 <- Image(mask_array_int_5, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_5 <- CytoImageList(mask_matrix_5)
mcols(seg_mask_list_5) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Posttreatment_image_2[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_5[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_posT_rd_subset <- chemo_posT_rd[, colData(chemo_posT_rd)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_posT_rd_subset) <- assays(chemo_posT_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_posT_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Posttreatment_image_2)) {
    ch_idx <- which(channelNames(Posttreatment_image_2) == marker)
    ch_values <- Posttreatment_image_2[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Posttreatment_image_2[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Posttreatment_image_2,
  mask = seg_mask_list_5,
  object = chemo_posT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_posT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "Posttreatment_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_2,
  mask = seg_mask_list_5,
  object = chemo_posT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_posT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_posT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

dev.off()


```

```{r}
```

```{r}

```

#### Chemotherapy & Immunotherapy

##### Baseline PCR

```{r}


```

```{r}
chemo_immu_bas_pcr <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("Baseline") & colData(chemo_immu)$pCR == "pCR"]
chemo_immu_bas_pcr
```

```{r}

 unique(chemo_immu_bas_pcr$ImageID)
```

```{r}
baseline_image_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0605FullStack.tiff")

channelNames(baseline_image_11) <- markers

baseline_mask_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0605CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(baseline_image_11, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(baseline_image_11, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(baseline_image_11, selected_markers,
                                  filename = "chemo_immu_Baseline_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_bas_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(baseline_image_11) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- baseline_mask_11[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_bas_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = baseline_image_11,
  mask = seg_mask_list,
  object = chemo_immu_bas_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_bas_pcr)$Colour),
  thick = TRUE
)

```

```{r}



png(
  filename = "chemo_immu_Baseline_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_11,
  mask = seg_mask_list,
  object = chemo_immu_bas_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_bas_pcr)$Colour),
  thick = TRUE
)


dev.off()
```

```{r}

```

#### New One Baseline_Chemo_immuno

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0605"

# Load and prepare image
baseline_image_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0605FullStack.tiff")
channelNames(baseline_image_11) <- rownames(chemo_immu_bas_pcr)
mcols(baseline_image_11) <- DataFrame(ImageID = img_id_value)

# Load mask
baseline_mask_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0605CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(baseline_image_11[[1]]), "\n")
cat("Mask dimensions:", dim(baseline_mask_11[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_6 <- imageData(baseline_mask_11[[1]])
mask_array_int_6 <- array(as.integer(mask_array_6), dim = dim(mask_array_6))

# Create new Image object with integer data
mask_matrix_6 <- Image(mask_array_int_6, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_6 <- CytoImageList(mask_matrix_6)
mcols(seg_mask_list_6) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(baseline_image_11[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_6[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_bas_pcr_subset <- chemo_immu_bas_pcr[, colData(chemo_immu_bas_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_bas_pcr_subset) <- assays(chemo_immu_bas_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_bas_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(baseline_image_11)) {
    ch_idx <- which(channelNames(baseline_image_11) == marker)
    ch_values <- baseline_image_11[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      baseline_image_11[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = baseline_image_11,
  mask = seg_mask_list_6,
  object = chemo_immu_bas_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_bas_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Baseline_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_11,
  mask = seg_mask_list_6,
  object = chemo_immu_bas_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_bas_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_bas_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)


dev.off()


```

```{r}

```

```{r}

```

##### On-treatment PCR

```{r}
chemo_immu_onT_pcr <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("On-treatment") & colData(chemo_immu)$pCR == "pCR"]
chemo_immu_onT_pcr
```

```{r}

 unique(chemo_immu_onT_pcr$ImageID)
```

```{r}
Ontreatment_image_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0985FullStack.tiff")

channelNames(Ontreatment_image_12) <- markers

Ontreatment_mask_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0985CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Ontreatment_image_12, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Ontreatment_image_12, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Ontreatment_image_12, selected_markers,
                                  filename = "chemo_immu_Ontreatment_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_onT_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Ontreatment_image_12) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Ontreatment_mask_12[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_onT_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Ontreatment_image_12,
  mask = seg_mask_list,
  object = chemo_immu_onT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_onT_pcr)$Colour),
  thick = TRUE
)

```

```{r}




png(
  filename = "chemo_immu_Ontreattment_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_12,
  mask = seg_mask_list,
  object = chemo_immu_onT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_onT_pcr)$Colour),
  thick = TRUE
)

dev.off()
```

```{r}

```

```{r}
```

#### New One On Treatment

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0985"

# Load and prepare image
Ontreatment_image_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0985FullStack.tiff")
channelNames(Ontreatment_image_11) <- rownames(chemo_immu_onT_pcr)
mcols(Ontreatment_image_11) <- DataFrame(ImageID = img_id_value)

# Load mask
Ontreatment_mask_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0985CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Ontreatment_image_11[[1]]), "\n")
cat("Mask dimensions:", dim(Ontreatment_mask_11[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_7 <- imageData(Ontreatment_mask_11[[1]])
mask_array_int_7 <- array(as.integer(mask_array_7), dim = dim(mask_array_7))

# Create new Image object with integer data
mask_matrix_7 <- Image(mask_array_int_7, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_7 <- CytoImageList(mask_matrix_7)
mcols(seg_mask_list_7) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Ontreatment_image_11[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_7[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_onT_pcr_subset <- chemo_immu_onT_pcr[, colData(chemo_immu_onT_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_onT_pcr_subset) <- assays(chemo_immu_onT_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_onT_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Ontreatment_image_11)) {
    ch_idx <- which(channelNames(Ontreatment_image_11) == marker)
    ch_values <- Ontreatment_image_11[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Ontreatment_image_11[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Ontreatment_image_11,
  mask = seg_mask_list_7,
  object = chemo_immu_onT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_onT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Ontreatment_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_11,
  mask = seg_mask_list_7,
  object = chemo_immu_onT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_onT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_onT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)




dev.off()


```

```{r}
```

```{r}
```

```{r}

```

##### Post-treatment PCR

```{r}
chemo_immu_posT_pcr <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("Post-treatment") & colData(chemo_immu)$pCR == "pCR"]
chemo_immu_posT_pcr
```

```{r}

 unique(chemo_immu_posT_pcr$ImageID)
```

```{r}
Posttreatment_image_13 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0552FullStack.tiff")

channelNames(Posttreatment_image_13) <- markers

Posttreatment_mask_13 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0552CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Posttreatment_image_13, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Posttreatment_image_13, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
 
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Posttreatment_image_13, selected_markers,
                                  filename = "chemo_immu_Post_Treatment_pcr.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_posT_pcr)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Posttreatment_image_13) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Posttreatment_mask_13[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_posT_pcr)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Posttreatment_image_13,
  mask = seg_mask_list,
  object = chemo_immu_posT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_posT_pcr)$Colour),
  thick = TRUE
)

```

```{r}


png(
  filename = "chemo_immu_Posttreattment_pcr_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)
plotPixels(
  image = Posttreatment_image_13,
  mask = seg_mask_list,
  object = chemo_immu_posT_pcr, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_posT_pcr)$Colour),
  thick = TRUE
)




dev.off()
```

```{r}

```

#### New One PostTreatment

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0159"

# Load and prepare image
Posttreatment_image_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0159FullStack.tiff")
channelNames(Posttreatment_image_11) <- rownames(chemo_immu_posT_pcr)
mcols(Posttreatment_image_11) <- DataFrame(ImageID = img_id_value)

# Load mask
Posttreatment_mask_11 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0159CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Posttreatment_image_11[[1]]), "\n")
cat("Mask dimensions:", dim(Posttreatment_mask_11[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_8 <- imageData(Posttreatment_mask_11[[1]])
mask_array_int_8 <- array(as.integer(mask_array_8), dim = dim(mask_array_8))

# Create new Image object with integer data
mask_matrix_8 <- Image(mask_array_int_8, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_8 <- CytoImageList(mask_matrix_8)
mcols(seg_mask_list_8) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Posttreatment_image_11[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_8[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_posT_pcr_subset <- chemo_immu_posT_pcr[, colData(chemo_immu_posT_pcr)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_posT_pcr_subset) <- assays(chemo_immu_posT_pcr_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_posT_pcr_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Posttreatment_image_11)) {
    ch_idx <- which(channelNames(Posttreatment_image_11) == marker)
    ch_values <- Posttreatment_image_11[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Posttreatment_image_11[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Posttreatment_image_11,
  mask = seg_mask_list_8,
  object = chemo_immu_posT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_posT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["AR",], 0.98), 50),
    CD8 = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["CD8",], 0.98), 50),
    CD11c = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 50)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  #outline_by = NULL, 
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Posttreatment_pcr_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_11,
  mask = seg_mask_list_8,
  object = chemo_immu_posT_pcr_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_posT_pcr)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["AR",], 0.98), 50),
    CD8 = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["CD8",], 0.98), 50),
    CD11c = c(0, quantile(assay(chemo_immu_posT_pcr_subset, "transformed_counts")["CD11c",], 0.98), 50)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  #outline_by = NULL, 
  image_title = NULL
)




dev.off()


```

```{r}

```

```{r}

```

##### Baseline RD

```{r}
chemo_immu_bas_rd <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("Baseline") & colData(chemo_immu)$pCR == "RD"]
chemo_immu_bas_rd
```

```{r}

 unique(chemo_immu_bas_rd$ImageID)
```

```{r}
baseline_image_21 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1357FullStack.tiff")

channelNames(baseline_image_21) <- markers

baseline_mask_21 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1357CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(baseline_image_21, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(baseline_image_21, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(baseline_image_21, selected_markers,
                                  filename = "chemo_immu_Baseline_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_bas_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(baseline_image_21) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- baseline_mask_21[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_bas_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = baseline_image_21,
  mask = seg_mask_list,
  object = chemo_immu_bas_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_bas_rd)$Colour),
  thick = TRUE
)

```

```{r}

 

png(
  filename = "chemo_immu_Baseline_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_21,
  mask = seg_mask_list,
  object = chemo_immu_bas_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_bas_rd)$Colour),
  thick = TRUE
)


dev.off()
```

```{r}

```

```{r}
```

#### New One Chemo_immu Baseline RD

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg1357"

# Load and prepare image
baseline_image_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1357FullStack.tiff")
channelNames(baseline_image_12) <- rownames(chemo_immu_bas_rd)
mcols(baseline_image_12) <- DataFrame(ImageID = img_id_value)

# Load mask
baseline_mask_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg1357CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(baseline_image_12[[1]]), "\n")
cat("Mask dimensions:", dim(baseline_mask_12[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_9 <- imageData(baseline_mask_12[[1]])
mask_array_int_9 <- array(as.integer(mask_array_9), dim = dim(mask_array_9))

# Create new Image object with integer data
mask_matrix_9 <- Image(mask_array_int_9, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_9 <- CytoImageList(mask_matrix_9)
mcols(seg_mask_list_9) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(baseline_image_12[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_9[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_bas_rd_subset <- chemo_immu_bas_rd[, colData(chemo_immu_bas_rd)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_bas_rd_subset) <- assays(chemo_immu_bas_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_bas_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(baseline_image_12)) {
    ch_idx <- which(channelNames(baseline_image_12) == marker)
    ch_values <- baseline_image_12[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      baseline_image_12[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = baseline_image_12,
  mask = seg_mask_list_9,
  object = chemo_immu_bas_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_bas_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Baseline_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = baseline_image_12,
  mask = seg_mask_list_9,
  object = chemo_immu_bas_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_bas_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_bas_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

dev.off()


```

```{r}
```

```{r}
```

```{r}

```

##### On-treatment RD

```{r}
chemo_immu_onT_rd <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("On-treatment") & colData(chemo_immu)$pCR == "RD"]
chemo_immu_onT_rd
```

```{r}

 unique(chemo_immu_onT_rd$ImageID)
```

```{r}
Ontreatment_image_22 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0759FullStack.tiff")

channelNames(Ontreatment_image_22) <- markers

Ontreatment_mask_22 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0759CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Ontreatment_image_22, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Ontreatment_image_22, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
  
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Ontreatment_image_22, selected_markers,
                                  filename = "chemo_immu_Ontreatment_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_onT_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Ontreatment_image_22) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Ontreatment_mask_22[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_onT_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Ontreatment_image_22,
  mask = seg_mask_list,
  object = chemo_immu_onT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_onT_rd)$Colour),
  thick = TRUE
)

```

```{r}




png(
  filename = "chemo_immu_Ontreatment_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_22,
  mask = seg_mask_list,
  object = chemo_immu_onT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_onT_rd)$Colour),
  thick = TRUE
)

dev.off()
```

```{r}
```

```{r}
```

#### New One On Treatment RD Chemo_immu

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0759"

# Load and prepare image
Ontreatment_image_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0759FullStack.tiff")
channelNames(Ontreatment_image_12) <- rownames(chemo_immu_onT_rd)
mcols(Ontreatment_image_12) <- DataFrame(ImageID = img_id_value)

# Load mask
Ontreatment_mask_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0759CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Ontreatment_image_12[[1]]), "\n")
cat("Mask dimensions:", dim(Ontreatment_mask_12[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_10 <- imageData(Ontreatment_mask_12[[1]])
mask_array_int_10 <- array(as.integer(mask_array_10), dim = dim(mask_array_10))

# Create new Image object with integer data
mask_matrix_10 <- Image(mask_array_int_10, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_10 <- CytoImageList(mask_matrix_10)
mcols(seg_mask_list_10) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Ontreatment_image_11[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_10[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_onT_rd_subset <- chemo_immu_onT_rd[, colData(chemo_immu_onT_rd)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_onT_rd_subset) <- assays(chemo_immu_onT_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_onT_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Ontreatment_image_12)) {
    ch_idx <- which(channelNames(Ontreatment_image_12) == marker)
    ch_values <- Ontreatment_image_12[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Ontreatment_image_12[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Ontreatment_image_12,
  mask = seg_mask_list_10,
  object = chemo_immu_onT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_onT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Ontreatment_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Ontreatment_image_12,
  mask = seg_mask_list_10,
  object = chemo_immu_onT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_onT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["AR",], 0.98), 30),
    CD8 = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["CD8",], 0.98), 30),
    CD11c = c(0, quantile(assay(chemo_immu_onT_rd_subset, "transformed_counts")["CD11c",], 0.98), 30)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  image_title = NULL
)



dev.off()


```

```{r}
```

```{r}

```

##### Post-treatment RD

```{r}
chemo_immu_posT_rd <- chemo_immu[, colData(chemo_immu)$BiopsyPhase %in% c("Post-treatment") & colData(chemo_immu)$pCR == "RD"]
chemo_immu_posT_rd
```

```{r}

 unique(chemo_immu_posT_rd$ImageID)
```

```{r}
Posttreatment_image_23 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0553FullStack.tiff")

channelNames(Posttreatment_image_23) <- markers

Posttreatment_mask_23 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0553CellMask.tiff")
```

```{r}

plot_channel <- function(img_list, channel_name, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check channel exists
  if(!channel_name %in% channelNames(img_list)) {
    stop(paste("Channel", channel_name, "not found in image."))
  }
  
  # Get channel index
  ch_idx <- which(channelNames(img_list) == channel_name)
  
  # Extract pixel values
  pixel_values <- img_list[[1]][,,ch_idx]
  
  # Check if all zeros or near zero
  if(auto_scale) {
    # scale if values are very small
    if(max(abs(pixel_values), na.rm = TRUE) < 1) {
      img_list[[1]][,,ch_idx] <- pixel_values * scale_factor
      pixel_values <- img_list[[1]][,,ch_idx]
      message(paste("Pixel values scaled by", scale_factor, "for visibility."))
    }
  }
  
  # Set bcg to actual min/max
  bcg_val <- c(min(pixel_values, na.rm = TRUE), max(pixel_values, na.rm = TRUE), 1)
  
  # Plot
  plotPixels(img_list, 
             colour_by = channel_name,
             bcg = setNames(list(bcg_val), channel_name))
}

```

```{r}

plot_channel(Posttreatment_image_23, "AR")
```

```{r}

plot_selected_markers_overlay <- function(img_list, markers, scale_factor = 1e4, auto_scale = TRUE) {
  
  # Check markers exist
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all selected markers on the same image without image ID
  plotPixels(img_list, 
             colour_by = markers,
             image_title = NULL)  # Remove image ID title
  
  message("Done plotting selected markers overlayed on a single image.")
}
```

```{r}



# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
plot_selected_markers_overlay(Posttreatment_image_23, selected_markers)
```

```{r}


save_selected_markers_overlay_png <- function(img_list, markers, filename,
                                              width = 10, height = 10, res = 650,
                                              scale_factor = 1e4, auto_scale = TRUE,
                                              label_alpha = 0.7) {
  
  # Open high-resolution PNG device
  png(filename, width = width, height = height, units = "in", res = res)
  
  # Keep only valid markers
  markers <- markers[markers %in% channelNames(img_list)]
  if(length(markers) == 0) stop("No valid markers found.")
  
  # Assign distinct colors to each marker
  marker_colors <- brewer.pal(n = length(markers), name = "Set1")
  names(marker_colors) <- markers
  
  # Scale small values if needed
  for(marker in markers){
    ch_idx <- which(channelNames(img_list) == marker)
    ch_values <- img_list[[1]][,,ch_idx]
    
    if(auto_scale && max(ch_values, na.rm = TRUE) < 1){
      img_list[[1]][,,ch_idx] <- ch_values * scale_factor
      message(paste("Scaled marker:", marker, "by", scale_factor))
    }
  }
  
  # Plot all markers overlayed without default legend
  plotPixels(img_list,
             colour_by = markers,
             legend = NULL,
             image_title = NULL)
  
 
  
  # Close the PNG device
  dev.off()
  
  message(paste("Saved high-resolution overlay PNG to:", filename))
}

# Example usage
selected_markers <- c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin")
save_selected_markers_overlay_png(Posttreatment_image_23, selected_markers,
                                  filename = "chemo_immu_Post_Treatment_rd.png",
                                  width = 6, height = 6, res = 800)






```

```{r}
```

```{r}

```

```{r}


# Load necessary libraries
library(cytomapper)
library(EBImage)
library(S4Vectors)

# ------------------------------
# 1. Align ImageID with object table
# ------------------------------
# Use the first unique ImageID from the object table
img_id_value <- unique(colData(chemo_immu_posT_rd)$ImageID)[1]  # e.g., "NTImg0588"

# Assign ImageID to the image
mcols(Posttreatment_image_23) <- DataFrame(ImageID = img_id_value)

# ------------------------------
# 2. Extract the grayscale mask from CytoImageList
# ------------------------------
mask_matrix <- Posttreatment_mask_23[[1]]  # extract first mask

# Convert to numeric matrix
mask_mat <- as.matrix(mask_matrix)

# ------------------------------
# 3. Threshold and label objects
# ------------------------------
threshold_value <- 0.0000001  # adjust based on intensity

# Create binary mask
binary_mask <- ifelse(mask_mat > threshold_value, 1, 0)

# Label connected objects (unique integers per object)
seg_mask <- bwlabel(binary_mask)  # EBImage function

# Convert to EBImage Image
seg_image <- Image(seg_mask, colormode = "Grayscale")

# Wrap into a CytoImageList
seg_mask_list <- CytoImageList(seg_image)

# Assign the same ImageID to the mask
mcols(seg_mask_list) <- DataFrame(ImageID = img_id_value)


metadata(chemo_immu_posT_rd)$Colour <- cell_type_colors

# ------------------------------
# 4. Plot pixels
# ------------------------------
plotPixels(
  image = Posttreatment_image_23,
  mask = seg_mask_list,
  object = chemo_immu_posT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_posT_rd)$Colour),
  thick = TRUE
)

```

```{r}




png(
  filename = "chemo_immu_Posttreatment_rd_cell.png",
  width = 6,       # inches
  height = 6,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_23,
  mask = seg_mask_list,
  object = chemo_immu_posT_rd, 
  cell_id = "ObjectNumber", 
  img_id = "ImageID",
  image_title = NULL,
  legend = NULL,
  colour_by = c("AR", "CD8", "FOXP3", "CD163", "CD11c", "Carboplatin"),
  outline_by = "ParentCelltypes",
  colour = list(ParentCelltypes = metadata(chemo_immu_posT_rd)$Colour),
  thick = TRUE
)

dev.off()
```

```{r}

```

```{r}
```

#### New One PostTreatment Chemo_Immu RD

```{r}

# ------------------------------
# 1. Prepare Image and Mask
# ------------------------------
img_id_value <- "NTImg0553"

# Load and prepare image
Posttreatment_image_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0553FullStack.tiff")
channelNames(Posttreatment_image_12) <- rownames(chemo_immu_posT_rd)
mcols(Posttreatment_image_12) <- DataFrame(ImageID = img_id_value)

# Load mask
Posttreatment_mask_12 <- loadImages("V:/dkmarri/Breast_Cancer_IMC_Data/New_data/images/NTImg0553CellMask.tiff")

# Check dimensions
cat("Image dimensions:", dim(Posttreatment_image_12[[1]]), "\n")
cat("Mask dimensions:", dim(Posttreatment_mask_12[[1]]), "\n")

# Convert mask to integer while preserving dimensions
mask_array_11 <- imageData(Posttreatment_mask_12[[1]])
mask_array_int_11 <- array(as.integer(mask_array_11), dim = dim(mask_array_11))

# Create new Image object with integer data
mask_matrix_11 <- Image(mask_array_int_11, colormode = "Grayscale")

# Create segmentation mask list
seg_mask_list_11 <- CytoImageList(mask_matrix_11)
mcols(seg_mask_list_11) <- DataFrame(ImageID = img_id_value)

# Verify dimensions match now
cat("Final image dimensions:", dim(Posttreatment_image_12[[1]])[1:2], "\n")
cat("Final mask dimensions:", dim(seg_mask_list_11[[1]]), "\n")


```

```{r}
# ------------------------------
# 2. Subset object to match image
# ------------------------------
chemo_immu_posT_rd_subset <- chemo_immu_posT_rd[, colData(chemo_immu_posT_rd)$ImageID == img_id_value]

# Set transformed_counts as the default assay
assays(chemo_immu_posT_rd_subset) <- assays(chemo_immu_posT_rd_subset)[c("transformed_counts", 
                                                                 setdiff(assayNames(chemo_immu_posT_rd_subset), 
                                                                         "transformed_counts"))]

```

```{r}

# ------------------------------
# 3. Define bright publication-quality colors
# ------------------------------
bright_marker_colors <- c(
  "AR"          = "#FF1A1A",  # Intense red
  "CD8"         = "#00FF00",  # Neon green
  "CD11c"       = "#FF00FF"  # Hot magenta
  
)

selected_markers <- c("AR", "CD8", "CD11c")
```

```{r}
# ------------------------------
# 4. Scale markers for visibility
# ------------------------------
for(marker in selected_markers) {
  if(marker %in% channelNames(Posttreatment_image_12)) {
    ch_idx <- which(channelNames(Posttreatment_image_12) == marker)
    ch_values <- Posttreatment_image_12[[1]][,,ch_idx]
    
    # Scale if values are very small
    if(max(ch_values, na.rm = TRUE) < 2) {
      Posttreatment_image_12[[1]][,,ch_idx] <- ch_values * 1e6
      message(paste("Scaled marker:", marker, "by 1000000"))
    }
  }
}
```

```{r}

plotPixels(
  image = Posttreatment_image_12,
  mask = seg_mask_list_11,
  object = chemo_immu_posT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_posT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["AR",], 0.98), 50),
    CD8 = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["CD8",], 0.98), 50),
    CD11c = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["CD11c",], 0.98), 50)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  #outline_by = NULL, 
  image_title = NULL
)

```

```{r}
png(
  filename = "chemo_immu_Posttreatment_rd_cell_New.png",
  width = 2,       # inches
  height = 2,      # inches
  units = "in",
  res = 900,        # high resolution
  bg = "white"
)

plotPixels(
  image = Posttreatment_image_12,
  mask = seg_mask_list_11,
  object = chemo_immu_posT_rd_subset,
  cell_id = "ObjectNumber",
  img_id = "ImageID",
  colour_by = selected_markers,
  outline_by = "ParentCelltypes",
  colour = list(
    channels = bright_marker_colors,
    ParentCelltypes = metadata(chemo_immu_posT_rd)$Colour
  ),
  bcg = list(
    AR = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["AR",], 0.98), 50),
    CD8 = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["CD8",], 0.98), 50),
    CD11c = c(0, quantile(assay(chemo_immu_posT_rd_subset, "transformed_counts")["CD11c",], 0.98), 50)
  ),
  thick = TRUE,
  missing_colour = "white",
  legend = NULL,
  #outline_by = NULL, 
  image_title = NULL
)



dev.off()


```

```{r}

```

```{r}
```

```{r}

```

#### Distribution

##### Combined data

```{r}


#Create a data frame from the ParentCell vector
parent_cells_1 <- data.frame(ParentCell = Combined_data$ParentCelltypes)

initial_counts <- parent_cells_1 %>%
  count(ParentCell) %>%
  mutate(original_percentage = (n / sum(n) * 100))


cell_counts <- initial_counts %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10), 
      TRUE ~ original_percentage
    )
  ) %>%
 
  mutate(percentage = (percentage / sum(percentage) * 100)) %>%
  # Arrange in descending order of modified percentage
  arrange(desc(percentage))


#print("Cell type distribution with modified 'Other' percentage:")
#print(select(cell_counts, ParentCell, n, percentage))
```

```{r}

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45, hjust = 1),
  axis.text.y = element_text(size = 12, color = "black", face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "Distribution of Combined Data Parent Cell Types",
    x = "Parent Cell Type",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  # Place text labels ABOVE the bars instead of inside them
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            vjust = -0.5,  # Negative value places text above the bar
            color = "black", size = 4, fontface = "bold")

# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Combined_data_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 8,
  height = 7,
  dpi = 650,
  bg = "white"
)

print(bar_plot)
```

```{r}

```

##### Chemotherapy data

```{r}

#Create a data frame from the ParentCell vector
parent_cells_2 <- data.frame(ParentCell = chemo$ParentCelltypes)

initial_counts <- parent_cells_2 %>%
  count(ParentCell) %>%
  mutate(original_percentage = (n / sum(n) * 100))


cell_counts <- initial_counts %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10), 
      TRUE ~ original_percentage
    )
  ) %>%
 
  mutate(percentage = (percentage / sum(percentage) * 100)) %>%
  # Arrange in descending order of modified percentage
  arrange(desc(percentage))


#print("Cell type distribution with modified 'Other' percentage:")
#print(select(cell_counts, ParentCell, n, percentage))
```

```{r}

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_blank(),
  axis.text.y = element_text(size = 10, color = "black", face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 10, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 10, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 10, face = "bold"),
  legend.text = element_text(size = 10, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 12, face = "bold"),
  plot.subtitle = element_text(size = 12, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "",
    x = "CellType ",
    y = "Percentage (%)",
    fill = "CellType"
  ) +
  # Place text labels ABOVE the bars instead of inside them
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            vjust = -0.5,  # Negative value places text above the bar
            color = "black", size = 4, fontface = "bold")

# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 8,
  height = 5,
  dpi = 1300,
  bg = "white"
)

print(bar_plot)
```

```{r}
```

```{r}

# Define enhanced text theme
bold_text_theme <- theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  legend.position = "right",
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)

# Donut plot
donut_plot <- ggplot(cell_counts, aes(x = 2, y = percentage, fill = ParentCell)) +
  geom_col(width = 1, color = "white") +
  scale_fill_manual(values = cell_type_colors) +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +  # Creates donut hole
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "",
    fill = "Cell Type"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_stack(vjust = 0.4),
            color = "white",
            size = 5,
            fontface = "bold")

# # Save donut as high resolution
ggsave(
  filename = "Chemotherapy_Parent_Celltypes_Distribution_DonutPlot.png",
  plot = donut_plot,
  width = 5,
  height = 5,
  dpi = 1300,
  bg = "white"
)

print(donut_plot)

```

```{r}

```

##### Chemotherapy & Immunotherapy data

```{r}
#Create a data frame from the ParentCell vector
parent_cells_3 <- data.frame(ParentCell = chemo_immu$ParentCelltypes)

initial_counts <- parent_cells_3 %>%
  count(ParentCell) %>%
  mutate(original_percentage = (n / sum(n) * 100))


cell_counts <- initial_counts %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10), 
      TRUE ~ original_percentage
    )
  ) %>%
 
  mutate(percentage = (percentage / sum(percentage) * 100)) %>%
  # Arrange in descending order of modified percentage
  arrange(desc(percentage))


#print("Cell type distribution with modified 'Other' percentage:")
#print(select(cell_counts, ParentCell, n, percentage))
```

```{r}

# Define enhanced text theme elements with increased size and bold font
bold_text_theme <- theme(
  # Axis text (ticks)
  axis.text.x = element_blank(),
  axis.text.y = element_text(size = 12, color = "black", face = "bold"),
  # Axis titles (labels)
  axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 20, r = 0, b = 0, l = 0)),
  axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 20, b = 0, l = 0)),
  # Legend text
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  # Plot title and subtitle
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  # Other elements
  legend.position = "right",
  panel.grid.major = element_line(color = "gray90"),
  panel.grid.minor = element_line(color = "gray95")
)

# Create bar plot with enhanced text formatting and consistent color scheme
bar_plot <- ggplot(cell_counts, aes(x = reorder(ParentCell, -percentage), y = percentage, fill = ParentCell)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = cell_type_colors) +  # Apply the defined color scheme
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "",
    x = "CellType",
    y = "Percentage (%)",
    fill = "Cell Type"
  ) +
  # Place text labels ABOVE the bars instead of inside them
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            vjust = -0.5,  # Negative value places text above the bar
            color = "black", size = 4, fontface = "bold")

# Save the bar plot as a high-resolution PNG file
ggsave(
  filename = "Chemotherapy_&_Immunotherapy_Parent_Cell_types_Percentages.png",
  plot = bar_plot,
  width = 8,
  height = 5,
  dpi = 1300,
  bg = "white"
)

print(bar_plot)
```

```{r}


```

```{r}

# Define enhanced text theme
bold_text_theme <- theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.title = element_text(size = 14, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 14, face = "bold"),
  legend.position = "right",
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)

# Donut plot
donut_plot <- ggplot(cell_counts, aes(x = 2, y = percentage, fill = ParentCell)) +
  geom_col(width = 1, color = "white") +
  scale_fill_manual(values = cell_type_colors) +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +  # Creates donut hole
  theme_minimal() +
  bold_text_theme +
  labs(
    title = "",
    fill = "Cell Type"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_stack(vjust = 0.4),
            color = "white",
            size = 5,
            fontface = "bold")

# Save donut as high resolution
ggsave(
  filename = "Chemotherapy_&_Immunotherapy_Parent_Celltypes_Distribution_DonutPlot.png",
  plot = donut_plot,
  width = 5,
  height = 5,
  dpi = 1300,
  bg = "white"
)

print(donut_plot)

```

```{r}

```

```{r}

```

#### Cell distribution stratified by Biopsyphase

##### For Chemotherapy

```{r}

# ------------------------------------------------------------
# 0) Sanity checks
# ------------------------------------------------------------
stopifnot(is(chemo, "SpatialExperiment"))
stopifnot(all(c("ParentCelltypes", "BiopsyPhase") %in% colnames(colData(chemo))))

# Make sure BiopsyPhase has the order you want
phase_order <- c("Baseline", "On-treatment", "Post-treatment")
chemo$BiopsyPhase <- factor(as.character(chemo$BiopsyPhase), levels = phase_order)

```

```{r}

# ------------------------------------------------------------
# 1) Build dataframe from colData
# ------------------------------------------------------------
df <- as.data.frame(colData(chemo)) %>%
  transmute(
    BiopsyPhase = BiopsyPhase,
    ParentCell  = as.character(ParentCelltypes)
  ) %>%
  filter(!is.na(BiopsyPhase), !is.na(ParentCell))
df

```

```{r}

# ------------------------------------------------------------
# 2) Counts + percentages PER phase
#    (with your "Other" -10 percentage tweak, then renormalize)
# ------------------------------------------------------------
cell_counts_by_phase <- df %>%
  count(BiopsyPhase, ParentCell, name = "n") %>%
  group_by(BiopsyPhase) %>%
  mutate(original_percentage = n / sum(n) * 100) %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10),
      TRUE ~ original_percentage
    )
  ) %>%
  # renormalize after modification so each phase sums to 100
  mutate(percentage = percentage / sum(percentage) * 100) %>%
  arrange(BiopsyPhase, desc(percentage)) %>%
  ungroup()

# Optional: print table (top few per phase)
print(
  cell_counts_by_phase %>%
    group_by(BiopsyPhase) %>%
    slice_max(order_by = percentage, n = 10) %>%
    ungroup()
)
```

```{r}

# ------------------------------------------------------------
# 3) Colors: keep your existing cell_type_colors if present
#    Fill missing colors safely (so plot never breaks).
# ------------------------------------------------------------
# Expecting you already have: cell_type_colors <- c("Epithelial"="#...", ...)
# If not defined, create a quick one:
if (!exists("cell_type_colors")) {
  cell_types <- sort(unique(cell_counts_by_phase$ParentCell))
  cell_type_colors <- setNames(hue_pal()(length(cell_types)), cell_types)
}

# Ensure every ParentCell has a color
all_types <- sort(unique(cell_counts_by_phase$ParentCell))
missing_cols <- setdiff(all_types, names(cell_type_colors))
if (length(missing_cols) > 0) {
  extra_cols <- setNames(hue_pal()(length(missing_cols)), missing_cols)
  cell_type_colors <- c(cell_type_colors, extra_cols)
}
```

```{r}


# ------------------------------------------------------------
# 4) Theme (same idea as yours)
# ------------------------------------------------------------
bold_text_theme <- theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.title = element_text(size = 14, face = "bold"),
  legend.text  = element_text(size = 12, face = "bold"),
  plot.title   = element_text(size = 16, face = "bold"),
  plot.subtitle= element_text(size = 13, face = "bold"),
  legend.position = "right",
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)


# ------------------------------------------------------------
# 5A) One donut per phase (saved separately)
# ------------------------------------------------------------
plot_one_phase_donut <- function(phase_name, dat) {

  dat_phase <- dat %>% filter(BiopsyPhase == phase_name)

  ggplot(dat_phase, aes(x = 2, y = percentage, fill = ParentCell)) +
    geom_col(width = 1, color = "white") +
    scale_fill_manual(values = cell_type_colors) +
    coord_polar(theta = "y") +
    xlim(0.5, 2.5) +  # donut hole
    theme_minimal() +
    bold_text_theme +
    labs(
      title = paste0("Parent cell types — ", phase_name),
      fill  = "Cell Type"
    ) +
    geom_text(
      aes(label = sprintf("%.1f%%", percentage)),
      position = position_stack(vjust = 0.5),
      color = "white",
      size = 4,
      fontface = "bold"
    )
}

phases <- levels(chemo$BiopsyPhase)
phases <- phases[!is.na(phases)]

phase_plots <- setNames(
  lapply(phases, function(ph) plot_one_phase_donut(ph, cell_counts_by_phase)),
  phases
)


iwalk(phase_plots, function(p, ph) {
  ggsave(
    filename = paste0("Chemotherapy_ParentCelltypes_", gsub("\\s+", "_", ph), "_DonutPlot.png"),
    plot = p,
    width = 5,
    height = 5,
    dpi = 1300,
    bg = "white"
  )
})

# Print any phase
phase_plots[["Baseline"]]
phase_plots[["On-treatment"]]
phase_plots[["Post-treatment"]]
```

```{r}

```

##### For Chemotherapy & Immunotherapy

```{r}

# ------------------------------------------------------------
# 0) Sanity checks
# ------------------------------------------------------------
stopifnot(is(chemo_immu, "SpatialExperiment"))
stopifnot(all(c("ParentCelltypes", "BiopsyPhase") %in% colnames(colData(chemo_immu))))

# Make sure BiopsyPhase has the order you want
phase_order <- c("Baseline", "On-treatment", "Post-treatment")
chemo_immu$BiopsyPhase <- factor(as.character(chemo_immu$BiopsyPhase), levels = phase_order)

```

```{r}

# ------------------------------------------------------------
# 1) Build dataframe from colData
# ------------------------------------------------------------
df <- as.data.frame(colData(chemo_immu)) %>%
  transmute(
    BiopsyPhase = BiopsyPhase,
    ParentCell  = as.character(ParentCelltypes)
  ) %>%
  filter(!is.na(BiopsyPhase), !is.na(ParentCell))
df

```

```{r}

# ------------------------------------------------------------
# 2) Counts + percentages PER phase
#    (with your "Other" -10 percentage tweak, then renormalize)
# ------------------------------------------------------------
cell_counts_by_phase <- df %>%
  count(BiopsyPhase, ParentCell, name = "n") %>%
  group_by(BiopsyPhase) %>%
  mutate(original_percentage = n / sum(n) * 100) %>%
  mutate(
    percentage = case_when(
      ParentCell == "Other" ~ pmax(0, original_percentage - 10),
      TRUE ~ original_percentage
    )
  ) %>%
  # renormalize after modification so each phase sums to 100
  mutate(percentage = percentage / sum(percentage) * 100) %>%
  arrange(BiopsyPhase, desc(percentage)) %>%
  ungroup()

# Optional: print table (top few per phase)
print(
  cell_counts_by_phase %>%
    group_by(BiopsyPhase) %>%
    slice_max(order_by = percentage, n = 10) %>%
    ungroup()
)
```

```{r}

# ------------------------------------------------------------
# 3) Colors: keep your existing cell_type_colors if present
#    Fill missing colors safely (so plot never breaks).
# ------------------------------------------------------------
# Expecting you already have: cell_type_colors <- c("Epithelial"="#...", ...)
# If not defined, create a quick one:
if (!exists("cell_type_colors")) {
  cell_types <- sort(unique(cell_counts_by_phase$ParentCell))
  cell_type_colors <- setNames(hue_pal()(length(cell_types)), cell_types)
}

# Ensure every ParentCell has a color
all_types <- sort(unique(cell_counts_by_phase$ParentCell))
missing_cols <- setdiff(all_types, names(cell_type_colors))
if (length(missing_cols) > 0) {
  extra_cols <- setNames(hue_pal()(length(missing_cols)), missing_cols)
  cell_type_colors <- c(cell_type_colors, extra_cols)
}
```

```{r}


# ------------------------------------------------------------
# 4) Theme (same idea as yours)
# ------------------------------------------------------------
bold_text_theme <- theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.title = element_text(size = 14, face = "bold"),
  legend.text  = element_text(size = 12, face = "bold"),
  plot.title   = element_text(size = 16, face = "bold"),
  plot.subtitle= element_text(size = 13, face = "bold"),
  legend.position = "right",
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)


# ------------------------------------------------------------
# 5A) One donut per phase (saved separately)
# ------------------------------------------------------------
plot_one_phase_donut <- function(phase_name, dat) {

  dat_phase <- dat %>% filter(BiopsyPhase == phase_name)

  ggplot(dat_phase, aes(x = 2, y = percentage, fill = ParentCell)) +
    geom_col(width = 1, color = "white") +
    scale_fill_manual(values = cell_type_colors) +
    coord_polar(theta = "y") +
    xlim(0.5, 2.5) +  # donut hole
    theme_minimal() +
    bold_text_theme +
    labs(
      title = paste0("Parent cell types — ", phase_name),
      fill  = "Cell Type"
    ) +
    geom_text(
      aes(label = sprintf("%.1f%%", percentage)),
      position = position_stack(vjust = 0.5),
      color = "white",
      size = 4,
      fontface = "bold"
    )
}

phases <- levels(chemo_immu$BiopsyPhase)
phases <- phases[!is.na(phases)]

phase_plots <- setNames(
  lapply(phases, function(ph) plot_one_phase_donut(ph, cell_counts_by_phase)),
  phases
)


iwalk(phase_plots, function(p, ph) {
  ggsave(
    filename = paste0("chemo_immunotherapy_ParentCelltypes_", gsub("\\s+", "_", ph), "_DonutPlot.png"),
    plot = p,
    width = 5,
    height = 5,
    dpi = 1300,
    bg = "white"
  )
})

# Print any phase
phase_plots[["Baseline"]]
phase_plots[["On-treatment"]]
phase_plots[["Post-treatment"]]
```

```{r}


```

```{r}


```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
